---
title: "battle_clear-clear-data-checks"
author: "Inigo Peng, edited by Badhia Yunes-Katz"
date: "`r Sys.Date()`"
output: html_document
---


```{r include=FALSE}
library(DBI)
library(tidyverse)
library(EDIutils)
library(EML)
library(readr)
library(janitor)
load("filter_params.RData")
```

```{r include=FALSE}
con <- DBI::dbConnect(
  RPostgres::Postgres(),
  dbname = "jpedb-staging",
  host = "jpe-db.postgres.database.azure.com",
  port = 5432,
  user = Sys.getenv("jpe_db_user_id"),
  password = Sys.getenv("jpe_db_password")
)
```
# Database Tables
```{r echo=FALSE}
catch <- DBI::dbReadTable(con, "catch")
trap_visit <- DBI::dbReadTable(con, "trap_visit")
release <- DBI::dbReadTable(con, "release")
recaptured_fish <- DBI::dbReadTable(con, "recaptured_fish")
trap_location <- DBI::dbReadTable(con, "trap_location")
released_fish <- DBI::dbReadTable(con, "released_fish")
```

```{r include=FALSE}
trap_id <- trap_location |> 
  rename(trap_location_id = id) |>
  select(trap_location_id, stream, site, subsite) 

catch_with_stream <- catch |>
  left_join(trap_id)
```

## battle_clear {.tabset}
 
```{r eval=FALSE, include=FALSE}
#TODO, there is no recapture in database
recapture_with_stream <- recaptured_fish |>
  left_join(trap_id)
``` 

### catch

**- Findings catch:**

  - a lot of "not provided" values on EDI lifestage

```{r echo=FALSE}
catch_battle_clear_edi <- read_csv("data-raw/edi-zips/battle_clear_catch.csv") |>
  filter(common_name == "Chinook Salmon") |>
  mutate(stream="battle clear creek",
         date = as.Date(sample_date))
```

#### Date Range Check
 
*EDI `date`*
```{r}
range(catch_battle_clear_edi$date)
```

*DB `date`*
  
```{r echo=FALSE}
catch_battle_clear_db <- catch_with_stream |> 
  filter(stream == "battle creek" | stream == "clear creek")  

range(catch_battle_clear_db$date)
```

#### ForkLength
*EDI `fork_length`*

* `r round(sum(is.na(catch_battle_clear_edi$fork_length))/nrow(catch_battle_clear_edi), 3)*100` % of values in the `fork_length` column are NA.

```{r echo=FALSE}
summary(catch_battle_clear_edi$fork_length, na.rm = TRUE)
```

*DB `fork_length`*

* `r round(sum(is.na(catch_battle_clear_db$fork_length))/nrow(catch_battle_clear_db), 3)*100` % of values in the `fork_length` column are NA.

```{r echo=FALSE}
summary(catch_battle_clear_db$fork_length, na.rm = TRUE)
```

```{r include=FALSE}
catch_battle_clear_edi_edited <- catch_battle_clear_edi |>
  clean_names() |> 
  mutate(source = "EDI") 
 
catch_battle_clear_db_edited <- catch_battle_clear_db |>
  mutate(source = "DB")
```

*Forklength Exploration Plots* 

```{r echo=FALSE, warning=FALSE}
combined_catch_battle_clear <- catch_battle_clear_db_edited |> 
  select(date, fork_length, source) |> 
  bind_rows(catch_battle_clear_edi_edited |> select(date, fork_length, source)) 

ggplot(data = combined_catch_battle_clear, aes(x = date, y = fork_length, color = source)) +
  geom_point() +
  facet_wrap(~source)
```

#### Lifestage
(life stage in edi data has not gone through wrangling)

*EDI `life_stage`*

* `r round(sum(is.na(catch_battle_clear_edi$life_stage))/nrow(catch_battle_clear_edi), 3)*100` % of values in the `life_stage` column from the EDI data are NA.

*DB `lifestage_id`*

* `r round(sum(is.na(catch_battle_clear_db$lifestage_id))/nrow(catch_battle_clear_db), 3)*100` % of values in the `lifestage_id` column from the database are NA.

```{r echo=FALSE, warning=FALSE}
catch_battle_clear_db |>
   ggplot(aes(x = as.factor(lifestage_id), y = fork_length, fill = as.factor(lifestage_id))) +
  geom_boxplot(outlier.shape = 16, outlier.size = 1, alpha = 0.7) +
  coord_flip() +
  theme_minimal() +
  labs(x = "lifestage id", y = "fork length", title = "DB data - lifestage vs forklength")
```


```{r echo=FALSE, warning=FALSE}
catch_battle_clear_edi |>
   ggplot(aes(x = as.factor(life_stage), y = fork_length, fill = as.factor(life_stage))) +
  geom_boxplot(outlier.shape = 16, outlier.size = 1, alpha = 0.7) +
  coord_flip() +
  theme_minimal() +
  labs(x = "life_stage", y = "fork_length", title = "EDI data - life_stage vs forklength")
```

#### Count
TODO: EDI data contains lifestage value "mixed" but not in lifestage lookuptable (resolved)

*EDI `count`*

* `r round(sum(is.na(catch_battle_clear_edi$count))/nrow(catch_battle_clear_edi), 3)*100` % of values in the `count` column in edi data are NA.

```{r echo=FALSE}
summary(catch_battle_clear_edi$count)
```

*DB `count`*

* `r round(sum(is.na(catch_battle_clear_db$count))/nrow(catch_battle_clear_db), 3)*100` % of values in the `count` column in database data are NA.

```{r echo=FALSE}
summary(catch_battle_clear_db$count)
```

```{r echo=FALSE, warning=FALSE}
ggplot(catch_battle_clear_db_edited, aes(x = date, y = count, color = source)) +
  geom_jitter(alpha = 0.6, width = 0.3, height = 0) +
  theme_minimal() +
  labs(x = "date", y = "count", color = "Source")

ggplot() +
  geom_jitter(data = catch_battle_clear_db_edited,
              aes(x = date, y = count),
              color = "blue", alpha = 0.4, width = 10, height = 0.5, shape = 16, size = 2) +
  geom_jitter(data = catch_battle_clear_edi_edited,
              aes(x = date, y = count),
              color = "red", alpha = 0.4, width = 10, height = 0.5, shape = 17, size = 2) +
  theme_minimal() +
  labs(x = "Date",
       y = "count/n")
```

Random pick of data entry to compare between db and EDI

- test 1

```{r echo=FALSE}
# random pick to compare data
catch_battle_clear_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "{292A0C58-FB5F-11D4-B974006008C01BCF}") |>
  glimpse()

catch_battle_clear_edi |>
  mutate(source = "EDI") |>
  filter(catch_row_id == "{292A0C58-FB5F-11D4-B974006008C01BCF}") |>
  glimpse()
```

- test 2

```{r echo=FALSE}
# random pick to compare data
catch_battle_clear_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id  == "{2929F03F-FB5F-11D4-B974006008C01BCF}") |>
  glimpse()

catch_battle_clear_edi |>
  mutate(source = "EDI") |>
  filter(catch_row_id == "{2929F03F-FB5F-11D4-B974006008C01BCF}") |>
  glimpse()

```

- test 3

```{r echo=FALSE}
# random pick to compare data
catch_battle_clear_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id  == "{4B85B5E1-9066-496A-A7F52185657FBFDE}") |>
  glimpse()

catch_battle_clear_edi |>
  mutate(source = "EDI") |>
  filter(catch_row_id == "{4B85B5E1-9066-496A-A7F52185657FBFDE}") |>
  glimpse()
```

### Trap

**- Findings trap:**


```{r echo=FALSE}
trap_with_stream <- trap_visit |>
  left_join(trap_id)

trap_battle_clear_db <- trap_with_stream |> 
  filter(stream == "battle creek" | stream == "clear creek") 
```

```{r}
trap_battle_clear_edi <- read_csv("data-raw/edi-zips/battle_clear_trap.csv") |> 
  mutate(stream="battle clear creek",
         date = as.Date(sample_date)) |>
  filter(!is.na(station_code)) |> 
  arrange(trap_start_date) |>          
  group_by(sample_id, station_code) |>
  slice_tail(n = 1)  |>                   
  ungroup() |> glimpse()
```

#### Date Range Check

*DB `trap_visit_time_start`* 
```{r echo=FALSE}
range(trap_battle_clear_db$trap_visit_time_start, na.rm = TRUE)
```
*EDI `sample_date`*
```{r}
range(trap_battle_clear_edi$sample_date)
```

#### water_velocity

*DB `water_velocity`*

* `r round(sum(is.na(trap_battle_clear_db$water_velocity))/nrow(trap_battle_clear_db), 3)*100` % of values in the `water_velocity` column are NA.
```{r echo=FALSE}
summary(trap_battle_clear_db$water_velocity)
```

*EDI `velocity`*
* `r round(sum(is.na(trap_battle_clear_edi$velocity))/nrow(trap_battle_clear_edi), 3)*100` % of values in the `velocity` column are NA.
```{r echo=FALSE}
summary(trap_battle_clear_edi$velocity)
```

```{r}
trap_battle_clear_edi_clean<- trap_battle_clear_edi |>
  mutate(date = as.Date(sample_date),
         source = "EDI")
trap_battle_clear_db_clean <- trap_battle_clear_db |>
  mutate(date = as.Date(trap_visit_time_start),
         source = "DB")

ggplot() +
  geom_jitter(data = trap_battle_clear_db_clean,
              aes(x = date, y = water_velocity, color = source),
              color = "blue", alpha = 0.4, width = 10, height = 0.5, shape = 16, size = 2) +
  geom_jitter(data = trap_battle_clear_edi_clean,
              aes(x = date, y = velocity),
              color = "red", alpha = 0.4, width = 10, height = 0.5, shape = 17, size = 2) +
  theme_minimal() +
  labs(x = "Date",
       y = "water_velocity or velocity",
       color = "Source"
       )


```

### release

**- Findings release:**

  - date ranges are the same
  - all lifestage_id are NA in DB, and no similar field in EDI

```{r echo=FALSE}
release_with_stream <- release |>
  left_join(trap_id)

release_battle_clear_db <- release_with_stream |> 
  filter(stream == "battle creek" | stream == "clear creek")
```

```{r echo=FALSE}
release_battle_clear_edi <- read_csv("data-raw/edi-zips/battle_clear_release.csv") |>
  # filter(common_name == "Chinook Salmon") |>
  mutate(stream="battle clear creek",
         date = as.Date(date_released)) 
```

#### Date Range Check

*DB `date_released`* 
```{r echo=FALSE}
range(release_battle_clear_db$date_released, na.rm = TRUE)
```
*EDI `releaseTime`*
```{r}
range(release_battle_clear_edi$date)
```


#### markedLifeStage

*EDI `* - no similar field

*DB `lifestage_id`*

* `r round(sum(is.na(release_battle_clear_db$lifestage_id))/nrow(release_battle_clear_db), 3)*100` % of values in the `lifestage_id` column are NA.

```{r echo=FALSE}
summary(release_battle_clear_db$lifestage_id, na.rm = TRUE)
```

#### nReleased

*EDI `number_released` * 

* `r round(sum(is.na(release_battle_clear_edi$number_released))/nrow(release_battle_clear_edi), 3)*100` % of values in the `number_released` column are NA.

```{r echo=FALSE}
summary(release_battle_clear_edi$number_released, na.rm = TRUE)
```

*DB `number_released`*

* `r round(sum(is.na(release_battle_clear_db$number_released))/nrow(release_battle_clear_db), 3)*100` % of values in the `number_released` column are NA.

```{r echo=FALSE}
summary(release_battle_clear_db$number_released, na.rm = TRUE)
```

#### median_fork_length_released

*EDI `median_fork_length_released`*

* `r round(sum(is.na(release_battle_clear_edi$median_fork_length_released))/nrow(release_battle_clear_edi), 3)*100` % of values in the `median_fork_length_released` column are NA.

```{r echo=FALSE}
summary(release_battle_clear_edi$median_fork_length_released)
```

*DB `median_fork_length_released`*

* `r round(sum(is.na(release_battle_clear_db$median_fork_length_released))/nrow(release_battle_clear_db), 3)*100` % of values in the `median_fork_length_released` column are NA.

```{r echo=FALSE}
summary(release_battle_clear_db$median_fork_length_released, na.rm = TRUE)
```

*Exploration Plots* 

```{r echo=FALSE, warning=FALSE}
release_battle_clear_edi_clean <- release_battle_clear_edi |>
  mutate(date = as.Date(date_released),
         source = "EDI")
release_battle_clear_db_clean <- release_battle_clear_db |>
  mutate(date = as.Date(date_released),
         source = "DB")

ggplot() +
  geom_jitter(data = release_battle_clear_db_clean,
              aes(x = date, y = number_released, color = source),
              color = "blue", alpha = 0.4, width = 10, height = 0.5, shape = 16, size = 2) +
  geom_jitter(data = release_battle_clear_edi_clean,
              aes(x = date, y = number_released),
              color = "red", alpha = 0.4, width = 10, height = 0.5, shape = 17, size = 2) +
  theme_minimal() +
  labs(x = "Date",
       y = "number_released or nReleased",
       color = "Source"
       )
```

### releasefish

**- Findings releasefish:**

  - no release fish table on EDI or DB
  
### releasefish

**- Findings recapture:**

  - no release recapture table in DB  
