---
title: "butte-test"
author: "Inigo Peng"
date: "`r Sys.Date()`"
output: html_document
---

```{r include=FALSE}
library(DBI)
library(tidyverse)
library(EDIutils)
library(EML)
library(readr)
library(janitor)
load("filter_params.RData")
```

```{r include=FALSE}
con <- DBI::dbConnect(
  RPostgres::Postgres(),
  dbname = "jpedb-staging",
  host = "jpe-db.postgres.database.azure.com",
  port = 5432,
  user = Sys.getenv("jpe_db_user_id"),
  password = Sys.getenv("jpe_db_password")
)
```
# Database Tables
```{r echo=FALSE}
catch <- DBI::dbReadTable(con, "catch")
trap_visit <- DBI::dbReadTable(con, "trap_visit")
release <- DBI::dbReadTable(con, "release")
recaptured_fish <- DBI::dbReadTable(con, "recaptured_fish")
trap_location <- DBI::dbReadTable(con, "trap_location")
```

```{r include=FALSE}
trap_id <- trap_location |> 
  rename(trap_location_id = id) |>
  select(trap_location_id, stream, site, subsite) 

catch_with_stream <- catch |>
  left_join(trap_id) 
```

## Butte {.tabset}

### catch

**- Findings catch:**

  - date range is the same 
  - life stage in edi data has not gone through wrangling

```{r echo=FALSE}
catch_butte_edi <- read_csv("data-raw/edi-zips/butte_catch.csv") |> 
  filter(siteName != "coleman national fish hatchery") |>
  filter(commonName == "Chinook salmon") |>
  mutate(siteName = tolower(siteName),
         subSiteName = tolower(subSiteName)) |> 
  mutate(
    siteName = case_when(
    siteName %in% c("okie rst", "parrot-phelan", "parrott-phelan canal trap box", "parrot-phelan rst") ~ "okie dam",
    TRUE ~ siteName
    ),
    subSiteName = case_when(
      subSiteName == "pp rst" ~ "okie dam 1",
      subSiteName == "pp rst 2" ~ "okie dam 2",
      subSiteName == "okie rst" ~ NA_character_,
      subSiteName == "canal trap box" ~ "okie dam fyke trap",
      TRUE ~ subSiteName)) |> 
  mutate(stream="butte creek",
         date = as.Date(visitTime))

filter_params_butte <- filter_params$`butte creek`

for (param in filter_params_butte){
  catch_butte_edi <- subset(catch_butte_edi, 
      !(stream == "butte creek" &
        (
          (is.na(param$site) & is.na(siteName)) |
          (!is.na(param$site) & siteName == param$site)
        ) &
        (
          (is.na(param$subsite) & is.na(subSiteName)) |
          (!is.na(param$subsite) & subSiteName == param$subsite)
        ) &
        date >= as.Date(param$start_date) &
        date <= as.Date(param$end_date)
      )
    )  
}
```

#### Date Range Check
*EDI `visitTime`*
```{r}
range(catch_butte_edi$visitTime)
```

*DB `date`*
  
```{r echo=FALSE}
catch_butte_db <- catch_with_stream |> 
  filter(stream == "butte creek") 
  # filter(is.na(site) & is.na(subsite)) |> glimpse() 

range(catch_butte_db$date)
```

#### ForkLength
*EDI `forkLength`*

* `r round(sum(is.na(catch_butte_edi$forkLength))/nrow(catch_butte_edi), 3)*100` % of values in the `forkLength` column are NA.

```{r echo=FALSE}
summary(catch_butte_edi$forkLength, na.rm = TRUE)
```

*DB `fork_length`*

* `r round(sum(is.na(catch_butte_db$fork_length))/nrow(catch_butte_db), 3)*100` % of values in the `fork_length` column are NA.

```{r echo=FALSE}
summary(catch_butte_db$fork_length, na.rm = TRUE)
```

```{r include=FALSE}
catch_butte_edi_edited <- catch_butte_edi |>
  clean_names() |>
  # mutate(date = as.Date(visit_time)) |> 
  mutate(source = "EDI")
  # filter(date > ymd("1999-06-04")) 
 
catch_butte_db_edited <- catch_butte_db |>
  mutate(source = "DB")
```

*Forklength Exploration Plots* 

```{r echo=FALSE, warning=FALSE}
combined_catch_butte <- catch_butte_db_edited |> 
  select(date, fork_length, source) |> 
  bind_rows(catch_butte_edi_edited |> select(date, fork_length, source)) 

ggplot(data = combined_catch_butte, aes(x = date, y = fork_length, color = source)) +
  geom_point() +
  facet_wrap(~source)
```

#### Lifestage
(life stage in edi data has not gone through wrangling)

*EDI `lifeStage`*

* `r round(sum(is.na(catch_butte_edi$lifeStage))/nrow(catch_butte_edi), 3)*100` % of values in the `lifeStage` column from the EDI data are NA.

*DB `lifestage_id`*

* `r round(sum(is.na(catch_butte_db$lifestage_id))/nrow(catch_butte_db), 3)*100` % of values in the `lifestage_id` column from the database are NA.

```{r echo=FALSE, warning=FALSE}
catch_butte_db |>
   ggplot(aes(x = as.factor(lifestage_id), y = fork_length, fill = as.factor(lifestage_id))) +
  geom_boxplot(outlier.shape = 16, outlier.size = 1, alpha = 0.7) +
  coord_flip() +
  theme_minimal() +
  labs(x = "lifestage id", y = "fork length", title = "DB data - lifestage vs forklength")
```


```{r echo=FALSE, warning=FALSE}
catch_butte_edi |>
   ggplot(aes(x = as.factor(lifeStage), y = forkLength, fill = as.factor(lifeStage))) +
  geom_boxplot(outlier.shape = 16, outlier.size = 1, alpha = 0.7) +
  coord_flip() +
  theme_minimal() +
  labs(x = "lifestage id", y = "fork length", title = "EDI data - lifestage vs forklength")
```

#### Count

*EDI `n`*

* `r round(sum(is.na(catch_butte_edi$n))/nrow(catch_butte_edi), 3)*100` % of values in the `n` column in edi data are NA.

```{r echo=FALSE}
summary(catch_butte_edi$n)
```

*DB `count`*

* `r round(sum(is.na(catch_butte_db$count))/nrow(catch_butte_db), 3)*100` % of values in the `count` column in database data are NA.

```{r echo=FALSE}
summary(catch_butte_db$count)
```

```{r echo=FALSE, warning=FALSE}
# ggplot() +
  # geom_jitter(data = catch_butte_db_edited,
  #             aes(x = date, y = count, color = "Database"),
  #             alpha = 0.4, width = 10, height = 0.5, shape = 16, size = 2) +
  # geom_jitter(data = catch_butte_edi_edited,
  #             aes(x = date, y = n, color = "EDI"),
  #             alpha = 0.4, width = 10, height = 0.5, shape = 17, size = 2) +
  # scale_color_manual(values = c("Database" = "blue", "EDI" = "red")) +
  # theme_minimal() +
  # labs(x = "Date", y = "Count", color = "Source")

```

Random pick of data entry to compare between db and EDI

- test 1

```{r echo=FALSE}
# random pick to compare data
catch_butte_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "55171") |>
  glimpse()

catch_butte_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "55171") |>
  glimpse()
```

- test 2

```{r echo=FALSE}
# random pick to compare data
catch_butte_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id  == "518") |>
  glimpse()

catch_butte_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == 518) |>
  glimpse()

```

- test 3

```{r echo=FALSE}
# random pick to compare data
catch_butte_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id  == "51525") |>
  glimpse()

catch_butte_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "51525") |>
  glimpse()
```

### trap

**- Findings:**

```{r echo=FALSE}
trap_with_stream <- trap_visit |>
  left_join(trap_id)

trap_butte_db <- trap_with_stream |> 
  filter(stream == "butte creek") |> 
  filter(!is.na(site) & !is.na(subsite))
```

```{r echo=FALSE}
trap_butte_edi <- read_csv("data-raw/edi-zips/butte_trap.csv") |>
  mutate(siteName = tolower(siteName),
         subSiteName = tolower(subSiteName)) |> 
  mutate(
    siteName = case_when(
    siteName %in% c("okie rst", "parrot-phelan", "parrott-phelan canal trap box", "parrot-phelan rst") ~ "okie dam",
    TRUE ~ siteName
    ),
    subSiteName = case_when(
      subSiteName == "pp rst" ~ "okie dam 1",
      subSiteName == "pp rst 2" ~ "okie dam 2",
      subSiteName == "okie rst" ~ NA_character_,
      subSiteName == "canal trap box" ~ "okie dam fyke trap",
      TRUE ~ subSiteName)) |>
  mutate(stream="butte creek",
         visit_time=as.Date(visitTime))

filter_params_butte <- filter_params$`butte creek`
# 
for (param in filter_params_butte){
  trap_butte_edi <- subset(trap_butte_edi,
      !(stream == "butte creek" &
        (
          (is.na(param$site) & is.na(siteName)) |
          (!is.na(param$site) & siteName == param$site)
        ) &
        (
          (is.na(param$subsite) & is.na(subSiteName)) |
          (!is.na(param$subsite) & subSiteName == param$subsite)
        ) &
        (visit_time >= as.Date(param$start_date) &
        visit_time <= as.Date(param$end_date))
      )
    )
}
```

#### Date Range Check

*DB `trap_visit_time_start`* 
```{r echo=FALSE}
range(trap_butte_db$trap_visit_time_start, na.rm = TRUE)
```
*EDI `visitTime`*
```{r}
range(trap_butte_edi$visitTime)
```

#### water_velocity

*DB `water_velocity`*

* `r round(sum(is.na(trap_butte_db$water_velocity))/nrow(trap_butte_db), 3)*100` % of values in the `water_velocity` column are NA.
```{r echo=FALSE}
summary(trap_butte_db$water_velocity)
```

*EDI `waterVel`*
* `r round(sum(is.na(trap_butte_edi$waterVel))/nrow(trap_butte_edi), 3)*100` % of values in the `water_velocity` column are NA.
```{r echo=FALSE}
summary(trap_butte_edi$waterVel)
```

#### fishProcessed

*DB `fish_processed_id`*

* `r round(sum(is.na(trap_butte_db$fish_processed_id))/nrow(trap_butte_db), 3)*100` % of values in the `fish_processed_id` column are NA.
```{r echo=FALSE}
table(trap_butte_db$fish_processed_id)
```

*EDI `fishProcessed`*
* `r round(sum(is.na(trap_butte_edi$fishProcessed))/nrow(trap_butte_edi), 3)*100` % of values in the `fishProcessed` column are NA.
```{r echo=FALSE}
table(trap_butte_edi$fishProcessed)
```

Random pick of data entry by date to compare between db and EDI

  - test 1
```{r echo=FALSE}
# random pick to compare data
trap_butte_db |>
  mutate(source = "DB",
         trap_visit_time_start = as.Date(trap_visit_time_start)) |>
  # filter(id == "2217421") |>
  filter(trap_visit_time_start == "2017-11-10") |>
  glimpse()

trap_butte_edi |>
  mutate(source = "EDI",
         visitTime = as.Date(visitTime)) |>
  filter(visitTime == "2017-11-10") |>
  glimpse()
```

  - test 2
```{r echo=FALSE}
# random pick to compare data
trap_butte_db |>
  mutate(source = "DB") |>
  filter(trap_visit_id  == "2144") |>
  glimpse()

trap_butte_edi |>
  mutate(source = "EDI") |>
  filter(trapVisitID == "2144") |>
  glimpse()
```

  - test 3
```{r echo=FALSE}
# random pick to compare data
trap_butte_db |>
  mutate(source = "DB") |>
  filter(trap_visit_id  == "2327") |>
  glimpse()

trap_butte_edi |>
  mutate(source = "EDI") |>
  filter(trapVisitID == "2327") |>
  glimpse()
```

  - test 4
```{r echo=FALSE}
# random pick to compare data
trap_butte_db |>
  mutate(source = "DB") |>
  filter(trap_visit_id  == "3151") |>
  glimpse()

trap_butte_edi |>
  mutate(source = "EDI") |>
  filter(trapVisitID == "3151") |>
  glimpse()
```

### release

**- Findings:**

   - time ranges are exactly the same
   - all lifestage_id's are NA in DB
   
```{r echo=FALSE}
release_with_stream <- release |>
  left_join(trap_id)

release_butte_db <- release_with_stream |> 
  filter(stream == "butte creek")
```
```{r echo=FALSE}
release_butte_edi <- read_csv("data-raw/edi-zips/butte_release.csv") |>
  filter(commonName == "Chinook salmon") |>
  mutate(releaseSite = tolower(releaseSite),
         releaseSubSite = tolower(releaseSubSite)) |> 
  mutate(
    releaseSite = case_when(
    releaseSite %in% c("okie rst", "parrot-phelan", "parrott-phelan canal trap box", "parrot-phelan rst") ~ "okie dam",
    TRUE ~ releaseSite
    ),
    releaseSubSite = case_when(
      releaseSubSite == "pp rst" ~ "okie dam 1",
      releaseSubSite == "pp rst 2" ~ "okie dam 2",
      releaseSubSite == "okie rst" ~ NA_character_,
      releaseSubSite == "canal trap box" ~ "okie dam fyke trap",
      TRUE ~ releaseSubSite)) |>
  mutate(stream="butte creek",
         date_released=as.Date(releaseTime))

filter_params_butte <- filter_params$`butte creek`
# 
for (param in filter_params_butte){
  release_butte_edi <- subset(release_butte_edi,
      !(stream == "butte creek" &
        (
          (is.na(param$site) & is.na(releaseSite)) |
          (!is.na(param$site) & releaseSite == param$site)
        ) &
        (
          (is.na(param$subsite) & is.na(releaseSubSite)) |
          (!is.na(param$subsite) & releaseSubSite == param$subsite)
        ) &
        (date_released >= as.Date(param$start_date) &
        date_released <= as.Date(param$end_date))
      )
    )
}
```

#### Date Range Check

*DB `date_released`* 
```{r echo=FALSE}
range(release_butte_db$date_released, na.rm = TRUE)
```
*EDI `releaseTime`*
```{r}
range(release_butte_edi$releaseTime)
```


#### markedLifeStage

*EDI `markedLifeStage`*

* `r round(sum(is.na(release_butte_edi$markedLifeStage))/nrow(release_butte_edi), 3)*100` % of values in the `markedLifeStage` column are NA.

```{r echo=FALSE}
table(release_butte_edi$markedLifeStage)
```

*DB `lifestage_id`*

* `r round(sum(is.na(release_butte_db$lifestage_id))/nrow(release_butte_db), 3)*100` % of values in the `fork_length` column are NA.

```{r echo=FALSE}
summary(release_butte_db$lifestage_id, na.rm = TRUE)
```

#### nReleased

*EDI `nReleased`*

* `r round(sum(is.na(release_butte_edi$nReleased))/nrow(release_butte_edi), 3)*100` % of values in the `nReleased` column are NA.

```{r echo=FALSE}
summary(release_butte_edi$nReleased)
```

*DB `number_released`*

* `r round(sum(is.na(release_butte_db$number_released))/nrow(release_butte_db), 3)*100` % of values in the `number_released` column are NA.

```{r echo=FALSE}
summary(release_butte_db$number_released, na.rm = TRUE)
```

*Exploration Plots* 

```{r echo=FALSE}
release_butte_edi_clean<- release_butte_edi |>
  mutate(date = as.Date(releaseTime),
         source = "EDI")
release_butte_db_clean <- release_butte_db |>
  mutate(date = as.Date(date_released),
         source = "DB")

ggplot() +
  geom_jitter(data = release_butte_db_clean,
              aes(x = date, y = number_released, color = source),
              color = "blue", alpha = 0.4, width = 10, height = 0.5, shape = 16, size = 2) +
  geom_jitter(data = release_butte_edi_clean,
              aes(x = date, y = nReleased),
              color = "red", alpha = 0.4, width = 10, height = 0.5, shape = 17, size = 2) +
  theme_minimal() +
  labs(x = "Date",
       y = "number_released or nReleased",
       color = "Source"
       )
```

Random pick of data entry by date to compare between db and EDI

  - test 1
```{r echo=FALSE}
# random pick to compare data
release_butte_db |>
  mutate(source = "DB") |>
  filter(release_id == "267") |>
  glimpse()

release_butte_edi |>
  mutate(source = "EDI") |>
  filter(releaseID == "267") |>
  glimpse()
```

  - test 2
```{r echo=FALSE}
# random pick to compare data
release_butte_db |>
  mutate(source = "DB") |>
  filter(release_id == "284") |>
  glimpse()

release_butte_edi |>
  mutate(source = "EDI") |>
  filter(releaseID == "284") |>
  glimpse()
```

  - test 3
```{r echo=FALSE}
# random pick to compare data
release_butte_db |>
  mutate(source = "DB") |>
  filter(release_id == "256") |>
  glimpse()

release_butte_edi |>
  mutate(source = "EDI") |>
  filter(releaseID == "256") |>
  glimpse()
```

### recapture

**- Findings:**

   - folk length values all are NA in both DB and EDI

```{r echo=FALSE}
recapture_with_stream <- recaptured_fish |>
  left_join(trap_id)

recapture_butte_db <- recapture_with_stream |> 
  filter(stream == "butte creek") 
```
```{r echo=FALSE}
recapture_butte_edi <- read_csv("data-raw/edi-zips/butte_recapture.csv") |>
  filter(commonName == "Chinook salmon") |>
  mutate(siteName = tolower(siteName),
         subSiteName = tolower(subSiteName)) |> 
  mutate(
    siteName = case_when(
    siteName %in% c("okie rst", "parrot-phelan", "parrott-phelan canal trap box", "parrot-phelan rst") ~ "okie dam",
    TRUE ~ siteName
    ),
    subSiteName = case_when(
      subSiteName == "pp rst" ~ "okie dam 1",
      subSiteName == "pp rst 2" ~ "okie dam 2",
      subSiteName == "okie rst" ~ NA_character_,
      subSiteName == "canal trap box" ~ "okie dam fyke trap",
      TRUE ~ subSiteName)) |> 
  mutate(stream="butte creek",
         date = as.Date(visitTime))

filter_params_butte <- filter_params$`butte creek`
# 
for (param in filter_params_butte){
  recapture_butte_edi <- subset(recapture_butte_edi,
      !(stream == "butte creek" &
        (
          (is.na(param$site) & is.na(siteName)) |
          (!is.na(param$site) & siteName == param$site)
        ) &
        (
          (is.na(param$subsite) & is.na(subSiteName)) |
          (!is.na(param$subsite) & subSiteName == param$subsite)
        ) &
        (date >= as.Date(param$start_date) &
        date <= as.Date(param$end_date))
      )
    )
}
```

#### Date Range Check

*DB `date`* 
```{r echo=FALSE}
range(recapture_butte_db$date)
```
*EDI `date`*
```{r}
range(recapture_butte_edi$date)
```

#### Count

*DB `count`* 

* `r round(sum(is.na(recapture_butte_db$count))/nrow(recapture_butte_db), 3)*100` % of values in the `count` column are NA.
```{r echo=FALSE}
summary(recapture_butte_db$count, na.rm = TRUE)
```

*EDI `n`*

* `r round(sum(is.na(recapture_butte_edi$n))/nrow(recapture_butte_edi), 3)*100` % of values in the `n` column are NA.
```{r}
summary(recapture_butte_edi$n)
```



#### lifeStage

*EDI `lifeStage`*

* `r round(sum(is.na(recapture_butte_edi$lifeStage))/nrow(recapture_butte_edi), 3)*100` % of values in the `lifeStage` column are NA.

```{r echo=FALSE}
table(recapture_butte_edi$lifeStage)
```

*DB `lifestage_id`*

* `r round(sum(is.na(recapture_butte_db$lifestage_id))/nrow(recapture_butte_db), 3)*100` % of values in the `lifestage_id` column are NA.

```{r echo=FALSE}
summary(recapture_butte_db$lifestage_id, na.rm = TRUE)
```


```{r echo=FALSE, warning=FALSE}
ggplot() +
  geom_point(data = recapture_butte_db,
              aes(x = date, y = count),
              color = "blue", alpha = 0.4, width = 10, height = 0.5, shape = 16, size = 2) +
  geom_point(data = recapture_butte_edi,
              aes(x = as.Date(visitTime), y = n),
              color = "red", alpha = 0.4, width = 10, height = 0.5, shape = 17, size = 2) +
  theme_minimal() +
  labs(x = "Date Recaptured",
       y = "count/number_recaptured")
```

#### forkLength

*EDI `forkLength`*

* `r round(sum(is.na(recapture_butte_edi$forkLength))/nrow(recapture_butte_edi), 3)*100` % of values in the `forkLength` column are NA.

*DB `fork_length`*

* `r round(sum(is.na(recapture_butte_db$fork_length))/nrow(recapture_butte_db), 3)*100` % of values in the `fork_length` column are NA.

Random pick of data entry to compare between db and EDI

  - test 1
```{r echo=FALSE}
# random pick to compare data
recapture_butte_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "48132") |>
  glimpse()

recapture_butte_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "48132") |>
  glimpse()
```

  - test 2
```{r echo=FALSE}
# random pick to compare data
recapture_butte_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "48250") |>
  glimpse()

recapture_butte_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "48250") |>
  glimpse()
```

  - test 3
```{r echo=FALSE}
# random pick to compare data
recapture_butte_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "46604") |>
  glimpse()

recapture_butte_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "46604") |>
  glimpse()
```

  - test 4
```{r echo=FALSE}
# random pick to compare data
recapture_butte_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "45421") |>
  glimpse()

recapture_butte_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "45421") |>
  glimpse()
```

