---
title: "feather-data-checks"
author: "Inigo Peng, edited by Badhia Yunes-Katz"
date: "`r Sys.Date()`"
output: html_document
---


```{r include=FALSE}
library(DBI)
library(tidyverse)
library(EDIutils)
library(EML)
library(readr)
library(janitor)
load("filter_params.RData")
```

```{r include=FALSE}
con <- DBI::dbConnect(
  RPostgres::Postgres(),
  dbname = "jpedb-staging",
  host = "jpe-db.postgres.database.azure.com",
  port = 5432,
  user = Sys.getenv("jpe_db_user_id"),
  password = Sys.getenv("jpe_db_password")
)
```
# Database Tables
```{r echo=FALSE}
catch <- DBI::dbReadTable(con, "catch")
trap_visit <- DBI::dbReadTable(con, "trap_visit")
release <- DBI::dbReadTable(con, "release")
recaptured_fish <- DBI::dbReadTable(con, "recaptured_fish")
trap_location <- DBI::dbReadTable(con, "trap_location")
released_fish <- DBI::dbReadTable(con, "released_fish")
```

```{r include=FALSE}
trap_id <- trap_location |> 
  rename(trap_location_id = id) |>
  select(trap_location_id, stream, site, subsite) 

catch_with_stream <- catch |>
  left_join(trap_id)
```

## Feather {.tabset}
### catch

**- Findings catch:**

  - Date ranges are the same
  - life stage in edi data has not gone through wrangling

```{r echo=FALSE}
catch_feather_edi <- read_csv("data-raw/edi-zips/feather_catch.csv") |>  
  filter(commonName == "Chinook salmon") |>
  mutate(stream="feather river",
         date = as.Date(visitTime)) 
```

#### Date Range Check
*EDI `date`*
```{r}
range(catch_feather_edi$date)
```

*DB `date`*
  
```{r echo=FALSE}
catch_feather_db <- catch_with_stream |> 
  filter(stream == "feather river") 

range(catch_feather_db$date)
```

#### ForkLength
*EDI `forkLength`*

* `r round(sum(is.na(catch_feather_edi$forkLength))/nrow(catch_feather_edi), 3)*100` % of values in the `forkLength` column are NA.

```{r echo=FALSE}
summary(catch_feather_edi$forkLength, na.rm = TRUE)
```

*DB `fork_length`*

* `r round(sum(is.na(catch_feather_db$fork_length))/nrow(catch_feather_db), 3)*100` % of values in the `fork_length` column are NA.

```{r echo=FALSE}
summary(catch_feather_db$fork_length, na.rm = TRUE)
```

```{r include=FALSE}
catch_feather_edi_edited <- catch_feather_edi |>
  clean_names() |>
  # mutate(date = as.Date(visit_time)) |> 
  mutate(source = "EDI")
  # filter(date > ymd("1999-06-04")) 
 
catch_feather_db_edited <- catch_feather_db |>
  mutate(source = "DB")
```

*Forklength Exploration Plots* 
```{r echo=FALSE, warning=FALSE}
combined_catch_feather <- catch_feather_db_edited |> 
  select(date, fork_length, source) |> 
  bind_rows(catch_feather_edi_edited |> select(date, fork_length, source)) 

ggplot(data = combined_catch_feather, aes(x = date, y = fork_length, color = source)) +
  geom_point() +
  facet_wrap(~source)
```

#### Lifestage
(life stage in edi data has not gone through wrangling)

*EDI `lifeStage`*

* `r round(sum(is.na(catch_feather_edi$lifeStage))/nrow(catch_feather_edi), 3)*100` % of values in the `lifeStage` column from the EDI data are NA.

*DB `lifestage_id`*

* `r round(sum(is.na(catch_feather_db$lifestage_id))/nrow(catch_feather_db), 3)*100` % of values in the `lifestage_id` column from the database are NA.

#### Count
*EDI `n`*

* `r round(sum(is.na(catch_feather_edi$n))/nrow(catch_feather_edi), 3)*100` % of values in the `n` column in edi data are NA.

```{r echo=FALSE}
summary(catch_feather_edi$n)
```

*DB `count`*

* `r round(sum(is.na(catch_feather_db$count))/nrow(catch_feather_db), 3)*100` % of values in the `count` column in database data are NA.

```{r echo=FALSE}
summary(catch_feather_db$count)
```

```{r echo=FALSE, warning=FALSE}
ggplot(catch_feather_db_edited, aes(x = date, y = count, color = source)) +
  geom_jitter(alpha = 0.6, width = 0.3, height = 0) +
  theme_minimal() +
  labs(x = "date", y = "count", color = "Source")

ggplot() +
  geom_jitter(data = catch_feather_db_edited,
              aes(x = date, y = count),
              color = "blue", alpha = 0.4, width = 10, height = 0.5, shape = 16, size = 2) +
  geom_jitter(data = catch_feather_edi_edited,
              aes(x = date, y = n),
              color = "red", alpha = 0.4, width = 10, height = 0.5, shape = 17, size = 2) +
  theme_minimal() +
  labs(x = "Date",
       y = "count/n")
```

Random pick of data entry to compare between db and EDI

- test 1

```{r echo=FALSE}
# random pick to compare data
catch_feather_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "55171") |>
  glimpse()

catch_feather_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "55171") |>
  glimpse()
```

- test 2

```{r echo=FALSE}
# random pick to compare data
catch_feather_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id  == "73") |>
  glimpse()

catch_feather_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == 73) |>
  glimpse()

```

- test 3

```{r echo=FALSE}
# random pick to compare data
catch_feather_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id  == "41525") |>
  glimpse()

catch_feather_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "41525") |>
  glimpse()
```

### trap

**- Findings trap:**

  - there is no waterVel in EDI data

```{r echo=FALSE}
trap_with_stream <- trap_visit |>
  left_join(trap_id)

trap_feather_db <- trap_with_stream |> 
  filter(stream == "feather river") 
```

```{r echo=FALSE}
trap_feather_edi <- read_csv("data-raw/edi-zips/feather_trap.csv") |>  
  mutate(stream="feather river",
         date = as.Date(visitTime))
```

#### Date Range Check

*DB `trap_visit_time_start`* 
```{r echo=FALSE}
range(trap_feather_db$trap_visit_time_start, na.rm = TRUE)
```
*EDI `visitTime`*
```{r}
range(trap_feather_edi$visitTime)
```

#### water_velocity

*DB `water_velocity`*

* `r round(sum(is.na(trap_feather_db$water_velocity))/nrow(trap_feather_db), 3)*100` % of values in the `water_velocity` column are NA.
```{r echo=FALSE}
summary(trap_feather_db$water_velocity)
```

*EDI `waterVel`* - no column present
* `r round(sum(is.na(trap_feather_edi$waterVel))/nrow(trap_feather_edi), 3)*100` % of values in the `water_velocity` column are NA.
```{r echo=FALSE}
summary(trap_feather_edi$waterVel)
```

#### fishProcessed

*DB `fish_processed_id`*

* `r round(sum(is.na(trap_feather_db$fish_processed_id))/nrow(trap_feather_db), 3)*100` % of values in the `fish_processed_id` column are NA.
```{r echo=FALSE}
table(trap_feather_db$fish_processed_id)
```

*EDI `fishProcessed`*
* `r round(sum(is.na(trap_feather_edi$fishProcessed))/nrow(trap_feather_edi), 3)*100` % of values in the `fishProcessed` column are NA.
```{r echo=FALSE}
table(trap_feather_edi$fishProcessed)
```

Random pick of data entry by date to compare between db and EDI

  - test 1
```{r echo=FALSE}
trap_feather_db |>
  mutate(source = "DB") |>
  filter(trap_visit_id  == "23") |>
  glimpse()

trap_feather_edi |>
  mutate(source = "EDI") |>
  filter(trapVisitID == "23") |>
  glimpse()
```

  - test 2
```{r echo=FALSE}
# random pick to compare data
trap_feather_db |>
  mutate(source = "DB") |>
  filter(trap_visit_id  == "2144") |>
  glimpse()

trap_feather_edi |>
  mutate(source = "EDI") |>
  filter(trapVisitID == "2144") |>
  glimpse()
```

  - test 3
```{r echo=FALSE}
# random pick to compare data
trap_feather_db |>
  mutate(source = "DB") |>
  filter(trap_visit_id  == "2327") |>
  glimpse()

trap_feather_edi |>
  mutate(source = "EDI") |>
  filter(trapVisitID == "2327") |>
  glimpse()
```

### recapture

**- Findings recapture:**

  - Date ranges are the same

```{r echo=FALSE}
recapture_with_stream <- recaptured_fish |>
  left_join(trap_id)

recapture_feather_db <- recapture_with_stream |> 
  filter(stream == "feather river") 
```

```{r echo=FALSE}
recapture_feather_edi <- read_csv("data-raw/edi-zips/feather_recapture.csv") |>  
  filter(commonName == "Chinook salmon") |>
  mutate(stream="feather river",
         date = as.Date(visitTime))
```

#### Date Range Check

*DB `date`* 
```{r echo=FALSE}
range(recapture_feather_db$date, na.rm = TRUE)
```
*EDI `date`*
```{r}
range(recapture_feather_edi$date)
```

#### Count

*DB `count`* 

* `r round(sum(is.na(recapture_feather_db$count))/nrow(recapture_feather_db), 3)*100` % of values in the `count` column are NA.
```{r echo=FALSE}
summary(recapture_feather_db$count, na.rm = TRUE)
```

*EDI `n`*

* `r round(sum(is.na(recapture_feather_edi$n))/nrow(recapture_feather_edi), 3)*100` % of values in the `n` column are NA.
```{r}
summary(recapture_feather_edi$n)
```

#### lifeStage

*EDI `lifeStage`*

* `r round(sum(is.na(recapture_feather_edi$lifeStage))/nrow(recapture_feather_edi), 3)*100` % of values in the `lifeStage` column are NA.

```{r echo=FALSE}
table(recapture_feather_edi$lifeStage)
```

*DB `lifestage_id`*

* `r round(sum(is.na(recapture_feather_db$lifestage_id))/nrow(recapture_feather_db), 3)*100` % of values in the `lifestage_id` column are NA.

```{r echo=FALSE}
summary(recapture_feather_db$lifestage_id)
table(recapture_feather_db$lifestage_id)
```


```{r echo=FALSE, warning=FALSE}
ggplot() +
  geom_point(data = recapture_feather_db,
              aes(x = date, y = count),
              color = "blue", alpha = 0.4, width = 10, height = 0.5, shape = 16, size = 2) +
  geom_point(data = recapture_feather_edi,
              aes(x = as.Date(visitTime), y = n),
              color = "red", alpha = 0.4, width = 10, height = 0.5, shape = 17, size = 2) +
  theme_minimal() +
  labs(x = "Date Recaptured",
       y = "count/number_recaptured")
```

#### forkLength

*EDI `forkLength`*

* `r round(sum(is.na(recapture_feather_edi$forkLength))/nrow(recapture_feather_edi), 3)*100` % of values in the `forkLength` column are NA.

*DB `fork_length`*

* `r round(sum(is.na(recapture_feather_db$fork_length))/nrow(recapture_feather_db), 3)*100` % of values in the `fork_length` column are NA.

#### Run

*EDI `run`*

* `r round(sum(is.na(recapture_feather_edi$run))/nrow(recapture_feather_edi), 3)*100` % of values in the `run` column are NA.

```{r echo=FALSE}
summary(recapture_feather_edi$run)
table(recapture_feather_edi$run)
```

*DB `run_id`*

* `r round(sum(is.na(recapture_feather_db$run_id))/nrow(recapture_feather_db), 3)*100` % of values in the `run_id` column are NA.
```{r echo=FALSE}
summary(recapture_feather_db$run_id)
table(recapture_feather_db$run_id)
```

Random pick of data entry to compare between db and EDI

  - test 1
```{r echo=FALSE}
# random pick to compare data
recapture_feather_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "215015") |>
  glimpse()

recapture_feather_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "215015") |>
  glimpse()
```

  - test 2
```{r echo=FALSE}
# random pick to compare data
recapture_feather_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "234245") |>
  glimpse()

recapture_feather_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "234245") |>
  glimpse()
```

  - test 3
```{r echo=FALSE}
# random pick to compare data
recapture_feather_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "76916066") |>
  glimpse()

recapture_feather_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "76916066") |>
  glimpse()
```

### release

**- Findings release:**

  - No markedLifeStage in EDI, all lifestage_id are NA in DB
  - There is no release_id == 140 or == 231 in EDI (look at test 1 and 2)

```{r include=FALSE}
release_with_stream <- release |>
  left_join(trap_id)

release_feather_db <- release_with_stream |> 
  filter(stream == "feather river") 
```

```{r include=FALSE}
release_feather_edi <- read_csv("data-raw/edi-zips/feather_release.csv") |>
  filter(commonName == "Chinook salmon") |>
  mutate(stream="feather river",
         date = as.Date(releaseTime))
```


#### Date Range Check

*DB `date_released`* 
```{r echo=FALSE}
range(release_feather_db$date_released, na.rm = TRUE)
```
*EDI `releaseTime`*
```{r}
range(release_feather_edi$releaseTime)
```


#### markedLifeStage

*EDI `markedLifeStage`*

* `r round(sum(is.na(release_feather_edi$markedLifeStage))/nrow(release_feather_edi), 3)*100` % of values in the `markedLifeStage` column are NA.

*DB `lifestage_id`*

* `r round(sum(is.na(release_feather_db$lifestage_id))/nrow(release_feather_db), 3)*100` % of values in the `fork_length` column are NA.

```{r echo=FALSE}
summary(release_feather_db$lifestage_id, na.rm = TRUE)
```

#### nReleased

*EDI `nReleased`*

* `r round(sum(is.na(release_feather_edi$nReleased))/nrow(release_feather_edi), 3)*100` % of values in the `nReleased` column are NA.

```{r echo=FALSE}
summary(release_feather_edi$nReleased)
```

*DB `number_released`*

* `r round(sum(is.na(release_feather_db$number_released))/nrow(release_feather_db), 3)*100` % of values in the `number_released` column are NA.

```{r echo=FALSE}
summary(release_feather_db$number_released, na.rm = TRUE)
```

*Exploration Plots* 

```{r echo=FALSE}
release_feather_edi_clean <- release_feather_edi |>
  mutate(date = as.Date(releaseTime),
         source = "EDI")
release_feather_db_clean <- release_feather_db |>
  mutate(date = as.Date(date_released),
         source = "DB")

ggplot() +
  geom_jitter(data = release_feather_db_clean,
              aes(x = date, y = number_released, color = source),
              color = "blue", alpha = 0.4, width = 10, height = 0.5, shape = 16, size = 2) +
  geom_jitter(data = release_feather_edi_clean,
              aes(x = date, y = nReleased),
              color = "red", alpha = 0.4, width = 10, height = 0.5, shape = 17, size = 2) +
  theme_minimal() +
  labs(x = "Date",
       y = "number_released or nReleased",
       color = "Source"
       )
```

Random pick of data entry by date to compare between db and EDI

  - test 1
```{r echo=FALSE}
# random pick to compare data
release_feather_db |>
  mutate(source = "DB") |>
  filter(release_id == "140") |>
  glimpse()

release_feather_edi |>
  mutate(source = "EDI") |>
  filter(releaseID == "140") |>
  glimpse()
```

  - test 2
```{r echo=FALSE}
# random pick to compare data
release_feather_db |>
  mutate(source = "DB") |>
  filter(release_id == "231") |>
  glimpse()

release_feather_edi |>
  mutate(source = "EDI") |>
  filter(releaseID == "231") |>
  glimpse()
```

  - test 3
```{r echo=FALSE}
# random pick to compare data
release_feather_db |>
  mutate(source = "DB") |>
  filter(release_id == "547") |>
  glimpse()

release_feather_edi |>
  mutate(source = "EDI") |>
  filter(releaseID == "547") |>
  glimpse()
```

  - test 4
```{r echo=FALSE}
# random pick to compare data
release_feather_db |>
  mutate(source = "DB") |>
  filter(release_id == "901") |>
  glimpse()

release_feather_edi |>
  mutate(source = "EDI") |>
  filter(releaseID == "901") |>
  glimpse()
```

### releasefish

**- Findings releasefish:**

```{r echo=FALSE}
release_fish_with_stream <- released_fish |>
  left_join(trap_id)

release_fish_feather_db <- release_fish_with_stream |> 
  filter(stream == "feather river") 
```

```{r echo=FALSE}
release_fish_feather_edi <- read_csv("data-raw/edi-zips/feather_releasefish.csv") |>
  # filter(commonName == "Chinook salmon") |>
  mutate(stream="feather river")
```

#### column names

*EDI*
```{r echo=FALSE}
colnames(release_fish_feather_edi)
```

*DB*
```{r echo=FALSE}
colnames(release_fish_feather_db)
```

#### forkLength

*EDI `forkLength`*

* `r round(sum(is.na(release_fish_feather_edi$forkLength))/nrow(release_fish_feather_edi), 3)*100` % of values in the `forkLength` column are NA.

```{r echo=FALSE}
summary(release_fish_feather_edi$forkLength, na.rm = TRUE)
```

*DB `fork_length`*

* `r round(sum(is.na(release_fish_feather_db$fork_length))/nrow(release_fish_feather_db), 3)*100` % of values in the `fork_length` column are NA.

```{r echo=FALSE}
summary(release_fish_feather_db$fork_length, na.rm = TRUE)
```