---
title: "feather-data-checks"
author: "Inigo Peng"
date: "`r Sys.Date()`"
output: html_document
---


```{r include=FALSE}
library(DBI)
library(tidyverse)
library(EDIutils)
library(EML)
library(readr)
library(janitor)
load("filter_params.RData")
```

```{r include=FALSE}
con <- DBI::dbConnect(
  RPostgres::Postgres(),
  dbname = "jpedb-staging",
  host = "jpe-db.postgres.database.azure.com",
  port = 5432,
  user = Sys.getenv("jpe_db_user_id"),
  password = Sys.getenv("jpe_db_password")
)
```
# Database Tables
```{r echo=FALSE}
catch <- DBI::dbReadTable(con, "catch")
trap_visit <- DBI::dbReadTable(con, "trap_visit")
release <- DBI::dbReadTable(con, "release")
recaptured_fish <- DBI::dbReadTable(con, "recaptured_fish")
trap_location <- DBI::dbReadTable(con, "trap_location")
released_fish <- DBI::dbReadTable(con, "released_fish")
```

```{r include=FALSE}
trap_id <- trap_location |> 
  rename(trap_location_id = id) |>
  select(trap_location_id, stream, site, subsite) 

catch_with_stream <- catch |>
  left_join(trap_id)
```

# Feather

## catch

**- Findings catch:**

```{r echo=FALSE}
catch_feather_edi <- read_csv("data-raw/edi-zips/feather_catch.csv") |>  
  filter(commonName == "Chinook salmon") |>
  mutate(stream="feather river",
         date = as.Date(visitTime))
```

### Date Range Check
*EDI `date`*
```{r}
range(catch_feather_edi$date)
```

*DB `date`*
  
```{r echo=FALSE}
catch_feather_db <- catch_with_stream |> 
  filter(stream == "feather river") 

range(catch_feather_db$date)
```

### ForkLength
*EDI `forkLength`*

* `r round(sum(is.na(catch_feather_edi$forkLength))/nrow(catch_feather_edi), 3)*100` % of values in the `forkLength` column are NA.

```{r echo=FALSE}
summary(catch_feather_edi$forkLength, na.rm = TRUE)
```

*DB `fork_length`*

* `r round(sum(is.na(catch_feather_db$fork_length))/nrow(catch_feather_db), 3)*100` % of values in the `fork_length` column are NA.

```{r echo=FALSE}
summary(catch_feather_db$fork_length, na.rm = TRUE)
```

```{r include=FALSE}
catch_feather_edi_edited <- catch_feather_edi |>
  clean_names() |>
  # mutate(date = as.Date(visit_time)) |> 
  mutate(source = "EDI")
  # filter(date > ymd("1999-06-04")) 
 
catch_feather_db_edited <- catch_feather_db |>
  mutate(source = "DB")
```
*Forklength Exploration Plots* 

```{r echo=FALSE, warning=FALSE}
combined_catch_feather <- catch_feather_db_edited |> 
  select(date, fork_length, source) |> 
  bind_rows(catch_feather_edi_edited |> select(date, fork_length, source)) 

ggplot(data = combined_catch_feather, aes(x = date, y = fork_length, color = source)) +
  geom_point() +
  facet_wrap(~source)
```

### Lifestage
(life stage in edi data has not gone through wrangling)

*EDI `lifeStage`*

* `r round(sum(is.na(catch_feather_edi$lifeStage))/nrow(catch_feather_edi), 3)*100` % of values in the `lifeStage` column from the EDI data are NA.

*DB `lifestage_id`*

* `r round(sum(is.na(catch_feather_db$lifestage_id))/nrow(catch_feather_db), 3)*100` % of values in the `lifestage_id` column from the database are NA.

```{r echo=FALSE, warning=FALSE}
catch_feather_db |>
   ggplot(aes(x = as.factor(lifestage_id), y = fork_length, fill = as.factor(lifestage_id))) +
  geom_boxplot(outlier.shape = 16, outlier.size = 1, alpha = 0.7) +
  coord_flip() +
  theme_minimal() +
  labs(x = "lifestage id", y = "fork length", title = "DB data - lifestage vs forklength")
```


```{r echo=FALSE, warning=FALSE}
catch_feather_edi |>
   ggplot(aes(x = as.factor(lifeStage), y = forkLength, fill = as.factor(lifeStage))) +
  geom_boxplot(outlier.shape = 16, outlier.size = 1, alpha = 0.7) +
  coord_flip() +
  theme_minimal() +
  labs(x = "lifestage id", y = "fork length", title = "EDI data - lifestage vs forklength")
```

### Count
TODO: EDI data contains lifestage value "mixed" but not in lifestage lookuptable

*EDI `n`*

* `r round(sum(is.na(catch_feather_edi$n))/nrow(catch_feather_edi), 3)*100` % of values in the `n` column in edi data are NA.

```{r echo=FALSE}
summary(catch_feather_edi$n)
```

*DB `count`*

* `r round(sum(is.na(catch_feather_db$count))/nrow(catch_feather_db), 3)*100` % of values in the `count` column in database data are NA.

```{r echo=FALSE}
summary(catch_feather_db$count)
```

```{r echo=FALSE, warning=FALSE}
ggplot(catch_feather_db_edited, aes(x = date, y = count, color = source)) +
  geom_jitter(alpha = 0.6, width = 0.3, height = 0) +
  theme_minimal() +
  labs(x = "date", y = "count", color = "Source")

ggplot() +
  geom_jitter(data = catch_feather_db_edited,
              aes(x = date, y = count),
              color = "blue", alpha = 0.4, width = 10, height = 0.5, shape = 16, size = 2) +
  geom_jitter(data = catch_feather_edi_edited,
              aes(x = date, y = n),
              color = "red", alpha = 0.4, width = 10, height = 0.5, shape = 17, size = 2) +
  theme_minimal() +
  labs(x = "Date",
       y = "count/n")
```

Random pick of data entry to compare between db and EDI

- test 1

```{r echo=FALSE}
# random pick to compare data
catch_feather_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "55171") |>
  glimpse()

catch_feather_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "55171") |>
  glimpse()
```

- test 2

```{r echo=FALSE}
# random pick to compare data
catch_feather_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id  == "73") |>
  glimpse()

catch_feather_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == 73) |>
  glimpse()

```

- test 3

```{r echo=FALSE}
# random pick to compare data
catch_feather_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id  == "41525") |>
  glimpse()

catch_feather_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "41525") |>
  glimpse()
```

## TODO 
## trap

**- Findings trap:**

```{r echo=FALSE}
trap_with_stream <- trap_visit |>
  left_join(trap_id)

trap_feather_db <- trap_with_stream |> 
  filter(stream == "feather river") 
``

```{r echo=FALSE}
trap_feather_edi <- read_csv("data-raw/edi-zips/feather_trap.csv") |>  
  mutate(stream="feather river",
         date = as.Date(visitTime))
```

## TODO 
## recapture

**- Findings recapture:**

```{r echo=FALSE}
recapture_with_stream <- recaptured_fish |>
  left_join(trap_id)

recapture_feather_db <- recapture_with_stream |> 
  filter(stream == "feather river") 
```

```{r echo=FALSE}
recapture_feather_edi <- read_csv("data-raw/edi-zips/feather_recapture.csv") |>  
  filter(commonName == "Chinook salmon") |>
  mutate(stream="feather river",
         date = as.Date(visitTime))
```
## TODO
## release

**- Findings release:**

```{r echo=FALSE}
release_with_stream <- release |>
  left_join(trap_id)

release_feather_db <- release_with_stream |> 
  filter(stream == "feather river") 
```

```{r echo=FALSE}
release_feather_edi <- read_csv("data-raw/edi-zips/feather_release.csv") |>
  filter(commonName == "Chinook salmon") |>
  mutate(stream="feather river",
         date = as.Date(releaseTime))
```
## TODO
## releasefish

**- Findings releasefish:**

```{r echo=FALSE}
release_fish_with_stream <- released_fish |>
  left_join(trap_id)

release_fish_feather_db <- release_fish_with_stream |> 
  filter(stream == "feather river") 
```

```{r echo=FALSE}
release_fish_feather_edi <- read_csv("data-raw/edi-zips/feather_releasefish.csv") |>
  # filter(commonName == "Chinook salmon") |>
  mutate(stream="feather river")
```
