---
title: "tisdale-data-checks"
author: "Inigo Peng, edited by Badhia Yunes-Katz"
date: "`r Sys.Date()`"
output: html_document
---
```{r include=FALSE}
library(DBI)
library(tidyverse)
library(EDIutils)
library(EML)
library(readr)
library(janitor)
load("filter_params.RData")
```

```{r include=FALSE}
con <- DBI::dbConnect(
  RPostgres::Postgres(),
  dbname = "jpedb-staging",
  host = "jpe-db.postgres.database.azure.com",
  port = 5432,
  user = Sys.getenv("jpe_db_user_id"),
  password = Sys.getenv("jpe_db_password")
)
```
# Database Tables
```{r echo=FALSE}
catch <- DBI::dbReadTable(con, "catch")
trap_visit <- DBI::dbReadTable(con, "trap_visit")
release <- DBI::dbReadTable(con, "release")
recaptured_fish <- DBI::dbReadTable(con, "recaptured_fish")
trap_location <- DBI::dbReadTable(con, "trap_location")
released_fish <- DBI::dbReadTable(con, "released_fish")
```

```{r include=FALSE}
trap_id <- trap_location |> 
  rename(trap_location_id = id) |>
  select(trap_location_id, stream, site, subsite) 

catch_with_stream <- catch |>
  left_join(trap_id)
```


## tisdale {.tabset}

### TODOS
  - map tisdale trap water_vel to water_velocity (done)
  - release_subsite == "1 rm up" this is not in lookup table
  
### catch

**- Findings catch:**

  - date range is the same 
  - potential outlier value on folk_length
  - life stage in edi data has not gone through wrangling

```{r echo=FALSE}
catch_tisdale_edi <- read_csv("data-raw/edi-zips/tisdale_catch.csv") |>
  filter(commonName == "Chinook salmon") |>
  mutate(siteName = tolower(siteName),
         subSiteName = tolower(subSiteName)) |> 
  mutate(
    siteName = case_when(
      siteName == "tisdale weir rst" ~ "tisdale",
      TRUE ~ siteName)) |> 
  mutate(stream="sacramento river",
         date = as.Date(visitTime))
```

#### Date Range Check
*EDI `date`*
```{r}
range(catch_tisdale_edi$date)
```

*DB `date`*
  
```{r echo=FALSE}
catch_tisdale_db <- catch_with_stream |> 
  filter(site == "tisdale") 

range(catch_tisdale_db$date)
```

#### ForkLength

*EDI `forkLength`*

* `r round(sum(is.na(catch_tisdale_edi$forkLength))/nrow(catch_tisdale_edi), 3)*100` % of values in the `forkLength` column are NA.

```{r echo=FALSE}
summary(catch_tisdale_edi$forkLength, na.rm = TRUE)
```

*DB `fork_length`*

* `r round(sum(is.na(catch_tisdale_db$fork_length))/nrow(catch_tisdale_db), 3)*100` % of values in the `fork_length` column are NA.

```{r echo=FALSE}
summary(catch_tisdale_db$fork_length, na.rm = TRUE)
```

```{r include=FALSE}
catch_tisdale_edi_edited <- catch_tisdale_edi |>
  clean_names() |>
  mutate(source = "EDI")
 
catch_tisdale_db_edited <- catch_tisdale_db |>
  mutate(source = "DB")
```

*Forklength Exploration Plots* 

```{r echo=FALSE, warning=FALSE}
combined_catch_tisdale <- catch_tisdale_db_edited |> 
  select(date, fork_length, source) |> 
  bind_rows(catch_tisdale_edi_edited |> select(date, fork_length, source)) 

ggplot(data = combined_catch_tisdale, aes(x = date, y = fork_length, color = source)) +
  geom_point() +
  facet_wrap(~source)
```

#### Lifestage

*EDI `lifeStage`*

* `r round(sum(is.na(catch_tisdale_edi$lifeStage))/nrow(catch_tisdale_edi), 3)*100` % of values in the `lifeStage` column from the EDI data are NA.

*DB `lifestage_id`*

* `r round(sum(is.na(catch_tisdale_db$lifestage_id))/nrow(catch_tisdale_db), 3)*100` % of values in the `lifestage_id` column from the database are NA.

```{r echo=FALSE, warning=FALSE}
catch_tisdale_db |>
   ggplot(aes(x = as.factor(lifestage_id), y = fork_length, fill = as.factor(lifestage_id))) +
  geom_boxplot(outlier.shape = 16, outlier.size = 1, alpha = 0.7) +
  coord_flip() +
  theme_minimal() +
  labs(x = "lifestage id", y = "fork length", title = "DB data - lifestage vs forklength")
```


```{r echo=FALSE, warning=FALSE}
catch_tisdale_edi |>
   ggplot(aes(x = as.factor(lifeStage), y = forkLength, fill = as.factor(lifeStage))) +
  geom_boxplot(outlier.shape = 16, outlier.size = 1, alpha = 0.7) +
  coord_flip() +
  theme_minimal() +
  labs(x = "lifestage id", y = "fork length", title = "EDI data - lifestage vs forklength")
```

#### Count

*EDI `n`*

* `r round(sum(is.na(catch_tisdale_edi$n))/nrow(catch_tisdale_edi), 3)*100` % of values in the `n` column in edi data are NA.

```{r echo=FALSE}
summary(catch_tisdale_edi$n)
```

*DB `count`*

* `r round(sum(is.na(catch_tisdale_db$count))/nrow(catch_tisdale_db), 3)*100` % of values in the `count` column in database data are NA.

```{r echo=FALSE}
summary(catch_tisdale_db$count)
```

```{r echo=FALSE, warning=FALSE}
ggplot(catch_tisdale_db_edited, aes(x = date, y = count, color = source)) +
  geom_jitter(alpha = 0.6, width = 0.3, height = 0) +
  theme_minimal() +
  labs(x = "date", y = "count", color = "Source")

ggplot() +
  geom_jitter(data = catch_tisdale_db_edited,
              aes(x = date, y = count),
              color = "blue", alpha = 0.4, width = 10, height = 0.5, shape = 16, size = 2) +
  geom_jitter(data = catch_tisdale_edi_edited,
              aes(x = date, y = n),
              color = "red", alpha = 0.4, width = 10, height = 0.5, shape = 17, size = 2) +
  theme_minimal() +
  labs(x = "Date",
       y = "count/n")
```

Random pick of data entry to compare between db and EDI

- test 1

```{r echo=FALSE}
# random pick to compare data
catch_tisdale_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "400") |>
  glimpse()

catch_tisdale_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "400") |>
  glimpse()
```

- test 2
```{r echo=FALSE}
# random pick to compare data
catch_tisdale_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id  == "18663") |>
  glimpse()

catch_tisdale_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == 18663) |>
  glimpse()

```

- test 3

```{r echo=FALSE}
# random pick to compare data
catch_tisdale_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id  == "1600") |>
  glimpse()

catch_tisdale_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == 1600) |>
  glimpse()
```

### trap

**- Findings:**

  - All water velocity values are NA in DB

```{r include=FALSE}
trap_with_stream <- trap_visit |>
  left_join(trap_id)

trap_tisdale_db <- trap_with_stream |> 
  filter(site == "tisdale") 
```
```{r include=FALSE}
trap_tisdale_edi <- read_csv("data-raw/edi-zips/tisdale_trap.csv") |>
  mutate(siteName = tolower(siteName),
         subSiteName = tolower(subSiteName)) |> 
  mutate(
    siteName = case_when(
      siteName == "tisdale weir rst" ~ "tisdale",
      TRUE ~ siteName)) |> 
  mutate(stream="sacramento river",
         date = as.Date(visitTime))

```

#### Date Range Check

*DB `trap_visit_time_start`* 
```{r echo=FALSE}
range(trap_tisdale_db$trap_visit_time_start, na.rm = TRUE)
```
*EDI `visitTime`*
```{r}
range(trap_tisdale_edi$visitTime)
```

#### water_velocity

*DB `water_velocity`*

* `r round(sum(is.na(trap_tisdale_db$water_velocity))/nrow(trap_tisdale_db), 3)*100` % of values in the `water_velocity` column are NA.
```{r echo=FALSE}
summary(trap_tisdale_db$water_velocity)
```

*EDI `waterVel`*
* `r round(sum(is.na(trap_tisdale_edi$waterVel))/nrow(trap_tisdale_edi), 3)*100` % of values in the `water_velocity` column are NA.
```{r echo=FALSE}
summary(trap_tisdale_edi$waterVel)
```

```{r echo=FALSE, warning=FALSE}
trap_tisdale_edi_clean <- trap_tisdale_edi |>
  mutate(date = as.Date(visitTime),
         source = "EDI")
trap_tisdale_db_clean <- trap_tisdale_db |>
  mutate(date = as.Date(trap_visit_time_start),
         source = "DB")

ggplot() +
  geom_jitter(data = trap_tisdale_db_clean,
              aes(x = date, y = water_velocity, color = source),
              color = "blue", alpha = 0.4, width = 10, height = 0.5, shape = 16, size = 2) +
  geom_jitter(data = trap_tisdale_edi_clean,
              aes(x = date, y = waterVel),
              color = "red", alpha = 0.4, width = 10, height = 0.5, shape = 17, size = 2) +
  theme_minimal() +
  labs(x = "Date",
       y = "water_velocity or waterVel",
       color = "Source"
       )
```

#### fishProcessed

*DB `fish_processed_id`*

* `r round(sum(is.na(trap_tisdale_db$fish_processed_id))/nrow(trap_tisdale_db), 3)*100` % of values in the `fish_processed_id` column are NA.
```{r echo=FALSE}
table(trap_tisdale_db$fish_processed_id)
```

*EDI `fishProcessed`*
* `r round(sum(is.na(trap_tisdale_edi$fishProcessed))/nrow(trap_tisdale_edi), 3)*100` % of values in the `fishProcessed` column are NA.
```{r echo=FALSE}
table(trap_tisdale_edi$fishProcessed)
```
Random pick of data entry by date to compare between db and EDI

  - test 1
```{r echo=FALSE}
# random pick to compare data
trap_tisdale_db |>
  mutate(source = "DB",
         trap_visit_time_start = as.Date(trap_visit_time_start)) |>
  # filter(id == "2217421") |>
  filter(trap_visit_id == "380") |>
  glimpse()

trap_tisdale_edi |>
  mutate(source = "EDI",
         visitTime = as.Date(visitTime)) |>
  filter(trapVisitID == "380") |>
  glimpse()
```

  - test 2
```{r echo=FALSE}
# random pick to compare data
trap_tisdale_db |>
  mutate(source = "DB",
         trap_visit_time_start = as.Date(trap_visit_time_start)) |>
  # filter(id == "2217421") |>
  filter(trap_visit_id == "2037") |>
  glimpse()

trap_tisdale_edi |>
  mutate(source = "EDI",
         visitTime = as.Date(visitTime)) |>
  filter(trapVisitID == "2037") |>
  glimpse()
```

  - test 3
```{r echo=FALSE}
# random pick to compare data
trap_tisdale_db |>
  mutate(source = "DB",
         trap_visit_time_start = as.Date(trap_visit_time_start)) |>
  # filter(id == "2217421") |>
  filter(trap_visit_id == "2015") |>
  glimpse()

trap_tisdale_edi |>
  mutate(source = "EDI",
         visitTime = as.Date(visitTime)) |>
  filter(trapVisitID == "2015") |>
  glimpse()
```

  - test 4
```{r echo=FALSE}
# random pick to compare data
trap_tisdale_db |>
  mutate(source = "DB",
         trap_visit_time_start = as.Date(trap_visit_time_start)) |>
  # filter(id == "2217421") |>
  filter(trap_visit_id == "6919") |>
  glimpse()

trap_tisdale_edi |>
  mutate(source = "EDI",
         visitTime = as.Date(visitTime)) |>
  filter(trapVisitID == "6919") |>
  glimpse()
```

### release

**- Findings:**

  - date range is the same  
  - most markedLifeStage values in EDI are NA, but no values are NA in DB

```{r include=FALSE}
release_with_stream <- release |>
  left_join(trap_id)

release_tisdale_db <- release_with_stream |> 
  filter(site == "tisdale") 
```
```{r include=FALSE}
release_tisdale_edi <- read_csv("data-raw/edi-zips/tisdale_release.csv") |> 
  filter(commonName == "Chinook salmon") |>
  mutate(releaseSite = tolower(releaseSite),
         releaseSubSite = tolower(releaseSubSite)) |> 
  mutate(
    releaseSite = case_when(
      releaseSite == "tisdale weir release site" ~ "tisdale",
      TRUE ~ releaseSite)) |> 
  mutate(stream="sacramento river",
         date = as.Date(releaseTime))
```

#### Date Range Check

*DB `date_released`* 
```{r echo=FALSE}
range(release_tisdale_db$date_released, na.rm = TRUE)
```
*EDI `releaseTime`*
```{r}
range(release_tisdale_edi$releaseTime)
```


#### markedLifeStage

*EDI `markedLifeStage`*

* `r round(sum(is.na(release_tisdale_edi$markedLifeStage))/nrow(release_tisdale_edi), 3)*100` % of values in the `markedLifeStage` column are NA.

```{r echo=FALSE}
table(release_tisdale_edi$markedLifeStage)
```

*DB `lifestage_id`*

* `r round(sum(is.na(release_tisdale_db$lifestage_id))/nrow(release_tisdale_db), 3)*100` % of values in the `fork_length` column are NA.

```{r echo=FALSE}
summary(release_tisdale_db$lifestage_id, na.rm = TRUE)
```

#### nReleased

*EDI `nReleased`*

* `r round(sum(is.na(release_tisdale_edi$nReleased))/nrow(release_tisdale_edi), 3)*100` % of values in the `nReleased` column are NA.

```{r echo=FALSE}
summary(release_tisdale_edi$nReleased)
```

*DB `number_released`*

* `r round(sum(is.na(release_tisdale_db$number_released))/nrow(release_tisdale_db), 3)*100` % of values in the `number_released` column are NA.

```{r echo=FALSE}
summary(release_tisdale_db$number_released, na.rm = TRUE)
```

*Exploration Plots* 

```{r echo=FALSE}
release_tisdale_edi_clean<- release_tisdale_edi |>
  mutate(date = as.Date(releaseTime),
         source = "EDI")
release_tisdale_db_clean <- release_tisdale_db |>
  mutate(date = as.Date(date_released),
         source = "DB")

ggplot() +
  geom_jitter(data = release_tisdale_db_clean,
              aes(x = date, y = number_released, color = source),
              color = "blue", alpha = 0.4, width = 10, height = 0.5, shape = 16, size = 2) +
  geom_jitter(data = release_tisdale_edi_clean,
              aes(x = date, y = nReleased),
              color = "red", alpha = 0.4, width = 10, height = 0.5, shape = 17, size = 2) +
  theme_minimal() +
  labs(x = "Date",
       y = "number_released or nReleased",
       color = "Source"
       )
```

Random pick of data entry by date to compare between db and EDI

  - test 1
```{r echo=FALSE}
# random pick to compare data
release_tisdale_db |>
  mutate(source = "DB") |>
  filter(release_id == "297") |>
  glimpse()

release_tisdale_edi |>
  mutate(source = "EDI") |>
  filter(releaseID == "297") |>
  glimpse()
```

  - test 2
```{r echo=FALSE}
# random pick to compare data
release_tisdale_db |>
  mutate(source = "DB") |>
  filter(release_id == "297") |>
  glimpse()

release_tisdale_edi |>
  mutate(source = "EDI") |>
  filter(releaseID == "297") |>
  glimpse()
```

  - test 3
```{r echo=FALSE}
# random pick to compare data
release_tisdale_db |>
  mutate(source = "DB") |>
  filter(release_id == "322") |>
  glimpse()

release_tisdale_edi |>
  mutate(source = "EDI") |>
  filter(releaseID == "322") |>
  glimpse()
```

  - test 4
```{r echo=FALSE}
# random pick to compare data
release_tisdale_db |>
  mutate(source = "DB") |>
  filter(release_id == "295") |>
  glimpse()

release_tisdale_edi |>
  mutate(source = "EDI") |>
  filter(releaseID == "295") |>
  glimpse()
```

### recapture

**- Findings:**

- date range is the same  

```{r include=FALSE}
recapture_with_stream <- recaptured_fish |>
  left_join(trap_id)

recapture_tisdale_db <- recapture_with_stream |> 
  filter(site == "tisdale") 
```
```{r include=FALSE}
recapture_tisdale_edi <- read_csv("data-raw/edi-zips/tisdale_recapture.csv") |>
  filter(commonName == "Chinook salmon") |>
  mutate(siteName = tolower(siteName),
         subSiteName = tolower(subSiteName)) |> 
  mutate(
    siteName = case_when(
      siteName == "tisdale weir rst" ~ "tisdale",
      TRUE ~ siteName)) |> 
  mutate(stream="sacramento river",
         date = as.Date(visitTime))
```

#### Date Range Check

*DB `date`* 
```{r echo=FALSE}
range(recapture_tisdale_db$date)
```
*EDI `date`*
```{r}
range(recapture_tisdale_edi$date)
```

#### Count

*DB `count`* 

* `r round(sum(is.na(recapture_tisdale_db$count))/nrow(recapture_tisdale_db), 3)*100` % of values in the `count` column are NA.
```{r echo=FALSE}
summary(recapture_tisdale_db$count, na.rm = TRUE)
```

*EDI `n`*

* `r round(sum(is.na(recapture_tisdale_edi$n))/nrow(recapture_tisdale_edi), 3)*100` % of values in the `n` column are NA.
```{r}
summary(recapture_tisdale_edi$n)
```

#### lifeStage

*EDI `lifeStage`*

* `r round(sum(is.na(recapture_tisdale_edi$lifeStage))/nrow(recapture_tisdale_edi), 3)*100` % of values in the `lifeStage` column are NA.

```{r echo=FALSE}
table(recapture_tisdale_edi$lifeStage)
```

*DB `lifestage_id`*

* `r round(sum(is.na(recapture_tisdale_db$lifestage_id))/nrow(recapture_tisdale_db), 3)*100` % of values in the `lifestage_id` column are NA.

```{r echo=FALSE}
summary(recapture_tisdale_db$lifestage_id, na.rm = TRUE)
```


```{r echo=FALSE, warning=FALSE}
ggplot() +
  geom_point(data = recapture_tisdale_db,
              aes(x = date, y = count),
              color = "blue", alpha = 0.4, width = 10, height = 0.5, shape = 16, size = 2) +
  geom_point(data = recapture_tisdale_edi,
              aes(x = as.Date(visitTime), y = n),
              color = "red", alpha = 0.4, width = 10, height = 0.5, shape = 17, size = 2) +
  theme_minimal() +
  labs(x = "Date Recaptured",
       y = "count/number_recaptured")
```

#### forkLength

*EDI `forkLength`*

* `r round(sum(is.na(recapture_tisdale_edi$forkLength))/nrow(recapture_tisdale_edi), 3)*100` % of values in the `forkLength` column are NA.

*DB `fork_length`*

* `r round(sum(is.na(recapture_tisdale_db$fork_length))/nrow(recapture_tisdale_db), 3)*100` % of values in the `fork_length` column are NA.

Random pick of data entry to compare between db and EDI

  - test 1
```{r echo=FALSE}
# random pick to compare data
recapture_tisdale_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "55215") |>
  glimpse()

recapture_tisdale_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "55215") |>
  glimpse()
```

  - test 2
```{r echo=FALSE}
# random pick to compare data
recapture_tisdale_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "12772") |>
  glimpse()

recapture_tisdale_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "12772") |>
  glimpse()
```

  - test 3
```{r echo=FALSE}
# random pick to compare data
recapture_tisdale_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "41490") |>
  glimpse()

recapture_tisdale_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "41490") |>
  glimpse()
```


  - test 4
```{r echo=FALSE}
# random pick to compare data
recapture_tisdale_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "88918") |>
  glimpse()

recapture_tisdale_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "88918") |>
  glimpse()
```