---
title: "yuba-data-checks"
author: "Inigo Peng, edited by Badhia Yunes-Katz"
date: "`r Sys.Date()`"
output: html_document
---


```{r include=FALSE}
library(DBI)
library(tidyverse)
library(EDIutils)
library(EML)
library(readr)
library(janitor)
load("filter_params.RData")
```

```{r include=FALSE}
con <- DBI::dbConnect(
  RPostgres::Postgres(),
  dbname = "jpedb-staging",
  host = "jpe-db.postgres.database.azure.com",
  port = 5432,
  user = Sys.getenv("jpe_db_user_id"),
  password = Sys.getenv("jpe_db_password")
)
```
# Database Tables
```{r echo=FALSE}
catch <- DBI::dbReadTable(con, "catch")
trap_visit <- DBI::dbReadTable(con, "trap_visit")
release <- DBI::dbReadTable(con, "release")
recaptured_fish <- DBI::dbReadTable(con, "recaptured_fish")
trap_location <- DBI::dbReadTable(con, "trap_location")
released_fish <- DBI::dbReadTable(con, "released_fish")
```

```{r include=FALSE}
trap_id <- trap_location |> 
  rename(trap_location_id = id) |>
  select(trap_location_id, stream, site, subsite) 

catch_with_stream <- catch |>
  left_join(trap_id)
```

## Yuba {.tabset}

### catch

**- Findings catch:**

  - life stage in edi data has not gone through wrangling
  - catch_raw_id in database has a decimal number (e.g 302.0)

```{r echo=FALSE}
catch_yuba_edi <- read_csv("data-raw/edi-zips/yuba_catch.csv") |>
  filter(commonName == "Chinook salmon") |>
  mutate(siteName = tolower(siteName),
         subSiteName = tolower(subSiteName)) |> 
  mutate(
    subSiteName = case_when(
      subSiteName == "yuba river" ~ "yub",
      subSiteName == "hallwood 1 rr" ~ "hal",
      subSiteName == "hallwood 2 rl" ~ "hal2",
      subSiteName == "hallwood 3" ~ "hal3",
      TRUE ~ subSiteName)) |> 
  mutate(stream="yuba river",
         date = as.Date(visitTime))

filter_params_yuba <- filter_params$`yuba river`

for (param in filter_params_yuba){
  catch_yuba_edi <- subset(catch_yuba_edi, 
      !(stream == "yuba river" &
        (
          (is.na(param$site) & is.na(siteName)) |
          (!is.na(param$site) & siteName == param$site)
        ) &
        (
          (is.na(param$subsite) & is.na(subSiteName)) |
          (!is.na(param$subsite) & subSiteName == param$subsite)
        ) &
        date >= as.Date(param$start_date) &
        date <= as.Date(param$end_date)
      )
    )  
}

```

#### Date Range Check
*EDI `date`*
```{r}
range(catch_yuba_edi$date)
```

*DB `date`*
  
```{r echo=FALSE}
catch_yuba_db <- catch_with_stream |> 
  filter(stream == "yuba river") |> 
  filter(!is.na(site) & !is.na(subsite))

range(catch_yuba_db$date)
```

#### ForkLength
*EDI `forkLength`*

* `r round(sum(is.na(catch_yuba_edi$forkLength))/nrow(catch_yuba_edi), 3)*100` % of values in the `forkLength` column are NA.

```{r echo=FALSE}
summary(catch_yuba_edi$forkLength, na.rm = TRUE)
```

*DB `fork_length`*

* `r round(sum(is.na(catch_yuba_db$fork_length))/nrow(catch_yuba_db), 3)*100` % of values in the `fork_length` column are NA.

```{r echo=FALSE}
summary(catch_yuba_db$fork_length, na.rm = TRUE)
```

```{r include=FALSE}
catch_yuba_edi_edited <- catch_yuba_edi |>
  clean_names() |>
  mutate(source = "EDI")
 
catch_yuba_db_edited <- catch_yuba_db |>
  mutate(source = "DB")
```

*Forklength Exploration Plots* 

```{r echo=FALSE, warning=FALSE}
combined_catch_yuba <- catch_yuba_db_edited |> 
  select(date, fork_length, source) |> 
  bind_rows(catch_yuba_edi_edited |> select(date, fork_length, source)) 

ggplot(data = combined_catch_yuba, aes(x = date, y = fork_length, color = source)) +
  geom_point() +
  facet_wrap(~source)
```

#### Lifestage
(life stage in edi data has not gone through wrangling)

*EDI `lifeStage`*

* `r round(sum(is.na(catch_yuba_edi$lifeStage))/nrow(catch_yuba_edi), 3)*100` % of values in the `lifeStage` column from the EDI data are NA.

*DB `lifestage_id`*

* `r round(sum(is.na(catch_yuba_db$lifestage_id))/nrow(catch_yuba_db), 3)*100` % of values in the `lifestage_id` column from the database are NA.


#### Count

*EDI `n`*

* `r round(sum(is.na(catch_yuba_edi$n))/nrow(catch_yuba_edi), 3)*100` % of values in the `n` column in edi data are NA.

```{r echo=FALSE}
summary(catch_yuba_edi$n)
```

*DB `count`*

* `r round(sum(is.na(catch_yuba_db$count))/nrow(catch_yuba_db), 3)*100` % of values in the `count` column in database data are NA.

```{r echo=FALSE}
summary(catch_yuba_db$count)
```

```{r echo=FALSE, warning=FALSE}
ggplot() +
  geom_jitter(data = catch_yuba_db_edited,
              aes(x = date, y = count),
              color = "blue", alpha = 0.4, width = 10, height = 0.5, shape = 16, size = 2) +
  geom_jitter(data = catch_yuba_edi_edited,
              aes(x = date, y = n),
              color = "red", alpha = 0.4, width = 10, height = 0.5, shape = 17, size = 2) +
  theme_minimal() +
  labs(x = "Date",
       y = "count/n")
```

Random pick of data entry to compare between db and EDI

- test 1

```{r echo=FALSE}
# random pick to compare data
catch_yuba_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "302.0") |>
  glimpse()

catch_yuba_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "302") |>
  glimpse()
```

- test 2

```{r echo=FALSE}
# random pick to compare data
catch_yuba_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id  == "18663.0") |>
  glimpse()

catch_yuba_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == 18663) |>
  glimpse()

```

- test 3

```{r echo=FALSE}
# random pick to compare data
catch_yuba_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id  == "1525.0") |>
  glimpse()

catch_yuba_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "1526") |>
  glimpse()
```

### trap

**- Findings:**
   - time ranges vary (on "max" value). db max 2025-05-23 vs EDI 2025-05-27
   - all water_velocity values in database are NA. There is no data entry row with similar name in EDI

```{r echo=FALSE}
trap_with_stream <- trap_visit |>
  left_join(trap_id)

trap_yuba_db <- trap_with_stream |> 
  filter(stream == "yuba river") 
```
```{r echo=FALSE}
trap_yuba_edi <- read_csv("data-raw/edi-zips/yuba_trap.csv") |>
  mutate(siteName = tolower(siteName),
         subSiteName = tolower(subSiteName)) |> 
  mutate(
    subSiteName = case_when(
      subSiteName == "yuba river" ~ "yub",
      subSiteName == "hallwood 1 rr" ~ "hal",
      subSiteName == "hallwood 2 rl" ~ "hal2",
      subSiteName == "hallwood 3" ~ "hal3",
      TRUE ~ subSiteName)) |> 
  mutate(stream="yuba river",
         date = as.Date(visitTime))

filter_params_yuba <- filter_params$`yuba river`
# 
for (param in filter_params_yuba){
  trap_yuba_edi <- subset(trap_yuba_edi,
      !(stream == "yuba river" &
        (
          (is.na(param$site) & is.na(siteName)) |
          (!is.na(param$site) & siteName == param$site)
        ) &
        (
          (is.na(param$subsite) & is.na(subSiteName)) |
          (!is.na(param$subsite) & subSiteName == param$subsite)
        ) &
        (date >= as.Date(param$start_date) &
        date <= as.Date(param$end_date))
      )
    )
}
```

#### Date Range Check

*DB `trap_visit_time_start`* 
```{r echo=FALSE}
range(trap_yuba_db$trap_visit_time_start, na.rm = TRUE)
```
*EDI `visitTime`*
```{r}
range(trap_yuba_edi$visitTime)
```

#### water_velocity

*DB `water_velocity`*

* `r round(sum(is.na(trap_yuba_db$water_velocity))/nrow(trap_yuba_db), 3)*100` % of values in the `water_velocity` column are NA.
```{r echo=FALSE}
summary(trap_yuba_db$water_velocity)
```

*EDI `waterVel`* - no column present


#### fishProcessed

*DB `fish_processed_id`*

* `r round(sum(is.na(trap_yuba_db$fish_processed_id))/nrow(trap_yuba_db), 3)*100` % of values in the `fish_processed_id` column are NA.
```{r echo=FALSE}
table(trap_yuba_db$fish_processed_id)
```

*EDI `fishProcessed`*
* `r round(sum(is.na(trap_yuba_edi$fishProcessed))/nrow(trap_yuba_edi), 3)*100` % of values in the `fishProcessed` column are NA.
```{r echo=FALSE}
table(trap_yuba_edi$fishProcessed)
```

Random pick of data entry by date to compare between db and EDI

  - test 1
```{r echo=FALSE}
# random pick to compare data
trap_yuba_db |>
  mutate(source = "DB") |>
  filter(trap_visit_id  == "986") |>
  glimpse()

trap_yuba_edi |>
  mutate(source = "EDI") |>
  filter(trapVisitID == "986") |>
  glimpse()
```

  - test 2
```{r echo=FALSE}
# random pick to compare data
trap_yuba_db |>
  mutate(source = "DB") |>
  filter(trap_visit_id  == "1270") |>
  glimpse()

trap_yuba_edi |>
  mutate(source = "EDI") |>
  filter(trapVisitID == "1270") |>
  glimpse()
```

  - test 3
```{r echo=FALSE}
# random pick to compare data
trap_yuba_db |>
  mutate(source = "DB") |>
  filter(trap_visit_id  == "549") |>
  glimpse()

trap_yuba_edi |>
  mutate(source = "EDI") |>
  filter(trapVisitID == "549") |>
  glimpse()
```

  - test 4
```{r echo=FALSE}
# random pick to compare data
trap_yuba_db |>
  mutate(source = "DB") |>
  filter(trap_visit_id  == "1071") |>
  glimpse()

trap_yuba_edi |>
  mutate(source = "EDI") |>
  filter(trapVisitID == "1071") |>
  glimpse()
```

### release

**- Findings:**

  - date and time range are the same
  - there is not markedLifeStage column in EDI
  
  
```{r echo=FALSE}
release_with_stream <- release |>
  left_join(trap_id)

release_yuba_db <- release_with_stream |> 
  filter(stream == "yuba river") 
```
```{r echo=FALSE}
release_yuba_edi <- read_csv("data-raw/edi-zips/yuba_release.csv") |> 
  filter(commonName == "Chinook salmon") |>
  mutate(releaseSite = tolower(releaseSite),
         releaseSubSite = tolower(releaseSubSite)) |> 
  mutate(
    releaseSubSite = case_when(
      releaseSubSite == "yuba river" ~ "yub",
      releaseSubSite == "hallwood 1 rr" ~ "hal",
      releaseSubSite == "hallwood 2 rl" ~ "hal2",
      releaseSubSite == "hallwood 3" ~ "hal3",
      TRUE ~ releaseSubSite)) |> 
  mutate(stream="yuba river",
         date = as.Date(releaseTime))

filter_params_yuba <- filter_params$`yuba river`
# 
for (param in filter_params_yuba){
  release_yuba_edi <- subset(release_yuba_edi,
      !(stream == "yuba river" &
        (
          (is.na(param$site) & is.na(releaseSite)) |
          (!is.na(param$site) & releaseSite == param$site)
        ) &
        (
          (is.na(param$subsite) & is.na(releaseSubSite)) |
          (!is.na(param$subsite) & releaseSubSite == param$subsite)
        ) &
        (date >= as.Date(param$start_date) &
        date <= as.Date(param$end_date))
      )
    )
}
```

#### Date Range Check

*DB `date_released`* 
```{r echo=FALSE}
range(release_yuba_db$date_released, na.rm = TRUE)
```
*EDI `releaseTime`*
```{r}
range(release_yuba_edi$releaseTime)
```

#### markedLifeStage

*EDI `markedLifeStage`* - no column name found


*DB `lifestage_id`*

* `r round(sum(is.na(release_yuba_db$lifestage_id))/nrow(release_yuba_db), 3)*100` % of values in the `fork_length` column are NA.

```{r echo=FALSE}
summary(release_yuba_db$lifestage_id, na.rm = TRUE)
```

#### nReleased

*EDI `nReleased`*

* `r round(sum(is.na(release_yuba_edi$nReleased))/nrow(release_yuba_edi), 3)*100` % of values in the `nReleased` column are NA.

```{r echo=FALSE}
summary(release_yuba_edi$nReleased)
```

*DB `number_released`*

* `r round(sum(is.na(release_yuba_db$number_released))/nrow(release_yuba_db), 3)*100` % of values in the `number_released` column are NA.

```{r echo=FALSE}
summary(release_yuba_db$number_released, na.rm = TRUE)
```

*Exploration Plots* 

```{r echo=FALSE}
release_yuba_edi_clean<- release_yuba_edi |>
  mutate(date = as.Date(releaseTime),
         source = "EDI")
release_yuba_db_clean <- release_yuba_db |>
  mutate(date = as.Date(date_released),
         source = "DB")

ggplot() +
  geom_point(data = release_yuba_db_clean,
              aes(x = date, y = number_released, color = source),
              color = "blue", alpha = 0.4, width = 10, height = 0.5, shape = 16, size = 2) +
  geom_point(data = release_yuba_edi_clean,
              aes(x = date, y = nReleased),
              color = "red", alpha = 0.4, width = 10, height = 0.5, shape = 17, size = 2) +
  theme_minimal() +
  labs(x = "Date",
       y = "number_released or nReleased",
       color = "Source"
       )
```

Random pick of data entry by date to compare between db and EDI

  - test 1
```{r echo=FALSE}
# random pick to compare data
release_yuba_db |>
  mutate(source = "DB") |>
  filter(release_id == "297") |>
  glimpse()

release_yuba_edi |>
  mutate(source = "EDI") |>
  filter(releaseID == "297") |>
  glimpse()
```

  - test 2
```{r echo=FALSE}
# random pick to compare data
release_yuba_db |>
  mutate(source = "DB") |>
  filter(release_id == "257") |>
  glimpse()

release_yuba_edi |>
  mutate(source = "EDI") |>
  filter(releaseID == "257") |>
  glimpse()
```

  - test 3
```{r echo=FALSE}
# random pick to compare data
release_yuba_db |>
  mutate(source = "DB") |>
  filter(release_id == "259") |>
  glimpse()

release_yuba_edi |>
  mutate(source = "EDI") |>
  filter(releaseID == "259") |>
  glimpse()
```

  - test 4
```{r echo=FALSE}
# random pick to compare data
release_yuba_db |>
  mutate(source = "DB") |>
  filter(release_id == "262") |>
  glimpse()

release_yuba_edi |>
  mutate(source = "EDI") |>
  filter(releaseID == "262") |>
  glimpse()
```

### recapture

**- Findings:**

  - Same date range


```{r echo=FALSE}
recapture_with_stream <- recaptured_fish |>
  left_join(trap_id)

recapture_yuba_db <- recapture_with_stream |> 
  filter(stream == "yuba river") 
```
```{r echo=FALSE}
recapture_yuba_edi <- read_csv("data-raw/edi-zips/yuba_recapture.csv") |>
  filter(commonName == "Chinook salmon") |>
  mutate(siteName = tolower(siteName),
         subSiteName = tolower(subSiteName)) |> 
  mutate(
    subSiteName = case_when(
      subSiteName == "yuba river" ~ "yub",
      subSiteName == "hallwood 1 rr" ~ "hal",
      subSiteName == "hallwood 2 rl" ~ "hal2",
      subSiteName == "hallwood 3" ~ "hal3",
      TRUE ~ subSiteName)) |> 
  mutate(stream="yuba river",
         date = as.Date(visitTime))

filter_params_yuba <- filter_params$`yuba river`
# 
for (param in filter_params_yuba){
  recapture_yuba_edi <- subset(recapture_yuba_edi,
      !(stream == "yuba river" &
        (
          (is.na(param$site) & is.na(siteName)) |
          (!is.na(param$site) & siteName == param$site)
        ) &
        (
          (is.na(param$subsite) & is.na(subSiteName)) |
          (!is.na(param$subsite) & subSiteName == param$subsite)
        ) &
        (date >= as.Date(param$start_date) &
        date <= as.Date(param$end_date))
      )
    )
}
```

#### Date Range Check

*DB `date`* 
```{r echo=FALSE}
range(recapture_yuba_db$date)
```
*EDI `date`*
```{r}
range(recapture_yuba_edi$date)
```

#### Count

*DB `count`* 

* `r round(sum(is.na(recapture_yuba_db$count))/nrow(recapture_yuba_db), 3)*100` % of values in the `count` column are NA.
```{r echo=FALSE}
summary(recapture_yuba_db$count, na.rm = TRUE)
```

*EDI `n`*

* `r round(sum(is.na(recapture_yuba_edi$n))/nrow(recapture_yuba_edi), 3)*100` % of values in the `n` column are NA.
```{r}
summary(recapture_yuba_edi$n)
```

#### lifeStage

*EDI `lifeStage`*

* `r round(sum(is.na(recapture_yuba_edi$lifeStage))/nrow(recapture_yuba_edi), 3)*100` % of values in the `lifeStage` column are NA.

```{r echo=FALSE}
table(recapture_yuba_edi$lifeStage)
```

*DB `lifestage_id`*

* `r round(sum(is.na(recapture_yuba_db$lifestage_id))/nrow(recapture_yuba_db), 3)*100` % of values in the `lifestage_id` column are NA.

```{r echo=FALSE}
summary(recapture_yuba_db$lifestage_id, na.rm = TRUE)
```


```{r echo=FALSE, warning=FALSE}
ggplot() +
  geom_point(data = recapture_yuba_db,
              aes(x = date, y = count),
              color = "blue", alpha = 0.4, width = 10, height = 0.5, shape = 16, size = 2) +
  geom_point(data = recapture_yuba_edi,
              aes(x = as.Date(visitTime), y = n),
              color = "red", alpha = 0.4, width = 10, height = 0.5, shape = 17, size = 2) +
  theme_minimal() +
  labs(x = "Date Recaptured",
       y = "count/number_recaptured")
```

#### forkLength

*EDI `forkLength`*

* `r round(sum(is.na(recapture_yuba_edi$forkLength))/nrow(recapture_yuba_edi), 3)*100` % of values in the `forkLength` column are NA.

*DB `fork_length`*

* `r round(sum(is.na(recapture_yuba_db$fork_length))/nrow(recapture_yuba_db), 3)*100` % of values in the `fork_length` column are NA.

Random pick of data entry to compare between db and EDI

  - test 1
```{r echo=FALSE}
# random pick to compare data
recapture_yuba_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "6045") |>
  glimpse()

recapture_yuba_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "6045") |>
  glimpse()
```

  - test 2
```{r echo=FALSE}
# random pick to compare data
recapture_yuba_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "16612") |>
  glimpse()

recapture_yuba_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "16612") |>
  glimpse()
```

  - test 3
```{r echo=FALSE}
# random pick to compare data
recapture_yuba_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "19678") |>
  glimpse()

recapture_yuba_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "19678") |>
  glimpse()
```

  - test 4
```{r echo=FALSE}
# random pick to compare data
recapture_yuba_db |>
  mutate(source = "DB") |>
  filter(catch_raw_id == "26312") |>
  glimpse()

recapture_yuba_edi |>
  mutate(source = "EDI") |>
  filter(catchRawID == "26312") |>
  glimpse()
```
