---
title: "Efficiency Trial Annual QC"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    embed-resources: true
params:
  review_run_year: 2026
---

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(lubridate)
library(here)
library(glue)
library(DT)

if (file.exists(here::here("DESCRIPTION"))) {
  devtools::load_all(here::here(), quiet = TRUE)
} else {
  library(SRJPEdata)
}

source(here::here("data-raw", "qc", "qc_helpers.R"))

review_run_year <- params$review_run_year
```

## Overview

Reviewing efficiency trial data. Run year being reviewed: **`r review_run_year`**.

Data sources: `weekly_efficiency` (raw trial data), `weekly_juvenile_abundance_efficiency_data` (model-ready), `years_to_exclude_rst_data`.

```{r}
#| label: overview-table
#| echo: false

weekly_efficiency |>
  mutate(run_year = as_run_year(
    as.Date(paste0(year_released, "-01-01") ) + lubridate::weeks(week_released - 1)
  )) |>
  group_by(stream, site) |>
  summarise(
    min_run_year   = min(run_year, na.rm = TRUE),
    max_run_year   = max(run_year, na.rm = TRUE),
    total_trials   = n(),
    total_released = sum(number_released, na.rm = TRUE),
    .groups        = "drop"
  ) |>
  arrange(stream, site) |>
  knitr::kable(caption = "Efficiency trial records by stream and site")
```

```{r}
#| label: prep-data
#| echo: false

# Add run_year to weekly_efficiency using week/year convention
eff <- weekly_efficiency |>
  mutate(run_year = if_else(week_released >= 45,
                             year_released + 1L,
                             year_released))
```

---

## Check 1: Recaptures Exceeding Releases

Flag trials where `number_recaptured > number_released`. These rows are already filtered out during the model data build, but logging them creates a traceable record for upstream correction.

```{r}
#| label: check-recaptures-exceed

recap_exceed <- eff |>
  filter(number_recaptured > number_released) |>
  group_by(stream, site, run_year) |>
  summarise(
    n_records    = n(),
    max_ratio    = round(max(number_recaptured / number_released), 2),
    .groups      = "drop"
  )

new_recap_exceed <- log_issues(
  recap_exceed |>
    mutate(
      data_type   = "efficiency",
      issue_type  = "recaptures_exceed_releases",
      field       = "number_recaptured",
      severity    = "critical",
      description = glue("{n_records} trial(s) in run year {run_year} at {site} ({stream}) have more recaptures than releases. Max recapture/release ratio: {max_ratio}. These rows are filtered at build time.")
    )
)
```

```{r}
#| label: plot-release-recapture
#| echo: false
#| fig-height: 5

eff_plot <- eff |>
  mutate(flagged = number_recaptured > number_released)

ggplot(eff_plot, aes(x = number_released, y = number_recaptured, color = flagged)) +
  geom_point(alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey40") +
  scale_color_manual(values = c("FALSE" = "steelblue", "TRUE" = "#d73027"),
                     labels = c("Normal", "Flagged"), name = NULL) +
  facet_wrap(~stream, scales = "free", ncol = 2) +
  labs(title = "Releases vs. recaptures per trial (dashed = 1:1 line)",
       x = "Number released", y = "Number recaptured") +
  theme_minimal()
```

`r if (nrow(recap_exceed) == 0) "No trials with recaptures exceeding releases." else glue("{sum(recap_exceed$n_records)} trial(s) flagged across {nrow(recap_exceed)} site/run-year combination(s).")`

---

## Check 2: Low Release Count

Flag trials where fewer than 50 fish were released. Small releases produce high-variance efficiency estimates.

```{r}
#| label: check-low-release

low_release <- eff |>
  filter(number_released < 50) |>
  group_by(stream, site, run_year) |>
  summarise(
    n_records   = n(),
    min_release = min(number_released),
    .groups     = "drop"
  )

new_low_release <- log_issues(
  low_release |>
    mutate(
      data_type   = "efficiency",
      issue_type  = "low_release_count",
      field       = "number_released",
      severity    = "minor",
      description = glue("{n_records} trial(s) with <50 fish released in run year {run_year} at {site} ({stream}). Minimum release count: {min_release}")
    )
)
```

```{r}
#| label: plot-release-histogram
#| echo: false
#| fig-height: 5

ggplot(eff, aes(x = number_released, fill = number_released < 50)) +
  geom_histogram(bins = 40, alpha = 0.8) +
  scale_fill_manual(values = c("FALSE" = "steelblue", "TRUE" = "#fdae61"),
                    labels = c(">= 50", "< 50 (flagged)"), name = NULL) +
  geom_vline(xintercept = 50, linetype = "dashed", color = "#d73027") +
  facet_wrap(~stream, scales = "free_y", ncol = 2) +
  labs(title = "Distribution of release counts (dashed = 50 fish threshold)",
       x = "Number released", y = "Count") +
  theme_minimal()
```

`r if (nrow(low_release) == 0) "No low-release trials detected." else glue("{sum(low_release$n_records)} trial(s) with <50 releases flagged.")`

---

## Check 3: Zero Recaptures

Flag trials with zero recaptures. Can be real (very low efficiency) but may also indicate a data entry issue.

```{r}
#| label: check-zero-recaptures

zero_recap <- eff |>
  filter(number_recaptured == 0) |>
  group_by(stream, site, run_year) |>
  summarise(
    n_records = n(),
    .groups   = "drop"
  )

new_zero_recap <- log_issues(
  zero_recap |>
    mutate(
      data_type   = "efficiency",
      issue_type  = "zero_recaptures",
      field       = "number_recaptured",
      severity    = "minor",
      description = glue("{n_records} trial(s) with 0 recaptures in run year {run_year} at {site} ({stream})")
    )
)
```

```{r}
#| label: plot-zero-recaptures
#| echo: false
#| fig-height: 6

zero_recap_by_year <- eff |>
  group_by(stream, site, run_year) |>
  summarise(
    total_trials      = n(),
    zero_recap_trials = sum(number_recaptured == 0),
    .groups           = "drop"
  )

ggplot(zero_recap_by_year, aes(x = run_year, y = zero_recap_trials)) +
  geom_col(fill = "#fdae61") +
  facet_wrap(~stream + site, scales = "free_y", ncol = 2) +
  labs(title = "Trials with zero recaptures by site and run year",
       x = "Run year", y = "Number of zero-recapture trials") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

`r if (nrow(zero_recap) == 0) "No zero-recapture trials detected." else glue("{sum(zero_recap$n_records)} zero-recapture trial(s) flagged.")`

---

## Check 4: Trial Coverage Per Run Year

Flag stream/site/run_years with fewer than 3 efficiency trials in the BTSPAS window (weeks 45–53 and 1–22). Indicates sparse coverage that may limit model reliability.

```{r}
#| label: check-trial-coverage

trial_coverage <- weekly_juvenile_abundance_efficiency_data |>
  filter(week %in% c(45:53, 1:22)) |>
  group_by(stream, site, run_year) |>
  summarise(n_trials = n(), .groups = "drop")

excluded_lookup <- years_to_exclude_rst_data |>
  distinct(stream, site, run_year) |>
  mutate(already_excluded = TRUE)

low_coverage <- trial_coverage |>
  filter(n_trials < 3) |>
  left_join(excluded_lookup, by = c("stream", "site", "run_year")) |>
  mutate(already_excluded = replace_na(already_excluded, FALSE))

new_low_coverage <- log_issues(
  low_coverage |>
    mutate(
      n_records   = n_trials,
      data_type   = "efficiency",
      issue_type  = "low_trial_coverage",
      field       = "n_trials",
      severity    = "moderate",
      description = glue("Only {n_trials} efficiency trial(s) in BTSPAS window for run year {run_year} at {site} ({stream}). Already excluded from modeling: {already_excluded}")
    )
)
```

```{r}
#| label: plot-trial-coverage
#| echo: false
#| fig-height: 8

coverage_plot <- trial_coverage |>
  left_join(excluded_lookup, by = c("stream", "site", "run_year")) |>
  mutate(
    already_excluded = replace_na(already_excluded, FALSE),
    label = case_when(
      n_trials >= 3              ~ "Adequate (>=3)",
      n_trials < 3 & already_excluded  ~ "Low (excluded)",
      n_trials < 3 & !already_excluded ~ "Low (not excluded)"
    )
  )

ggplot(coverage_plot, aes(x = run_year, y = site, fill = label)) +
  geom_tile(color = "white") +
  scale_fill_manual(
    values = c("Adequate (>=3)"        = "steelblue",
               "Low (excluded)"        = "#fdae61",
               "Low (not excluded)"    = "#d73027"),
    name = NULL
  ) +
  facet_wrap(~stream, scales = "free_y", ncol = 2) +
  labs(title = "Efficiency trial count in BTSPAS window by site and run year",
       x = "Run year", y = NULL) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
```

`r if (nrow(low_coverage) == 0) "No low-coverage run years detected." else glue("{nrow(low_coverage)} site/run-year combination(s) flagged ({sum(low_coverage$already_excluded)} already excluded from modeling).")`

---

## Check 5: Efficiency Rate Outliers

For each site with at least 5 years of history, flag run_years where the mean simple efficiency (`number_recaptured / number_released`) is more than 3 SD above the site historical mean.

```{r}
#| label: check-efficiency-outliers

eff_rates <- eff |>
  filter(number_released > 0, number_recaptured <= number_released) |>
  mutate(simple_eff = number_recaptured / number_released) |>
  group_by(stream, site, run_year) |>
  summarise(mean_eff = mean(simple_eff, na.rm = TRUE), .groups = "drop")

site_eff_stats <- eff_rates |>
  group_by(stream, site) |>
  summarise(
    n_years  = n(),
    hist_mean = mean(mean_eff, na.rm = TRUE),
    hist_sd   = sd(mean_eff, na.rm = TRUE),
    .groups   = "drop"
  ) |>
  filter(n_years >= 5)

eff_outliers <- eff_rates |>
  inner_join(site_eff_stats, by = c("stream", "site")) |>
  filter(mean_eff > hist_mean + 3 * hist_sd)

new_eff_outliers <- log_issues(
  eff_outliers |>
    mutate(
      n_records   = 1L,
      data_type   = "efficiency",
      issue_type  = "extreme_value",
      field       = "number_recaptured",
      severity    = "minor",
      description = glue("Mean efficiency of {round(mean_eff, 3)} in run year {run_year} at {site} ({stream}) is >3 SD above site historical mean ({round(hist_mean, 3)}, SD = {round(hist_sd, 3)})")
    )
)
```

```{r}
#| label: plot-efficiency-timeseries
#| echo: false
#| fig-height: 8

flagged_eff <- eff_outliers |> select(stream, site, run_year)

ggplot(eff_rates, aes(x = run_year, y = mean_eff)) +
  geom_line(color = "grey60") +
  geom_point(
    aes(color = paste(stream, site, run_year) %in%
          paste(flagged_eff$stream, flagged_eff$site, flagged_eff$run_year)),
    size = 2
  ) +
  scale_color_manual(values = c("FALSE" = "steelblue", "TRUE" = "#d73027"),
                     labels = c("Normal", "Flagged"), name = NULL) +
  facet_wrap(~stream + site, scales = "free_y", ncol = 2) +
  labs(title = "Mean annual simple efficiency by site with flagged years",
       x = "Run year", y = "Mean efficiency (recaptured / released)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

`r if (nrow(eff_outliers) == 0) "No efficiency rate outliers detected." else glue("{nrow(eff_outliers)} site/run-year combination(s) flagged.")`

---

## Issues Logged This Run

```{r}
#| label: issues-summary
#| echo: false

all_new <- bind_rows(
  new_recap_exceed, new_low_release, new_zero_recap,
  new_low_coverage, new_eff_outliers
)

if (nrow(all_new) == 0) {
  cat("No new issues logged this run.")
} else {
  DT::datatable(
    all_new |> select(log_id, stream, site, run_year, issue_type, field, severity, description),
    caption = glue("{nrow(all_new)} new issue(s) added to qc_log.csv"),
    options = list(pageLength = 25, scrollX = TRUE),
    rownames = FALSE
  )
}
```

### All Open Efficiency Issues

```{r}
#| label: open-issues
#| echo: false

full_log   <- readr::read_csv(here::here("data-raw", "qc", "qc_log.csv"), show_col_types = FALSE)
open_eff   <- full_log |> filter(data_type == "efficiency", status == "open")

if (nrow(open_eff) == 0) {
  cat("No open efficiency issues in the log.")
} else {
  DT::datatable(
    open_eff |> select(log_id, date_identified, stream, site, run_year,
                       issue_type, field, severity, description, reviewer_notes),
    caption = glue("{nrow(open_eff)} open efficiency issue(s)"),
    options = list(pageLength = 25, scrollX = TRUE),
    rownames = FALSE
  )
}
```
