---
title: "Flow & Temperature Annual QC"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    embed-resources: true
params:
  review_run_year: 2026
---

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(lubridate)
library(here)
library(glue)
library(DT)

if (file.exists(here::here("DESCRIPTION"))) {
  devtools::load_all(here::here(), quiet = TRUE)
} else {
  library(SRJPEdata)
}

source(here::here("data-raw", "qc", "qc_helpers.R"))

review_run_year <- params$review_run_year

EXPECTED_STREAMS <- c(
  "battle creek", "butte creek", "clear creek", "deer creek",
  "feather river", "mill creek", "sacramento river", "yuba river"
)

# RST season weeks (BTSPAS window: weeks 45-53 then 1-22)
RST_WEEKS <- c(45:53, 1:22)
```

## Overview

Reviewing flow and temperature data. Run year being reviewed: **`r review_run_year`**.

Data source: `environmental_data` — weekly data with columns `year`, `week`, `stream`, `site_group`, `parameter`, `statistic`, `value`.

```{r}
#| label: overview-table
#| echo: false

environmental_data |>
  group_by(stream, parameter, statistic) |>
  summarise(
    min_year  = min(year, na.rm = TRUE),
    max_year  = max(year, na.rm = TRUE),
    n_records = n(),
    .groups   = "drop"
  ) |>
  arrange(stream, parameter, statistic) |>
  knitr::kable(caption = "Environmental data coverage by stream, parameter, and statistic")
```

```{r}
#| label: prep-data
#| echo: false

# Add run_year: week >= 45 belongs to the following calendar year's run
env <- environmental_data |>
  mutate(run_year = if_else(week >= 45L, year + 1L, year))

flow_data <- env |> filter(parameter == "flow", statistic == "mean")
temp_data <- env |> filter(parameter == "temperature", statistic == "mean")
```

---

## Check 1: Stream Coverage

Verify all expected streams have flow data for the most recent run year in the dataset.

```{r}
#| label: check-stream-coverage

current_run_year_data <- max(flow_data$run_year, na.rm = TRUE)

stream_coverage <- flow_data |>
  filter(run_year == current_run_year_data) |>
  group_by(stream) |>
  summarise(
    n_weeks  = n(),
    max_year = max(year),
    max_week = max(week[year == max(year)]),
    .groups  = "drop"
  )

missing_streams <- tibble(stream = EXPECTED_STREAMS) |>
  anti_join(stream_coverage, by = "stream")

new_missing_streams <- log_issues(
  missing_streams |>
    mutate(
      site        = NA_character_,
      run_year    = current_run_year_data,
      n_records   = 0L,
      data_type   = "flow",
      issue_type  = "gap",
      field       = "stream_coverage",
      severity    = "critical",
      description = glue("Stream '{stream}' has no flow data in run year {current_run_year_data}")
    )
)
```

```{r}
#| label: table-stream-coverage
#| echo: false

tibble(stream = EXPECTED_STREAMS) |>
  left_join(stream_coverage, by = "stream") |>
  mutate(
    present  = if_else(is.na(n_weeks), "MISSING", "Present"),
    last_obs = if_else(is.na(max_year), NA_character_,
                       glue("year {max_year} week {max_week}"))
  ) |>
  select(stream, present, n_weeks, last_obs) |>
  knitr::kable(caption = glue("Flow coverage in run year {current_run_year_data}"))
```

`r if (nrow(missing_streams) == 0) "All expected streams present." else glue("{nrow(missing_streams)} stream(s) missing from run year {current_run_year_data}.")`

---

## Check 2: Flow — Missing Weeks in RST Season Window

For each stream/run_year, count how many weeks in the BTSPAS window (weeks 45–53 and 1–22) are missing flow data. Flag run_years where >20% of the 31 possible window weeks are missing (`gap`) or where 3+ consecutive weeks are missing (`extended_gap`).

```{r}
#| label: check-flow-gaps

# Helper: find max consecutive missing weeks given present weeks and expected weeks
max_consec_missing <- function(weeks_present, expected) {
  missing <- expected[!expected %in% weeks_present]
  if (length(missing) == 0L) return(0L)
  # For consecutive detection, treat the season as a linear sequence
  # Map to position in expected vector
  pos <- which(expected %in% missing)
  max(tabulate(cumsum(c(1L, diff(pos) != 1L))))
}

# Weeks present per stream/run_year (within RST window)
weeks_present <- flow_data |>
  filter(week %in% RST_WEEKS) |>
  group_by(stream, site_group, run_year) |>
  summarise(
    weeks_present   = list(week),
    n_present       = n(),
    .groups         = "drop"
  ) |>
  mutate(
    n_expected       = 31L,  # weeks 45-53 (9) + 1-22 (22)
    n_missing        = n_expected - n_present,
    pct_missing      = round(n_missing / n_expected * 100, 1),
    max_consec       = map_int(weeks_present, ~max_consec_missing(.x, RST_WEEKS))
  )

gap_issues     <- weeks_present |> filter(pct_missing > 20, max_consec < 3)
ext_gap_issues <- weeks_present |> filter(max_consec >= 3)

new_gaps <- log_issues(
  gap_issues |>
    mutate(
      site        = site_group,
      data_type   = "flow",
      issue_type  = "gap",
      field       = "week",
      n_records   = n_missing,
      severity    = "moderate",
      description = glue("{n_missing} of 31 RST window weeks ({pct_missing}%) missing flow data in run year {run_year} for {stream}")
    )
)

new_ext_gaps <- log_issues(
  ext_gap_issues |>
    mutate(
      site        = site_group,
      data_type   = "flow",
      issue_type  = "extended_gap",
      field       = "week",
      n_records   = max_consec,
      severity    = "critical",
      description = glue("{max_consec} consecutive RST window weeks missing flow data in run year {run_year} for {stream} ({n_missing} total missing weeks)")
    )
)
```

```{r}
#| label: plot-flow-availability
#| echo: false
#| fig-height: 8

# Show weekly flow availability by site_group, faceted by stream
avail_plot <- flow_data |>
  filter(week %in% RST_WEEKS) |>
  count(stream, site_group, run_year, name = "n_weeks_present") |>
  mutate(pct_present = round(n_weeks_present / 31 * 100))

ggplot(avail_plot, aes(x = run_year, y = site_group, fill = pct_present)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "#fee08b", high = "steelblue",
                      name = "% weeks\npresent", limits = c(0, 100)) +
  facet_wrap(~stream, scales = "free_y", ncol = 2) +
  labs(title = "Flow data coverage in RST season window (weeks 45–53 and 1–22)",
       x = "Run year", y = NULL) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(face = "bold"))
```

`r if (nrow(gap_issues) + nrow(ext_gap_issues) == 0) "No missing-week issues detected." else glue("{nrow(gap_issues)} gap(s) and {nrow(ext_gap_issues)} extended gap(s) flagged.")`

---

## Check 3: Flow — Zero or Negative Values

Flag stream/run_years with any flow value ≤ 0 in the RST season window. Zero flow during RST season is almost certainly a sensor or data issue.

```{r}
#| label: check-zero-flow

zero_flow <- flow_data |>
  filter(week %in% RST_WEEKS, value <= 0) |>
  group_by(stream, site_group, run_year) |>
  summarise(
    n_records = n(),
    min_value = min(value),
    .groups   = "drop"
  )

new_zero_flow <- log_issues(
  zero_flow |>
    mutate(
      site        = site_group,
      data_type   = "flow",
      issue_type  = "implausible_value",
      field       = "value",
      severity    = "critical",
      description = glue("{n_records} week(s) with flow <= 0 (min = {min_value} cfs) during RST season window in run year {run_year} for {stream}")
    )
)
```

`r if (nrow(zero_flow) == 0) "No zero or negative flow values detected." else glue("{nrow(zero_flow)} stream/run-year combination(s) flagged.")`

---

## Check 4: Temperature — Implausible Values

Flag stream/run_years with any weekly mean temperature outside the plausible range (< −5°C or > 35°C).

```{r}
#| label: check-temp-implausible

temp_impl <- temp_data |>
  filter(value < -5 | value > 35) |>
  group_by(stream, site_group, run_year) |>
  summarise(
    n_records = n(),
    min_temp  = round(min(value), 1),
    max_temp  = round(max(value), 1),
    .groups   = "drop"
  )

new_temp_impl <- log_issues(
  temp_impl |>
    mutate(
      site        = site_group,
      data_type   = "flow",
      issue_type  = "implausible_value",
      field       = "value",
      severity    = "critical",
      description = glue("{n_records} weekly temperature record(s) outside [-5, 35]°C in run year {run_year} for {stream}. Range: [{min_temp}, {max_temp}]°C")
    )
)
```

```{r}
#| label: plot-temp-distribution
#| echo: false
#| fig-height: 6

temp_plot <- temp_data |>
  mutate(implausible = value < -5 | value > 35)

ggplot(temp_plot, aes(x = value, fill = implausible)) +
  geom_histogram(bins = 60, alpha = 0.8) +
  scale_fill_manual(values = c("FALSE" = "steelblue", "TRUE" = "#d73027"),
                    labels = c("Normal", "Flagged"), name = NULL) +
  geom_vline(xintercept = c(-5, 35), linetype = "dashed", color = "#d73027") +
  facet_wrap(~stream, scales = "free_y", ncol = 2) +
  labs(title = "Weekly temperature distribution with implausible values flagged",
       x = "Temperature (°C)", y = "Count") +
  theme_minimal()
```

`r if (nrow(temp_impl) == 0) "No implausible temperature values detected." else glue("{nrow(temp_impl)} stream/run-year combination(s) flagged.")`

---

## Issues Logged This Run

```{r}
#| label: issues-summary
#| echo: false

all_new <- bind_rows(new_missing_streams, new_gaps, new_ext_gaps,
                     new_zero_flow, new_temp_impl)

if (nrow(all_new) == 0) {
  cat("No new issues logged this run.")
} else {
  DT::datatable(
    all_new |> select(log_id, stream, site, run_year, issue_type, field, severity, description),
    caption = glue("{nrow(all_new)} new issue(s) added to qc_log.csv"),
    options = list(pageLength = 25, scrollX = TRUE),
    rownames = FALSE
  )
}
```

### All Open Flow Issues

```{r}
#| label: open-issues
#| echo: false

full_log  <- readr::read_csv(here::here("data-raw", "qc", "qc_log.csv"), show_col_types = FALSE)
open_flow <- full_log |> filter(data_type == "flow", status == "open")

if (nrow(open_flow) == 0) {
  cat("No open flow issues in the log.")
} else {
  DT::datatable(
    open_flow |> select(log_id, date_identified, stream, site, run_year,
                        issue_type, field, severity, description, reviewer_notes),
    caption = glue("{nrow(open_flow)} open flow issue(s)"),
    options = list(pageLength = 25, scrollX = TRUE),
    rownames = FALSE
  )
}
```
