---
title: "Genetics Data Annual QC"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    embed-resources: true
params:
  review_run_year: 2026
---

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(lubridate)
library(here)
library(glue)
library(DT)

if (file.exists(here::here("DESCRIPTION"))) {
  devtools::load_all(here::here(), quiet = TRUE)
} else {
  library(SRJPEdata)
}

source(here::here("data-raw", "qc", "qc_helpers.R"))

review_run_year <- params$review_run_year

# Check whether genetics data is available in this build
genetics_available <- exists("completed_genetic_samples")
```

## Status

```{r}
#| echo: false

if (!genetics_available) {
  cat("Genetics data (completed_genetic_samples) is not yet integrated into this build.\n")
  cat("QC checks are stubbed out and will be activated once the dataset is available.\n")
  cat("See data-raw/pull_data_scripts/pull_genetic_run_assignment_data.R for the pull script.")
} else {
  cat(glue("Genetics data available. {nrow(completed_genetic_samples)} records found.\n"))
}
```

`r if (!genetics_available) knitr::knit_exit()`

---

## Overview

Run year being reviewed: **`r review_run_year`**.

Data source: `completed_genetic_samples` — columns include `sample_id`, `datetime_collected`, `fork_length_mm`, `sherlock_run_assignment`, `field_run_assignment`, `stream_name`.

```{r}
#| label: overview-table
#| echo: false

completed_genetic_samples |>
  mutate(year = lubridate::year(datetime_collected)) |>
  group_by(stream_name, year) |>
  summarise(n_samples = n(), .groups = "drop") |>
  arrange(stream_name, year) |>
  knitr::kable(caption = "Genetics samples by stream and year")
```

---

## Check 1: Sample Count — Low by Stream/Year

Flag stream/years with fewer samples than expected. Expected counts to be defined based on monitoring program targets (placeholder threshold: < 30 samples).

```{r}
#| label: check-sample-count

sample_counts <- completed_genetic_samples |>
  mutate(
    run_year = as_run_year(as.Date(datetime_collected))
  ) |>
  group_by(stream_name, run_year) |>
  summarise(n_samples = n(), .groups = "drop")

# TODO: replace placeholder threshold with program-specific targets
low_samples <- sample_counts |>
  filter(n_samples < 30)

new_low_samples <- log_issues(
  low_samples |>
    rename(stream = stream_name) |>
    mutate(
      site        = NA_character_,
      n_records   = n_samples,
      data_type   = "genetics",
      issue_type  = "low_sample_count",
      field       = "sample_id",
      severity    = "moderate",
      description = glue("Only {n_samples} sample(s) in run year {run_year} for {stream} (threshold: 30). Verify against program targets.")
    )
)
```

`r if (nrow(low_samples) == 0) "No low sample counts detected." else glue("{nrow(low_samples)} stream/run-year combination(s) flagged.")`

---

## Check 2: Missing Key Fields

Flag stream/years with high NA rates in `fork_length_mm`, `sherlock_run_assignment`, or `datetime_collected`.

```{r}
#| label: check-missing-fields

missing_fields <- completed_genetic_samples |>
  mutate(run_year = as_run_year(as.Date(datetime_collected))) |>
  group_by(stream_name, run_year) |>
  summarise(
    n_records             = n(),
    pct_na_fork_length    = round(mean(is.na(fork_length_mm)) * 100, 1),
    pct_na_sherlock       = round(mean(is.na(sherlock_run_assignment)) * 100, 1),
    pct_na_date           = round(mean(is.na(datetime_collected)) * 100, 1),
    .groups               = "drop"
  )

fl_issues <- missing_fields |>
  filter(pct_na_fork_length > 50) |>
  mutate(
    stream      = stream_name, site = NA_character_,
    data_type   = "genetics", issue_type = "high_na_rate", field = "fork_length_mm",
    severity    = "moderate",
    description = glue("{pct_na_fork_length}% of {n_records} sample(s) missing fork_length_mm in run year {run_year} for {stream_name}")
  )

sherlock_issues <- missing_fields |>
  filter(pct_na_sherlock > 50) |>
  mutate(
    stream      = stream_name, site = NA_character_,
    data_type   = "genetics", issue_type = "high_na_rate", field = "sherlock_run_assignment",
    severity    = "moderate",
    description = glue("{pct_na_sherlock}% of {n_records} sample(s) missing sherlock_run_assignment in run year {run_year} for {stream_name}")
  )

new_missing_fields <- bind_rows(
  log_issues(fl_issues),
  log_issues(sherlock_issues)
)
```

---

## Check 3: Run Assignment Mismatch

For records with both Sherlock and field assignment, compute the mismatch rate by stream/year. Flag stream/years where mismatch rate is notably higher than the site historical average.

```{r}
#| label: check-run-mismatch

mismatch <- completed_genetic_samples |>
  filter(!is.na(sherlock_run_assignment), !is.na(field_run_assignment)) |>
  mutate(
    run_year  = as_run_year(as.Date(datetime_collected)),
    mismatch  = sherlock_run_assignment != field_run_assignment
  ) |>
  group_by(stream_name, run_year) |>
  summarise(
    n_compared       = n(),
    n_mismatch       = sum(mismatch),
    pct_mismatch     = round(mean(mismatch) * 100, 1),
    .groups          = "drop"
  )

site_mismatch_stats <- mismatch |>
  group_by(stream_name) |>
  summarise(
    n_years       = n(),
    hist_mean_pct = mean(pct_mismatch),
    hist_sd_pct   = sd(pct_mismatch),
    .groups       = "drop"
  ) |>
  filter(n_years >= 3)

mismatch_outliers <- mismatch |>
  inner_join(site_mismatch_stats, by = "stream_name") |>
  filter(pct_mismatch > hist_mean_pct + 2 * hist_sd_pct)

new_mismatch <- log_issues(
  mismatch_outliers |>
    mutate(
      stream      = stream_name,
      site        = NA_character_,
      n_records   = n_mismatch,
      data_type   = "genetics",
      issue_type  = "run_assignment_mismatch",
      field       = "sherlock_run_assignment",
      severity    = "minor",
      description = glue("{pct_mismatch}% mismatch rate ({n_mismatch}/{n_compared} samples) in run year {run_year} for {stream_name}. Historical mean: {round(hist_mean_pct, 1)}%")
    )
)
```

---

## Issues Logged This Run

```{r}
#| label: issues-summary
#| echo: false

all_new <- bind_rows(new_low_samples, new_missing_fields, new_mismatch)

if (nrow(all_new) == 0) {
  cat("No new issues logged this run.")
} else {
  DT::datatable(
    all_new |> select(log_id, stream, site, run_year, issue_type, field, severity, description),
    caption = glue("{nrow(all_new)} new issue(s) added to qc_log.csv"),
    options = list(pageLength = 25, scrollX = TRUE),
    rownames = FALSE
  )
}
```

### All Open Genetics Issues

```{r}
#| label: open-issues
#| echo: false

full_log     <- readr::read_csv(here::here("data-raw", "qc", "qc_log.csv"), show_col_types = FALSE)
open_genetics <- full_log |> filter(data_type == "genetics", status == "open")

if (nrow(open_genetics) == 0) {
  cat("No open genetics issues in the log.")
} else {
  DT::datatable(
    open_genetics |> select(log_id, date_identified, stream, site, run_year,
                            issue_type, field, severity, description, reviewer_notes),
    caption = glue("{nrow(open_genetics)} open genetics issue(s)"),
    options = list(pageLength = 25, scrollX = TRUE),
    rownames = FALSE
  )
}
```
