---
title: "QC Summary: This Year vs. Historical Average"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    embed-resources: true
params:
  review_run_year: 2026
---

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(lubridate)
library(here)
library(glue)
library(DT)

if (file.exists(here::here("DESCRIPTION"))) {
  devtools::load_all(here::here(), quiet = TRUE)
} else {
  library(SRJPEdata)
}

source(here::here("data-raw", "qc", "qc_helpers.R"))

review_run_year <- params$review_run_year

RST_WEEKS <- c(45:53, 1:22)
```

## Overview

This report compares **run year `r review_run_year`** against the historical average across all other run years present in the underlying data. 

```{r}
#| label: rst-metrics
#| echo: false

rst_fl_metrics <- rst_catch |>
  filter(!is.na(date)) |>
  mutate(run_year = as_run_year(date)) |>
  group_by(run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(is.na(fork_length)), .groups = "drop") |>
  mutate(data_type = "rst", issue_type = "high_na_rate", field = "fork_length")

rst_fl_impl_metrics <- rst_catch |>
  filter(!is.na(fork_length), fork_length > 0) |>
  mutate(run_year = as_run_year(date)) |>
  group_by(run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(fork_length < 20 | fork_length > 200), .groups = "drop") |>
  mutate(data_type = "rst", issue_type = "implausible_value", field = "fork_length")

annual_catch <- rst_catch |>
  filter(!is.na(date), !is.na(count)) |>
  mutate(run_year = as_run_year(date)) |>
  group_by(stream, site, run_year) |>
  summarise(total_catch = sum(count, na.rm = TRUE), .groups = "drop")

catch_site_stats <- annual_catch |>
  group_by(stream, site) |>
  summarise(n_years = n(), median_catch = median(total_catch), .groups = "drop") |>
  filter(n_years >= 5, median_catch > 0)

annual_catch_eval <- annual_catch |> inner_join(catch_site_stats, by = c("stream", "site"))

rst_catch_high_metrics <- annual_catch_eval |>
  group_by(run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(total_catch > 4 * median_catch), .groups = "drop") |>
  mutate(data_type = "rst", issue_type = "extreme_value", field = "count")

efficiency_summary <- weekly_juvenile_abundance_efficiency_data |>
  filter(!is.na(number_recaptured), !is.na(number_released), number_released > 0) |>
  mutate(simple_eff = number_recaptured / number_released) |>
  group_by(stream, site) |>
  summarise(mean_simple_eff = mean(simple_eff, na.rm = TRUE), .groups = "drop") |>
  filter(mean_simple_eff > 0)

annual_expanded <- rst_catch |>
  filter(!is.na(date), !is.na(count)) |>
  mutate(run_year = as_run_year(date)) |>
  inner_join(efficiency_summary, by = c("stream", "site")) |>
  mutate(expanded_count = count / mean_simple_eff) |>
  group_by(stream, site, run_year) |>
  summarise(total_expanded = sum(expanded_count, na.rm = TRUE), .groups = "drop")

expanded_site_stats <- annual_expanded |>
  group_by(stream, site) |>
  summarise(n_years = n(), median_expanded = median(total_expanded), .groups = "drop") |>
  filter(n_years >= 5, median_expanded > 0)


weekly_catch_stats <- weekly_juvenile_abundance_catch_data |>
  filter(!is.na(count)) |>
  group_by(stream, site) |>
  mutate(
    mean_weekly_count = mean(count, na.rm = TRUE),
    sd_weekly_count   = sd(count, na.rm = TRUE),
    upper_limit       = mean_weekly_count + 2 * sd_weekly_count
  ) |>
  ungroup() |>
  filter(sd_weekly_count > 0)

rst_catch_weekly_metrics <- weekly_catch_stats |>
  group_by(run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(count > upper_limit), .groups = "drop") |>
  mutate(data_type = "rst", issue_type = "extreme_value", field = "count_weekly")

rst_sampling_metrics <- weekly_juvenile_abundance_catch_data |>
  filter(week %in% RST_WEEKS, !is.na(count)) |>
  group_by(stream, site, run_year) |>
  summarise(weeks_sampled = n_distinct(week), .groups = "drop") |>
  group_by(run_year) |>
  summarise(n_evaluated = n() * 31L, n_failed = sum(pmax(31L - weeks_sampled, 0L)), .groups = "drop") |>
  mutate(data_type = "rst", issue_type = "low_sampling_effort", field = "weeks_sampled")

rst_hours_metrics <- weekly_juvenile_abundance_catch_data |>
  filter(!is.na(hours_fished)) |>
  group_by(run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(hours_fished > 168), .groups = "drop") |>
  mutate(data_type = "rst", issue_type = "implausible_value", field = "hours_fished")
```

```{r}
#| label: flow-metrics
#| echo: false

EXPECTED_STREAMS <- c(
  "battle creek", "butte creek", "clear creek", "deer creek",
  "feather river", "mill creek", "sacramento river", "yuba river"
)

env <- environmental_data |> mutate(run_year = if_else(week >= 45L, year + 1L, year))
flow_data <- env |> filter(parameter == "flow", statistic == "mean")
temp_data <- env |> filter(parameter == "temperature", statistic == "mean")

flow_stream_coverage_metrics <- flow_data |>
  distinct(stream, run_year) |>
  group_by(run_year) |>
  summarise(n_present = n_distinct(stream), .groups = "drop") |>
  mutate(n_evaluated = length(EXPECTED_STREAMS), n_failed = pmax(n_evaluated - n_present, 0L)) |>
  select(run_year, n_evaluated, n_failed) |>
  mutate(data_type = "flow", issue_type = "gap", field = "stream_coverage")

flow_gap_metrics <- flow_data |>
  filter(week %in% RST_WEEKS) |>
  group_by(stream, site_group, run_year) |>
  summarise(n_present = n(), .groups = "drop") |>
  mutate(n_missing = pmax(31L - n_present, 0L)) |>
  group_by(run_year) |>
  summarise(n_evaluated = n() * 31L, n_failed = sum(n_missing), .groups = "drop") |>
  mutate(data_type = "flow", issue_type = "gap", field = "week")

flow_zero_metrics <- flow_data |>
  filter(week %in% RST_WEEKS) |>
  group_by(run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(value <= 0), .groups = "drop") |>
  mutate(data_type = "flow", issue_type = "implausible_value", field = "value")

flow_temp_metrics <- temp_data |>
  group_by(run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(value < -5 | value > 35), .groups = "drop") |>
  mutate(data_type = "flow", issue_type = "implausible_value", field = "temperature")
```

```{r}
#| label: efficiency-metrics
#| echo: false

eff <- weekly_efficiency |>
  mutate(run_year = if_else(week_released >= 45, year_released + 1L, year_released))

eff_recap_metrics <- eff |>
  group_by(run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(number_recaptured > number_released), .groups = "drop") |>
  mutate(data_type = "efficiency", issue_type = "recaptures_exceed_releases", field = "number_recaptured")

eff_low_release_metrics <- eff |>
  group_by(run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(number_released < 50), .groups = "drop") |>
  mutate(data_type = "efficiency", issue_type = "low_release_count", field = "number_released")

eff_zero_recap_metrics <- eff |>
  group_by(run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(number_recaptured == 0), .groups = "drop") |>
  mutate(data_type = "efficiency", issue_type = "zero_recaptures", field = "number_recaptured")

eff_trial_coverage_metrics <- weekly_juvenile_abundance_efficiency_data |>
  filter(week %in% RST_WEEKS) |>
  group_by(stream, site, run_year) |>
  summarise(n_trials = n(), .groups = "drop") |>
  group_by(run_year) |>
  summarise(n_evaluated = n() * 3L, n_failed = sum(pmax(3L - n_trials, 0L)), .groups = "drop") |>
  mutate(data_type = "efficiency", issue_type = "low_trial_coverage", field = "n_trials")

eff_rates <- eff |>
  filter(number_released > 0, number_recaptured <= number_released) |>
  mutate(simple_eff = number_recaptured / number_released) |>
  group_by(stream, site, run_year) |>
  summarise(mean_eff = mean(simple_eff, na.rm = TRUE), .groups = "drop")

eff_site_stats <- eff_rates |>
  group_by(stream, site) |>
  summarise(n_years = n(), hist_mean = mean(mean_eff, na.rm = TRUE), hist_sd = sd(mean_eff, na.rm = TRUE), .groups = "drop") |>
  filter(n_years >= 5)

eff_outlier_metrics <- eff_rates |>
  inner_join(eff_site_stats, by = c("stream", "site")) |>
  group_by(run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(mean_eff > hist_mean + 3 * hist_sd), .groups = "drop") |>
  mutate(data_type = "efficiency", issue_type = "extreme_value", field = "number_recaptured")

all_historical_run_years <- sort(unique(c(annual_catch$run_year, flow_data$run_year, eff$run_year)))

eff_no_trials_metrics <- map_dfr(all_historical_run_years, function(ry) {
  active <- rst_catch |>
    filter(!is.na(date), !is.na(count)) |>
    mutate(run_year = as_run_year(date)) |>
    filter(run_year == ry) |>
    distinct(stream, site)
  if (nrow(active) == 0) return(tibble())
  had_trials <- eff |> filter(run_year == ry) |> distinct(stream, site) |> mutate(has_trials = TRUE)
  active |>
    left_join(had_trials, by = c("stream", "site")) |>
    mutate(has_trials = replace_na(has_trials, FALSE)) |>
    summarise(n_evaluated = n(), n_failed = sum(!has_trials)) |>
    mutate(run_year = ry)
}) |>
  mutate(data_type = "efficiency", issue_type = "no_efficiency_trials", field = "number_released")
```

```{r}
#| label: combine-metrics
#| echo: false

metrics <- bind_rows(
  rst_fl_metrics, rst_fl_impl_metrics, rst_catch_high_metrics, rst_catch_weekly_metrics, rst_sampling_metrics, rst_hours_metrics,
  flow_stream_coverage_metrics, flow_gap_metrics, flow_zero_metrics, flow_temp_metrics,
  eff_recap_metrics, eff_low_release_metrics, eff_zero_recap_metrics, eff_trial_coverage_metrics,
  eff_outlier_metrics, eff_no_trials_metrics
) |>
  mutate(pct_failed = if_else(n_evaluated > 0, round(n_failed / n_evaluated * 100, 2), NA_real_)) |>
  left_join(CHECK_METADATA, by = c("data_type", "issue_type", "field"))

n_history_years <- metrics |> filter(run_year != review_run_year) |> distinct(run_year) |> nrow()
```

`r if (n_history_years == 0) "No prior run_years are available for comparison yet." else glue("Comparing against {n_history_years} prior run year(s).")`

---

## Headline: This Year vs. Historical Average, by Check

For each check, "this year" is the pooled `n_failed / n_evaluated` for `review_run_year`; "historical average" is the mean of that same pooled rate computed separately for each prior run_year, with `notably_worse` flagged when this year exceeds the historical mean by more than 2 historical SDs.

```{r}
#| label: headline-table

this_year <- metrics |>
  filter(run_year == review_run_year) |>
  select(data_type, issue_type, field, alert_level, severity, this_year_pct = pct_failed)

historical <- metrics |>
  filter(run_year != review_run_year) |>
  group_by(data_type, issue_type, field) |>
  summarise(
    n_hist_years = n(),
    hist_avg_pct = round(mean(pct_failed, na.rm = TRUE), 2),
    hist_sd_pct  = round(sd(pct_failed, na.rm = TRUE), 2),
    .groups      = "drop"
  )

headline <- this_year |>
  full_join(historical, by = c("data_type", "issue_type", "field")) |>
  mutate(
    delta_pct     = round(this_year_pct - hist_avg_pct, 2),
    notably_worse = !is.na(hist_avg_pct) & !is.na(hist_sd_pct) & hist_sd_pct > 0 &
                    !is.na(this_year_pct) & this_year_pct > hist_avg_pct + 2 * hist_sd_pct,
    check_label   = as.character(glue("{data_type} / {issue_type} / {field}"))
  ) |>
  arrange(desc(notably_worse), desc(delta_pct))

DT::datatable(
  headline |>
    select(check_label, alert_level, severity, this_year_pct, hist_avg_pct,
           n_hist_years, delta_pct, notably_worse),
  caption = glue("Check-level failure rates: run year {review_run_year} vs. historical average"),
  options = list(pageLength = 25, scrollX = TRUE),
  rownames = FALSE
) |>
  DT::formatStyle("notably_worse",
                  target = "row",
                  backgroundColor = DT::styleEqual(TRUE, "#fde0dc"))
```

---

## Visualization: This Year vs. Historical Average

One panel per `data_type`. Grey dot = historical average pooled failure rate; colored dot = this year (amber = warning-level check, red = error-level check). Bars show ±1 historical SD.

```{r}
#| label: plot-dumbbell
#| echo: false
#| fig-height: 10

plot_data <- headline |>
  filter(!is.na(this_year_pct) | !is.na(hist_avg_pct)) |>
  mutate(sort_key = pmax(this_year_pct, hist_avg_pct, na.rm = TRUE)) |>
  arrange(sort_key) |>
  mutate(check_label = factor(check_label, levels = unique(check_label)))

ggplot(plot_data) +
  geom_segment(aes(x = pmin(hist_avg_pct, this_year_pct, na.rm = TRUE),
                    xend = pmax(hist_avg_pct, this_year_pct, na.rm = TRUE),
                    y = check_label, yend = check_label),
               color = "grey70") +
  geom_errorbar(aes(xmin = hist_avg_pct - hist_sd_pct, xmax = hist_avg_pct + hist_sd_pct, y = check_label),
                orientation = "y", width = 0.15, color = "grey60", na.rm = TRUE) +
  geom_point(aes(x = hist_avg_pct, y = check_label), color = "grey40", size = 2, na.rm = TRUE) +
  geom_point(aes(x = this_year_pct, y = check_label, color = alert_level), size = 3, na.rm = TRUE) +
  scale_color_manual(values = c("warning" = "#fdae61", "error" = "#d73027"), name = "This year (by alert level)") +
  facet_wrap(~data_type, scales = "free_y", ncol = 1) +
  labs(title = glue("Check failure rate: run year {review_run_year} (colored) vs. historical average (grey)"),
       x = "% of records failed", y = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

---

## Open Issue Counts: This Year vs. Historical Average

```{r}
#| label: open-issue-counts
#| echo: false

full_log <- readr::read_csv(QC_LOG_PATH, show_col_types = FALSE) |>
  mutate(run_year = as.integer(run_year))

issues_per_year <- full_log |>
  count(data_type, alert_level, severity, run_year, name = "n_issues")

this_year_issues <- issues_per_year |>
  filter(run_year == review_run_year) |>
  select(data_type, alert_level, severity, this_year_n_issues = n_issues)

hist_issues <- issues_per_year |>
  filter(run_year != review_run_year) |>
  group_by(data_type, alert_level, severity) |>
  summarise(hist_avg_n_issues = round(mean(n_issues), 1), n_hist_years = n(), .groups = "drop")

issue_count_summary <- this_year_issues |>
  full_join(hist_issues, by = c("data_type", "alert_level", "severity")) |>
  arrange(data_type, alert_level, severity)

DT::datatable(
  issue_count_summary,
  caption = glue("New issues logged per run: run year {review_run_year} vs. historical average"),
  options = list(pageLength = 25, scrollX = TRUE),
  rownames = FALSE
)
```

### Checks Flagged as Notably Worse Than Historical Average

```{r}
#| label: notably-worse-table
#| echo: false

worse <- headline |> filter(notably_worse)

if (nrow(worse) == 0) {
  cat("No checks are notably worse than their historical average this run.")
} else {
  DT::datatable(
    worse |> select(check_label, alert_level, severity, this_year_pct, hist_avg_pct, hist_sd_pct, delta_pct),
    caption = glue("{nrow(worse)} check(s) notably worse than historical average (>2 SD)"),
    options = list(pageLength = 25, scrollX = TRUE),
    rownames = FALSE
  )
}
```

---

## All Open Issues, by Alert Level and Severity

```{r}
#| label: open-issues-table
#| echo: false

open_issues <- full_log |> filter(status == "open")

if (nrow(open_issues) == 0) {
  cat("No open issues in the log.")
} else {
  DT::datatable(
    open_issues |> count(data_type, alert_level, severity, name = "n_open") |>
      arrange(data_type, alert_level, severity),
    caption = glue("{nrow(open_issues)} total open issue(s) across all data types"),
    options = list(pageLength = 25, scrollX = TRUE),
    rownames = FALSE
  )
}
```
