---
title: "QC Summary: This Year vs. Historical Average"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    embed-resources: true
params:
  review_run_year: 2026
execute:
  warning: false
  message: false
---

```{r}
#| label: setup

library(tidyverse)
library(lubridate)
library(here)
library(glue)
library(DT)
library(plotly)
library(kableExtra)

if (file.exists(here::here("DESCRIPTION"))) {
  devtools::load_all(here::here(), quiet = TRUE)
} else {
  library(SRJPEdata)
}

source(here::here("data-raw", "qc", "qc_helpers.R"))

review_run_year <- params$review_run_year

RST_WEEKS <- c(45:53, 1:22)

# Earliest run_year shown in the trend plot and heatmaps below. Flow data goes
# back to 1961, but the pre-1990 span is sparse enough (few streams, few
# checks with any evaluable data) that including it mostly just compresses
# the years reviewers actually care about. The headline table is unaffected
# and still uses the full history.
MIN_TREND_YEAR <- 1990

# Rolls up per-(stream, run_year) check results into the pooled per-run_year
# metrics used by the headline table and trend plot. Kept as a helper since
# every check below is computed at the stream grain first (for the
# per-stream heatmap tabs) and then pooled the same way.
pool_across_streams <- function(by_stream) {
  by_stream |>
    dplyr::group_by(run_year, data_type, issue_type, field) |>
    dplyr::summarise(n_evaluated = sum(n_evaluated), n_failed = sum(n_failed), .groups = "drop")
}
```

## Overview

This report compares **run year `r review_run_year`** against the historical average across all other run years present in the underlying data. 

```{r}
#| label: rst-metrics
#| echo: false

rst_fl_metrics_by_stream <- rst_catch |>
  filter(!is.na(date)) |>
  mutate(run_year = as_run_year(date)) |>
  group_by(stream, run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(is.na(fork_length)), .groups = "drop") |>
  mutate(data_type = "rst", issue_type = "high_na_rate", field = "fork_length")

rst_fl_impl_metrics_by_stream <- rst_catch |>
  filter(!is.na(fork_length), fork_length > 0) |>
  mutate(run_year = as_run_year(date)) |>
  group_by(stream, run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(fork_length < 20 | fork_length > 200), .groups = "drop") |>
  mutate(data_type = "rst", issue_type = "implausible_value", field = "fork_length")

annual_catch <- rst_catch |>
  filter(!is.na(date), !is.na(count)) |>
  mutate(run_year = as_run_year(date)) |>
  group_by(stream, site, run_year) |>
  summarise(total_catch = sum(count, na.rm = TRUE), .groups = "drop")

catch_site_stats <- annual_catch |>
  group_by(stream, site) |>
  summarise(n_years = n(), median_catch = median(total_catch), .groups = "drop") |>
  filter(n_years >= 5, median_catch > 0)

annual_catch_eval <- annual_catch |> inner_join(catch_site_stats, by = c("stream", "site"))

rst_catch_high_metrics_by_stream <- annual_catch_eval |>
  group_by(stream, run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(total_catch > 4 * median_catch), .groups = "drop") |>
  mutate(data_type = "rst", issue_type = "extreme_value", field = "count")

efficiency_summary <- weekly_juvenile_abundance_efficiency_data |>
  filter(!is.na(number_recaptured), !is.na(number_released), number_released > 0) |>
  mutate(simple_eff = number_recaptured / number_released) |>
  group_by(stream, site) |>
  summarise(mean_simple_eff = mean(simple_eff, na.rm = TRUE), .groups = "drop") |>
  filter(mean_simple_eff > 0)

annual_expanded <- rst_catch |>
  filter(!is.na(date), !is.na(count)) |>
  mutate(run_year = as_run_year(date)) |>
  inner_join(efficiency_summary, by = c("stream", "site")) |>
  mutate(expanded_count = count / mean_simple_eff) |>
  group_by(stream, site, run_year) |>
  summarise(total_expanded = sum(expanded_count, na.rm = TRUE), .groups = "drop")

expanded_site_stats <- annual_expanded |>
  group_by(stream, site) |>
  summarise(n_years = n(), median_expanded = median(total_expanded), .groups = "drop") |>
  filter(n_years >= 5, median_expanded > 0)


weekly_catch_stats <- weekly_juvenile_abundance_catch_data |>
  filter(!is.na(count)) |>
  group_by(stream, site) |>
  mutate(
    mean_weekly_count = mean(count, na.rm = TRUE),
    sd_weekly_count   = sd(count, na.rm = TRUE),
    upper_limit       = mean_weekly_count + 2 * sd_weekly_count
  ) |>
  ungroup() |>
  filter(sd_weekly_count > 0)

rst_catch_weekly_metrics_by_stream <- weekly_catch_stats |>
  group_by(stream, run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(count > upper_limit), .groups = "drop") |>
  mutate(data_type = "rst", issue_type = "extreme_value", field = "count_weekly")

rst_sampling_metrics_by_stream <- weekly_juvenile_abundance_catch_data |>
  filter(week %in% RST_WEEKS, !is.na(count)) |>
  group_by(stream, site, run_year) |>
  summarise(weeks_sampled = n_distinct(week), .groups = "drop") |>
  group_by(stream, run_year) |>
  summarise(n_evaluated = n() * 31L, n_failed = sum(pmax(31L - weeks_sampled, 0L)), .groups = "drop") |>
  mutate(data_type = "rst", issue_type = "low_sampling_effort", field = "weeks_sampled")

```

```{r}
#| label: flow-metrics
#| echo: false

EXPECTED_STREAMS <- c(
  "battle creek", "butte creek", "clear creek", "deer creek",
  "feather river", "mill creek", "sacramento river", "yuba river"
)
environmental_data <- bind_rows(flow_data, temperature_data)
env <- environmental_data |> mutate(run_year = if_else(week >= 45L, year + 1L, year))
flow_data <- env |> filter(parameter == "flow", statistic == "mean")
temp_data <- env |> filter(parameter == "temperature", statistic == "mean")

# Special-cased rather than a simple group_by(stream, ...) add-on, since this
# check is inherently about whether each stream shows up at all in a given
# run_year - there's no site-level detail to roll up from. Every expected
# stream x run_year combination is scored, so a stream missing entirely from
# flow_data that year is a genuine (not implicit) failure at the stream grain.
flow_stream_coverage_by_stream <- tidyr::expand_grid(
    stream = EXPECTED_STREAMS,
    run_year = sort(unique(flow_data$run_year))
  ) |>
  left_join(flow_data |> distinct(stream, run_year) |> mutate(present = TRUE),
            by = c("stream", "run_year")) |>
  mutate(n_evaluated = 1L, n_failed = if_else(is.na(present), 1L, 0L),
         data_type = "flow", issue_type = "gap", field = "stream_coverage") |>
  select(stream, run_year, n_evaluated, n_failed, data_type, issue_type, field)

flow_gap_metrics_by_stream <- flow_data |>
  filter(week %in% RST_WEEKS) |>
  group_by(stream, site_group, run_year) |>
  summarise(n_present = n(), .groups = "drop") |>
  mutate(n_missing = pmax(31L - n_present, 0L)) |>
  group_by(stream, run_year) |>
  summarise(n_evaluated = n() * 31L, n_failed = sum(n_missing), .groups = "drop") |>
  mutate(data_type = "flow", issue_type = "gap", field = "week")

flow_zero_metrics_by_stream <- flow_data |>
  filter(week %in% RST_WEEKS) |>
  group_by(stream, run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(value <= 0), .groups = "drop") |>
  mutate(data_type = "flow", issue_type = "implausible_value", field = "value")

flow_temp_metrics_by_stream <- temp_data |>
  group_by(stream, run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(value < -5 | value > 35), .groups = "drop") |>
  mutate(data_type = "flow", issue_type = "implausible_value", field = "temperature")
```

```{r}
#| label: efficiency-metrics
#| echo: false

eff <- weekly_efficiency |>
  mutate(run_year = if_else(week_released >= 45, year_released + 1L, year_released))

eff_recap_metrics_by_stream <- eff |>
  group_by(stream, run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(number_recaptured > number_released), .groups = "drop") |>
  mutate(data_type = "efficiency", issue_type = "recaptures_exceed_releases", field = "number_recaptured")

eff_low_release_metrics_by_stream <- eff |>
  group_by(stream, run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(number_released < 50), .groups = "drop") |>
  mutate(data_type = "efficiency", issue_type = "low_release_count", field = "number_released")

eff_zero_recap_metrics_by_stream <- eff |>
  group_by(stream, run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(number_recaptured == 0), .groups = "drop") |>
  mutate(data_type = "efficiency", issue_type = "zero_recaptures", field = "number_recaptured")

eff_trial_coverage_metrics_by_stream <- weekly_juvenile_abundance_efficiency_data |>
  filter(week %in% RST_WEEKS) |>
  group_by(stream, site, run_year) |>
  summarise(n_trials = n(), .groups = "drop") |>
  group_by(stream, run_year) |>
  summarise(n_evaluated = n() * 3L, n_failed = sum(pmax(3L - n_trials, 0L)), .groups = "drop") |>
  mutate(data_type = "efficiency", issue_type = "low_trial_coverage", field = "n_trials")

eff_rates <- eff |>
  filter(number_released > 0, number_recaptured <= number_released) |>
  mutate(simple_eff = number_recaptured / number_released) |>
  group_by(stream, site, run_year) |>
  summarise(mean_eff = mean(simple_eff, na.rm = TRUE), .groups = "drop")

eff_site_stats <- eff_rates |>
  group_by(stream, site) |>
  summarise(n_years = n(), hist_mean = mean(mean_eff, na.rm = TRUE), hist_sd = sd(mean_eff, na.rm = TRUE), .groups = "drop") |>
  filter(n_years >= 5)

eff_outlier_metrics_by_stream <- eff_rates |>
  inner_join(eff_site_stats, by = c("stream", "site")) |>
  group_by(stream, run_year) |>
  summarise(n_evaluated = n(), n_failed = sum(mean_eff > hist_mean + 3 * hist_sd), .groups = "drop") |>
  mutate(data_type = "efficiency", issue_type = "extreme_value", field = "number_recaptured")

all_historical_run_years <- sort(unique(c(annual_catch$run_year, flow_data$run_year, eff$run_year)))

eff_no_trials_metrics_by_stream <- map_dfr(all_historical_run_years, function(ry) {
  active <- rst_catch |>
    filter(!is.na(date), !is.na(count)) |>
    mutate(run_year = as_run_year(date)) |>
    filter(run_year == ry) |>
    distinct(stream, site)
  if (nrow(active) == 0) return(tibble())
  had_trials <- eff |> filter(run_year == ry) |> distinct(stream, site) |> mutate(has_trials = TRUE)
  active |>
    left_join(had_trials, by = c("stream", "site")) |>
    mutate(has_trials = replace_na(has_trials, FALSE)) |>
    group_by(stream) |>
    summarise(n_evaluated = n(), n_failed = sum(!has_trials), .groups = "drop") |>
    mutate(run_year = ry)
}) |>
  mutate(data_type = "efficiency", issue_type = "no_efficiency_trials", field = "number_released")
```

```{r}
#| label: combine-metrics
#| echo: false

# Every check above is computed at the (stream, run_year) grain first. metrics
# (pooled across streams, as before) and metrics_by_stream (used by the
# per-stream heatmap tabs) are both derived from that same by-stream data, so
# the two views can never drift apart from each other.
metrics_by_stream_raw <- bind_rows(
  rst_fl_metrics_by_stream, rst_fl_impl_metrics_by_stream, rst_catch_high_metrics_by_stream,
  rst_catch_weekly_metrics_by_stream, rst_sampling_metrics_by_stream,
  flow_stream_coverage_by_stream, flow_gap_metrics_by_stream, flow_zero_metrics_by_stream, flow_temp_metrics_by_stream,
  eff_recap_metrics_by_stream, eff_low_release_metrics_by_stream, eff_zero_recap_metrics_by_stream,
  eff_trial_coverage_metrics_by_stream, eff_outlier_metrics_by_stream, eff_no_trials_metrics_by_stream
)

metrics_by_stream <- metrics_by_stream_raw |>
  mutate(pct_failed = if_else(n_evaluated > 0, round(n_failed / n_evaluated * 100, 2), NA_real_)) |>
  left_join(CHECK_METADATA, by = c("data_type", "issue_type", "field"))

metrics <- pool_across_streams(metrics_by_stream_raw) |>
  mutate(pct_failed = if_else(n_evaluated > 0, round(n_failed / n_evaluated * 100, 2), NA_real_)) |>
  left_join(CHECK_METADATA, by = c("data_type", "issue_type", "field"))

n_history_years <- metrics |> filter(run_year != review_run_year) |> distinct(run_year) |> nrow()
```

---

## Failure Rate Heatmap (All Checks x Run Years)

Every check (row) against every run year (column) from `r MIN_TREND_YEAR` onward, colored by % failed, split into an **Error** panel and a **Warning** panel by `alert_level`. The outlined column is `review_run_year`\. This plot provides a quick overview of how checks preform over time. Hover any tile for the exact numbers and a description of what the check measures.

The **Overall** tab pools every stream together, matching the tables and plots below. Each stream tab recomputes the same checks from just that stream's records, so you can tell whether something flagged in the pooled view is basin-wide or concentrated at one site.

```{r}
#| label: heatmap-helpers
#| echo: false

# Sequential ramps following this report's error = red / warning = amber
# color convention, each stretched into its own light->dark single-hue scale
# rather than used as a flat swatch - a heatmap needs a gradient, not a
# single color, to also show magnitude within each alert_level.
ERROR_RAMP   <- c(low = "#fff5f0", high = "#a50026")
WARNING_RAMP <- c(low = "#fffde7", high = "#f9a825")

# One alert_level's panel of a failure-rate heatmap: checks (rows) x run
# years (columns), colored on its own ramp, converted to plotly so each tile
# carries a hover tooltip (same info as the trend plot's tooltip below).
# Returns NULL if this alert_level has no rows in `data` (e.g. a single
# stream with only warning-level checks in its history), so
# stack_alert_level_panels() can skip it cleanly.
alert_level_heatmap_panel <- function(data, level, ramp) {
  d <- data |> filter(alert_level == level)
  if (nrow(d) == 0) return(NULL)

  d <- d |>
    mutate(
      check_label = factor(check_label, levels = rev(unique(check_label))),
      tooltip = if_else(
        is.na(pct_failed),
        as.character(glue(
          "{data_type} / {issue_type} / {field}<br>",
          "Run year: {run_year}<br>",
          "No evaluable data this year<br>",
          "{description}"
        )),
        as.character(glue(
          "{data_type} / {issue_type} / {field}<br>",
          "Run year: {run_year}<br>",
          "% failed: {pct_failed}% ({n_failed} / {n_evaluated})<br>",
          "Alert level: {alert_level} | Severity: {severity}<br>",
          "{description}"
        ))
      )
    )

  p <- ggplot(d, aes(x = factor(run_year), y = check_label, fill = pct_failed)) +
    geom_tile(aes(text = tooltip), color = "white", linewidth = 0.3) +
    geom_tile(data = filter(d, run_year == review_run_year), aes(text = tooltip),
              fill = NA, color = "black", linewidth = 0.9) +
    scale_fill_gradient(low = ramp[["low"]], high = ramp[["high"]],
                        na.value = "grey90", limits = c(0, 100)) +
    labs(title = str_to_title(level), x = "Run year", y = NULL) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          axis.text.y = element_text(size = 8),
          legend.position = "none")

  plotly::ggplotly(p, tooltip = "text")
}

# Stacks the error and warning panels, each on its own color scale (so a
# single facet_wrap() with one shared scale won't do here) and each already
# converted to plotly - combined with plotly::subplot() so the hover
# tooltip carries through (a static-only combiner like patchwork can't take
# plotly widgets as input). Panel heights stay proportional to how many
# checks each has: a 3-check panel and a 10-check panel would otherwise
# render at the same height and every tile in the smaller panel would look
# artificially tall.
stack_alert_level_panels <- function(data, title, height = 900, width = 1200) {
  n_error   <- data |> filter(alert_level == "error")   |> distinct(check_label) |> nrow()
  n_warning <- data |> filter(alert_level == "warning") |> distinct(check_label) |> nrow()

  p_error   <- alert_level_heatmap_panel(data, "error", ERROR_RAMP)
  p_warning <- alert_level_heatmap_panel(data, "warning", WARNING_RAMP)

  combined <- if (is.null(p_error)) {
    p_warning
  } else if (is.null(p_warning)) {
    p_error
  } else {
    plotly::subplot(p_error, p_warning, nrows = 2, shareX = TRUE,
                    titleX = TRUE, titleY = TRUE, margin = 0.04,
                    heights = c(n_error, n_warning) / (n_error + n_warning))
  }

  combined <- combined |> plotly::layout(title = title)
  # height/width must be set as top-level widget attributes, not via
  # layout() (that form is deprecated and silently no-ops), and subplot()
  # itself doesn't take sizing args - so they're assigned directly here.
  combined$width  <- width
  combined$height <- height
  combined
}

# check_label's alert_level/data_type/severity/description never vary within
# a check, so nesting() keeps them paired to check_label while still
# crossing every check against the full set of run_years - a plain
# complete(check_label, run_year) would null out alert_level (and the rest)
# on the synthetic gap rows it creates, silently dropping them from both
# panels above instead of showing as grey.
build_heatmap_data <- function(data) {
  data |>
    filter(run_year >= MIN_TREND_YEAR) |>
    mutate(check_label = as.character(glue("{data_type} / {issue_type} / {field}"))) |>
    select(check_label, data_type, issue_type, field, alert_level, severity, description,
           run_year, n_evaluated, n_failed, pct_failed) |>
    complete(nesting(check_label, data_type, issue_type, field, alert_level, severity, description), run_year) |>
    arrange(data_type)
}
```

::: {.panel-tabset}

### Overall

```{r}
#| label: plot-heatmap-overall
#| echo: false

heatmap_data <- build_heatmap_data(metrics)

stack_alert_level_panels(heatmap_data,
  title = glue("% failed by check across all run years (outlined column = run year {review_run_year})"))
```

```{r}
#| label: plot-heatmap-by-stream
#| echo: false
#| results: asis

for (s in sort(unique(metrics_by_stream$stream))) {
  cat(glue("\n\n### {str_to_title(s)}\n\n"))

  stream_heatmap_data <- build_heatmap_data(metrics_by_stream |> filter(stream == s))

  p_stream <- stack_alert_level_panels(stream_heatmap_data,
    title = glue("{str_to_title(s)}: % failed by check across all run years"),
    height = 700)

  # plain print() doesn't reliably emit an htmlwidget from inside a for loop
  # under results: asis - wrapping in tagList() is the standard fix.
  print(htmltools::tagList(p_stream))

  cat("\n\n")
}
```

:::

---

## Table Showing this Year vs. Historical Average

For each check, "this year" is the pooled `n_failed / n_evaluated` for `review_run_year`; "historical average" is the mean of that same pooled rate computed separately for each prior run_year, with `notably_worse` flagged when this year exceeds the historical mean by more than 2 historical SDs.

```{r}
#| label: headline-table

this_year <- metrics |>
  filter(run_year == review_run_year) |>
  select(data_type, issue_type, field, alert_level, severity, this_year_pct = pct_failed)

historical <- metrics |>
  filter(run_year != review_run_year) |>
  group_by(data_type, issue_type, field) |>
  summarise(
    n_hist_years = n(),
    hist_avg_pct = round(mean(pct_failed, na.rm = TRUE), 2),
    hist_sd_pct  = round(sd(pct_failed, na.rm = TRUE), 2),
    .groups      = "drop"
  )

headline <- this_year |>
  full_join(historical, by = c("data_type", "issue_type", "field")) |>
  mutate(
    delta_pct     = round(this_year_pct - hist_avg_pct, 2),
    notably_worse = !is.na(hist_avg_pct) & !is.na(hist_sd_pct) & hist_sd_pct > 0 &
                    !is.na(this_year_pct) & this_year_pct > hist_avg_pct + 2 * hist_sd_pct,
    check_label   = as.character(glue("{data_type} / {issue_type} / {field}"))
  ) |>
  arrange(desc(notably_worse), desc(delta_pct))

DT::datatable(
  headline |>
    select(check_label, alert_level, severity, this_year_pct, hist_avg_pct,
           n_hist_years, delta_pct, notably_worse),
  caption = glue("Check-level failure rates: run year {review_run_year} vs. historical average"),
  options = list(pageLength = 25, scrollX = TRUE),
  rownames = FALSE
) |>
  DT::formatStyle("notably_worse",
                  target = "row",
                  backgroundColor = DT::styleEqual(TRUE, "#fde0dc"))
```

---

## Trend Across All Run Years

One panel per `data_type`, one colored line per check, from `r MIN_TREND_YEAR` onward. This shows whether `review_run_year` continues a multi-year trend or is an isolated blip — something the headline table above and the heatmap below can't convey as directly, since the table collapses all prior years into a single average and the heatmap needs a glance-then-scan rather than a continuous line to read a trend. The ringed point marks `review_run_year` on each line. Hover any point for the exact numbers and a description of what the check measures.

```{r}
#| label: plot-trend
#| echo: false
#| fig-height: 16
#| fig-width: 10

trend_data <- metrics |>
  filter(!is.na(pct_failed), run_year >= MIN_TREND_YEAR) |>
  mutate(
    check_short = as.character(glue("{issue_type} / {field}")),
    tooltip = as.character(glue(
      "{data_type} / {issue_type} / {field}<br>",
      "Run year: {run_year}<br>",
      "% failed: {pct_failed}% ({n_failed} / {n_evaluated})<br>",
      "Alert level: {alert_level} | Severity: {severity}<br>",
      "{description}"
    ))
  )

p_trend <- ggplot(trend_data, aes(x = run_year, y = pct_failed, color = check_short, group = check_short)) +
  geom_line(alpha = 0.7) +
  geom_point(aes(text = tooltip), size = 1.8, alpha = 0.8) +
  geom_point(data = filter(trend_data, run_year == review_run_year),
             aes(text = tooltip),
             shape = 21, size = 4, stroke = 1.2, color = "black", fill = NA) +
  geom_vline(xintercept = review_run_year, linetype = "dashed", color = "grey50") +
  facet_wrap(~data_type, scales = "free_y", ncol = 1) +
  labs(title = glue("% failed by run year, per check (ringed point = run year {review_run_year})"),
       x = "Run year", y = "% of records failed") +
  theme_minimal()

# theme(legend.position = "none") on the ggplot above is not enough here -
# ggplotly() ignores it for a discrete color scale and marks every line
# trace showlegend = TRUE regardless, so it has to be forced off on the
# plotly side directly.
plotly::ggplotly(p_trend, tooltip = "text", height = 1100, width = 800) |>
  plotly::layout(showlegend = FALSE)
```

---

## Open Issue Counts - This Year vs. Historical Average

```{r}
#| label: open-issue-counts
#| echo: false

full_log <- read_full_qc_log() |>
  mutate(run_year = as.integer(run_year))

issues_per_year <- full_log |>
  count(data_type, alert_level, severity, run_year, name = "n_issues")

this_year_issues <- issues_per_year |>
  filter(run_year == review_run_year) |>
  select(data_type, alert_level, severity, this_year_n_issues = n_issues)

hist_issues <- issues_per_year |>
  filter(run_year != review_run_year) |>
  group_by(data_type, alert_level, severity) |>
  summarise(hist_avg_n_issues = round(mean(n_issues), 1), n_hist_years = n(), .groups = "drop")

issue_count_summary <- this_year_issues |>
  full_join(hist_issues, by = c("data_type", "alert_level", "severity")) |>
  arrange(data_type, alert_level, severity)

DT::datatable(
  issue_count_summary,
  caption = glue("New issues logged per run: run year {review_run_year} vs. historical average"),
  options = list(pageLength = 25, scrollX = TRUE),
  rownames = FALSE
)
```

### Checks Flagged as Notably Worse Than Historical Average

```{r}
#| label: notably-worse-table
#| echo: false

worse <- headline |> filter(notably_worse)

if (nrow(worse) == 0) {
  cat("No checks are notably worse than their historical average this run.")
} else {
  DT::datatable(
    worse |> select(check_label, alert_level, severity, this_year_pct, hist_avg_pct, hist_sd_pct, delta_pct),
    caption = glue("{nrow(worse)} check(s) notably worse than historical average (>2 SD)"),
    options = list(pageLength = 25, scrollX = TRUE),
    rownames = FALSE
  )
}
```

---

## All Open Issues

```{r}
#| label: open-issues-table
#| echo: false

open_issues <- full_log |> filter(status == "open")

if (nrow(open_issues) == 0) {
  cat("No open issues in the log.")
} else {
  DT::datatable(
    open_issues |> count(data_type, alert_level, severity, name = "n_open") |>
      arrange(data_type, alert_level, severity),
    caption = glue("{nrow(open_issues)} total open issue(s) across all data types"),
    options = list(pageLength = 25, scrollX = TRUE),
    rownames = FALSE
  )
}
```

---

## Error Log - Run Year `r review_run_year`

The individual error-level entries from `qc_errors.csv` for `review_run_year`, for reference alongside the checks above — everything else in this report is a computed rate, this is the actual log rows behind it. Rows still in `open` status are highlighted.

```{r}
#| label: error-log-table
#| echo: false

year_errors <- full_log |>
  filter(alert_level == "error", run_year == review_run_year) |>
  select(data_type, stream, site, issue_type, field, description, n_records, severity, status, reviewer_notes) |>
  arrange(data_type, stream, site, issue_type)

if (nrow(year_errors) == 0) {
  cat(glue("No error-level issues logged for run year {review_run_year}."))
} else {
  year_errors |>
    kableExtra::kbl(
      caption = glue("Error-level issues logged for run year {review_run_year} ({nrow(year_errors)} total)"),
      col.names = c("Data Type", "Stream", "Site", "Issue Type", "Field", "Description",
                    "# Records", "Severity", "Status", "Reviewer Notes"),
      row.names = FALSE
    ) |>
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) |>
    kableExtra::row_spec(which(year_errors$status == "open"), background = "#fde0dc")
}
```
