---
title: "RST Data Annual QC"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    embed-resources: true
params:
  review_run_year: 2026
---

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(lubridate)
library(here)
library(glue)
library(DT)

# Load package data (works from source or installed)
if (file.exists(here::here("DESCRIPTION"))) {
  devtools::load_all(here::here(), quiet = TRUE)
} else {
  library(SRJPEdata)
}

source(here::here("data-raw", "qc", "qc_helpers.R"))

review_run_year <- params$review_run_year
```

## Overview

Reviewing RST catch and trap data. Run year being reviewed: **`r review_run_year`**.

Data sources: `rst_catch`, `rst_trap`, `weekly_juvenile_abundance_catch_data`, `years_to_exclude_rst_data`.

```{r}
#| label: overview-table
#| echo: false

rst_catch |>
  mutate(run_year = as_run_year(date)) |>
  group_by(stream, site) |>
  summarise(
    min_run_year = min(run_year, na.rm = TRUE),
    max_run_year = max(run_year, na.rm = TRUE),
    total_records = n(),
    .groups = "drop"
  ) |>
  arrange(stream, site) |>
  knitr::kable(caption = "RST catch records by stream and site")
```

---

## Check 1: Fork Length — High NA Rate

Flag stream/site/run_years where >95% of records have no fork length recorded and total records exceed 50. Indicates a season where measurements were not taken.

```{r}
#| label: check-fl-na

fl_na_check <- rst_catch |>
  filter(!is.na(date)) |>
  mutate(run_year = as_run_year(date)) |>
  group_by(stream, site, run_year) |>
  summarise(
    n_records    = n(),
    n_na_fl      = sum(is.na(fork_length)),
    pct_na_fl    = round(mean(is.na(fork_length)) * 100, 1),
    .groups      = "drop"
  ) |>
  filter(pct_na_fl > 95, n_records > 50)

new_fl_na <- log_issues(
  fl_na_check |>
    mutate(
      data_type   = "rst",
      issue_type  = "high_na_rate",
      field       = "fork_length",
      severity    = "moderate",
      description = glue("{n_na_fl} of {n_records} records ({pct_na_fl}%) have no fork length in run year {run_year} at {site} ({stream})")
    )
)
```

```{r}
#| label: plot-fl-na
#| echo: false
#| fig-height: 6

fl_na_all <- rst_catch |>
  filter(!is.na(date)) |>
  mutate(run_year = as_run_year(date)) |>
  group_by(stream, site, run_year) |>
  summarise(pct_na_fl = round(mean(is.na(fork_length)) * 100, 1), .groups = "drop")

ggplot(fl_na_all, aes(x = run_year, y = site, fill = pct_na_fl)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "#d73027", name = "% NA") +
  facet_wrap(~stream, scales = "free_y", ncol = 2) +
  labs(title = "Fork length NA rate by site and run year",
       x = "Run year", y = NULL) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

`r if (nrow(fl_na_check) == 0) "No issues detected." else glue("{nrow(fl_na_check)} site/run-year combination(s) flagged.")`

---

## Check 2: Fork Length — Implausible Values

Flag records where `fork_length < 20mm` or `fork_length > 200mm`. These are biologically implausible for juvenile chinook at an RST.

```{r}
#| label: check-fl-implausible

fl_impl_check <- rst_catch |>
  filter(!is.na(fork_length), fork_length > 0) |>
  mutate(run_year = as_run_year(date)) |>
  mutate(implausible = fork_length < 20 | fork_length > 200) |>
  group_by(stream, site, run_year) |>
  summarise(
    n_records    = sum(implausible),
    min_fl       = min(fork_length[implausible], na.rm = TRUE),
    max_fl       = max(fork_length[implausible], na.rm = TRUE),
    .groups      = "drop"
  ) |>
  filter(n_records > 0)

new_fl_impl <- log_issues(
  fl_impl_check |>
    mutate(
      data_type   = "rst",
      issue_type  = "implausible_value",
      field       = "fork_length",
      severity    = "critical",
      description = glue("{n_records} record(s) with fork_length <20mm or >200mm in run year {run_year} at {site} ({stream}). Range of flagged values: [{min_fl}, {max_fl}]mm")
    )
)
```

```{r}
#| label: plot-fl-implausible
#| echo: false
#| fig-height: 5

fl_plot_data <- rst_catch |>
  filter(!is.na(fork_length), fork_length > 0) |>
  mutate(
    run_year    = as_run_year(date),
    implausible = fork_length < 20 | fork_length > 200
  )

ggplot(fl_plot_data, aes(x = fork_length)) +
  geom_histogram(aes(fill = implausible), bins = 80, alpha = 0.8) +
  scale_fill_manual(values = c("FALSE" = "steelblue", "TRUE" = "#d73027"),
                    labels = c("Normal", "Flagged"), name = NULL) +
  facet_wrap(~stream, scales = "free_y", ncol = 2) +
  labs(title = "Fork length distribution with flagged values",
       x = "Fork length (mm)", y = "Count") +
  theme_minimal()
```

`r if (nrow(fl_impl_check) == 0) "No issues detected." else glue("{nrow(fl_impl_check)} site/run-year combination(s) flagged.")`

---

## Check 3: Catch — Extreme Annual Totals

Flag run_years where total catch at a site exceeds 4× the site's historical median. Requires at least 5 years of history. Not necessarily an error — warrants review.

```{r}
#| label: check-catch-extreme

annual_catch <- rst_catch |>
  filter(!is.na(date), !is.na(count)) |>
  mutate(run_year = as_run_year(date)) |>
  group_by(stream, site, run_year) |>
  summarise(total_catch = sum(count, na.rm = TRUE), .groups = "drop")

site_stats <- annual_catch |>
  group_by(stream, site) |>
  summarise(
    n_years      = n(),
    median_catch = median(total_catch),
    .groups      = "drop"
  ) |>
  filter(n_years >= 5, median_catch > 0)

catch_extreme <- annual_catch |>
  inner_join(site_stats, by = c("stream", "site")) |>
  filter(total_catch > 4 * median_catch)

new_catch_extreme <- log_issues(
  catch_extreme |>
    mutate(
      n_records   = 1L,
      data_type   = "rst",
      issue_type  = "extreme_value",
      field       = "count",
      severity    = "minor",
      description = glue("Total catch of {total_catch} in run year {run_year} at {site} ({stream}) is {round(total_catch / median_catch, 1)}x the site median ({median_catch})")
    )
)
```

```{r}
#| label: plot-catch-timeseries
#| echo: false
#| fig-height: 8

flagged_years <- catch_extreme |> select(stream, site, run_year)

ggplot(annual_catch, aes(x = run_year, y = total_catch)) +
  geom_line(color = "grey60") +
  geom_point(aes(color = paste(stream, site, run_year) %in%
               paste(flagged_years$stream, flagged_years$site, flagged_years$run_year)),
             size = 2) +
  scale_color_manual(values = c("FALSE" = "steelblue", "TRUE" = "#d73027"),
                     labels = c("Normal", "Flagged"), name = NULL) +
  facet_wrap(~stream + site, scales = "free_y", ncol = 2) +
  labs(title = "Annual total catch by site with flagged years",
       x = "Run year", y = "Total catch") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

`r if (nrow(catch_extreme) == 0) "No issues detected." else glue("{nrow(catch_extreme)} site/run-year combination(s) flagged.")`

---

## Check 4: Season Coverage — Low Sampling Effort

Flag all stream/site/run_years with fewer than 3 weeks sampled in the BTSPAS window (weeks 45–53 and 1–22), including those already in `years_to_exclude_rst_data`. The description notes exclusion status.

```{r}
#| label: check-season-coverage

weeks_in_window <- weekly_juvenile_abundance_catch_data |>
  filter(week %in% c(45:53, 1:22), !is.na(count)) |>
  group_by(stream, site, run_year) |>
  summarise(weeks_sampled = n_distinct(week), .groups = "drop")

excluded_lookup <- years_to_exclude_rst_data |>
  distinct(stream, site, run_year) |>
  mutate(already_excluded = TRUE)

low_effort <- weeks_in_window |>
  filter(weeks_sampled < 3) |>
  left_join(excluded_lookup, by = c("stream", "site", "run_year")) |>
  mutate(already_excluded = replace_na(already_excluded, FALSE))

new_low_effort <- log_issues(
  low_effort |>
    mutate(
      n_records   = weeks_sampled,
      data_type   = "rst",
      issue_type  = "low_sampling_effort",
      field       = "weeks_sampled",
      severity    = "moderate",
      description = glue("Only {weeks_sampled} week(s) sampled in BTSPAS window for run year {run_year} at {site} ({stream}). Already excluded from modeling: {already_excluded}")
    )
)
```

```{r}
#| label: plot-season-coverage
#| echo: false
#| fig-height: 8

coverage_plot_data <- weeks_in_window |>
  left_join(excluded_lookup, by = c("stream", "site", "run_year")) |>
  mutate(
    already_excluded = replace_na(already_excluded, FALSE),
    label = case_when(
      weeks_sampled < 3 & already_excluded ~ "Low effort (excluded)",
      weeks_sampled < 3 & !already_excluded ~ "Low effort (not excluded)",
      TRUE ~ "Adequate"
    )
  )

ggplot(coverage_plot_data, aes(x = run_year, y = site, fill = label)) +
  geom_tile(color = "white") +
  scale_fill_manual(
    values = c("Adequate" = "steelblue",
               "Low effort (excluded)" = "#fdae61",
               "Low effort (not excluded)" = "#d73027"),
    name = NULL
  ) +
  facet_wrap(~stream, scales = "free_y", ncol = 2) +
  labs(title = "Weeks sampled in BTSPAS window by site and run year",
       x = "Run year", y = NULL) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
```

`r if (nrow(low_effort) == 0) "No issues detected." else glue("{nrow(low_effort)} site/run-year combination(s) flagged ({sum(low_effort$already_excluded)} already excluded from modeling).")`

---

## Issues Logged This Run

```{r}
#| label: issues-summary
#| echo: false

all_new <- bind_rows(new_fl_na, new_fl_impl, new_catch_extreme, new_low_effort)

if (nrow(all_new) == 0) {
  cat("No new issues logged this run.")
} else {
  DT::datatable(
    all_new |> select(log_id, stream, site, run_year, issue_type, field, severity, description),
    caption = glue("{nrow(all_new)} new issue(s) added to qc_log.csv"),
    options = list(pageLength = 25, scrollX = TRUE),
    rownames = FALSE
  )
}
```

### All Open RST Issues

```{r}
#| label: open-issues
#| echo: false

full_log <- readr::read_csv(here::here("data-raw", "qc", "qc_log.csv"), show_col_types = FALSE)
open_rst  <- full_log |> filter(data_type == "rst", status == "open")

if (nrow(open_rst) == 0) {
  cat("No open RST issues in the log.")
} else {
  DT::datatable(
    open_rst |> select(log_id, date_identified, stream, site, run_year,
                       issue_type, field, severity, description, reviewer_notes),
    caption = glue("{nrow(open_rst)} open RST issue(s)"),
    options = list(pageLength = 25, scrollX = TRUE),
    rownames = FALSE
  )
}
```
