---
title: "Lifestage Ruleset"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lifestage Ruleset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width=8, fig.height=5)
```

```{r setup, echo = FALSE, warning=FALSE, message=FALSE}
library(SRJPEdata)
library(tidyverse)
library(lubridate)
# Read in datasets 
rulesets <- read.csv(here::here("data-raw", "helper-tables", "weekly_yearling_cutoff.csv")) |> 
  mutate(date_for_plot = as_date(date_for_plot))
daily_ruleset_all <- read.csv(here::here("data-raw", "helper-tables", "daily_yearling_ruleset.csv"))
```

## Generate Yearling Rulesets 

Raw catch from the trap does not consistently differentiate between yearling and young of year (YOY) Chinook. FlowWest presented a proposed methodology at a life history diversity (LHD) ruleset workshop (see [lhd shiny](https://flowwest.shinyapps.io/lhd-workshop-shiny/) for workshop materials) and worked with watershed experts to define a methodology to systematically determine life history for each tributary (described below).

**Approach**

1) Set weekly cutoff values: Use visual determination on fork length over time scatter plots to set weekly cutoff values of yearlings vs YOY. 
2) Generate daily cutoff values: Use linear interpolation to extrapolate weekly cutoffs into daily values. 
3) Review & update: Share proposed cutoff values with watershed experts to review. Update as needed. 
4) Apply cutoff to catch data: Use daily cutoff values to generate a yearling column in the catch data. 

### Set Weekly Cutoff Values & Generate daily cutoff values

FlowWest proposed weekly cutoff values and used these weekly cutoff values to generate daily values using a linear approximation function, `approxfun`. 

`generate_cutoff <- approxfun(date, fork_length_cutoff, rule = 2)`

The plot below shows the updated cutoff values with linear interpolation of weekly cutoffs for Deer Creek. 

**Note: You can view all code used to generate plots and tables in this markdown [here.](https://github.com/SRJPE/SRJPEdata/blob/main/vignettes/lifestage_ruleset.Rmd)**

```{r}
# PLOT EXAMPLE FOR DEER 
stream_name <- "deer creek"

rulesets <- read.csv(here::here("data-raw", "helper-tables", "weekly_yearling_cutoff.csv")) |> 
  mutate(date_for_plot = as_date(date_for_plot))

rulesets_filtered <- rulesets |> 
    filter(stream == stream_name) 
  
data <- rst_catch |> 
  filter(fork_length < 300) |> # for graph scaling, takes out a lot of butte creek
  filter(stream != "sacramento river") |> 
  mutate(week = week(date),
         week = if_else(week == 53, 52, week), # all December 30th or 31st
         month = month(date),
         day = ifelse(day(date) < 10, paste0("0", day(date)), day(date)),
         fake_year = ifelse(month %in% 10:12, 1971, 1972),
         # create fake date for plotting all years on same x axis
         fake_date = ymd(paste0(fake_year, "-",month, "-", day))) |> 
  select(true_fl = fork_length, week, stream, fake_date) |> 
  filter(stream == stream_name) 

# interpolate daily cutoff line from weekly cutoff values
generate_cutoff <- approxfun(rulesets_filtered$date_for_plot,
                             rulesets_filtered$fl_cutoff, rule = 2)
ruleset_lines <- tibble(
    rulesets_dates = seq(as.Date("1971-10-01"), as.Date("1972-09-30"), by = "day"),
    cutoff_line = generate_cutoff(rulesets_dates))

# format daily rulesets
daily_ruleset <- ruleset_lines |> 
    mutate(month = month(rulesets_dates),
           day = day(rulesets_dates),
           cutoff = round(cutoff_line, 2),
           stream = stream_name) |> 
    select(stream, month, day, cutoff)
  
  # join interpolated line by date and then assign
  # yearling cutoffs based on the interpolated line
data <- data |> 
    left_join(ruleset_lines, by = c("fake_date" = "rulesets_dates")) |> 
    mutate(is_yearling = case_when(true_fl <= cutoff_line ~ "Subyearling",
                                   true_fl > cutoff_line ~ "Yearling"))
  
  # plot
 ggplot() + 
    geom_point(data, mapping = aes(x = fake_date, y = true_fl, 
                                   color = is_yearling), alpha = 0.6) +
    scale_x_date(date_breaks = "1 month", date_labels = "%b",
                 limits = c(ymd("1971-10-01"), ymd("1972-09-30"))) +
    labs(y = "fork length (mm)",
         x = "") +
    theme_minimal() +
    geom_line(ruleset_lines, mapping = aes(x = rulesets_dates, y = cutoff_line)) +
    ggtitle(paste0(str_to_title(stream_name))) + 
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "bottom") + 
    labs(color = "Yearling designation") + 
    geom_point(rulesets_filtered, mapping = aes(x = date_for_plot, y = fl_cutoff), shape = 15) + 
    scale_color_manual(values = c("#440154FF", "#FDE725FF", "#21908CFF"),
                       breaks = c("Subyearling", "Yearling", "Missing weekly cutoff value"))

```

```{r, echo = FALSE}
# Create for all tribs 
generate_daily_cutoff <- function(stream_name) {
  rulesets <- read.csv(here::here("data-raw", "helper-tables", "weekly_yearling_cutoff.csv")) |> 
    mutate(date_for_plot = as_date(date_for_plot))
  
  rulesets_filtered <- rulesets |> 
    filter(stream == stream_name) 
    
  catch_data <- rst_catch |> 
    filter(fork_length < 300) |> # for graph scaling, takes out a lot of butte creek
    filter(stream == stream_name) |> 
    mutate(week = week(date),
           week = if_else(week == 53, 52, week), # all December 30th or 31st
           month = month(date),
           day = ifelse(day(date) < 10, paste0("0", day(date)), day(date)),
           fake_year = ifelse(month %in% 10:12, 1971, 1972),
           # create fake date for plotting all years on same x axis
           fake_date = ymd(paste0(fake_year, "-",month, "-", day))) |> 
    select(true_fl = fork_length, week, stream, fake_date) 
  
  # interpolate daily cutoff line from weekly cutoff values
  generate_cutoff <- approxfun(rulesets_filtered$date_for_plot,
                               rulesets_filtered$fl_cutoff, rule = 2)
  ruleset_lines <- tibble(
      rulesets_dates = seq(as.Date("1971-10-01"), as.Date("1972-09-30"), by = "day"),
      cutoff_line = generate_cutoff(rulesets_dates))
  
  # format daily rulesets
  daily_yearling_ruleset <- ruleset_lines |> 
      mutate(month = month(rulesets_dates),
             day = day(rulesets_dates),
             cutoff = round(cutoff_line, 2),
             stream = stream_name) |> 
      select(stream, month, day, cutoff)
  return(daily_yearling_ruleset)
}
streams <- c("battle creek", "butte creek", "clear creek", "deer creek", 
             "feather river", "mill creek", "yuba river")

daily_yearling_ruleset <- purrr::map(streams, generate_daily_cutoff) |> reduce(bind_rows)

usethis::use_data(daily_yearling_ruleset, overwrite = TRUE)
```


### Review & Update

FlowWest shared above plot for each watershed and asked stream experts to review. We incorporated feedback and modified rulesets to better separate yearlings and YOY. 

### Apply Cutoff to Catch Data

FlowWest took the daily cutoff line (shown in plot above) and used it as a threshold to classify yearling vs YOY in historical catch data. We added an `is_yearling` column to the catch data and set `is_yearling = TRUE` for any fish with a fork length that exceeded the yearling cutoff on a given date. 

The following code is applied in the `weekly_data_summary` script. 

```{r, include = TRUE, echo = TRUE}
# Note this is not the final dataset as lifestage is added below
standard_catch_unmarked_w_yearling <- rst_catch |> 
  filter(species == "chinook") |>  # filter for only chinook
  mutate(month = month(date), 
         day = day(date)) |> 
  left_join(daily_yearling_ruleset) |> 
  mutate(is_yearling = case_when((fork_length <= cutoff & !run %in% c("fall","late fall", "winter")) ~ F,
                                 (fork_length > cutoff & !run %in% c("fall","late fall", "winter")) ~ T,
                                 (run %in% c("fall","late fall", "winter")) ~ NA,
                                 T ~ NA)) 
```


## Fry and Smolt Designations

In addition to differentiating between yearling and YOY it is important for the SR JPE to differentiate between fry and smolt as there will likely be a separate SR JPE for each lifestage. Some monitoring programs assign lifestage based on visual determination or fork length though not all RST data included lifestage data. To ensure lifestage was assigned consistently across streams and was complete, FlowWest developed an approach for differentiating between fry and smolt.

**Approach**

1. Create `lifestage` based on forklength cutoff of 45mm (< 45 - fry, > 45 - smolt)
2. Determine year specific proportions for fry, smolt, yearling for each stream, site, week and year
3. Determine general week proportions for fry, smolt, yearling for each stream, week
4. Apply proportions to data to fill in missing lifestage

### Create lifestage variable

The first step was to apply a lifestage cutoff to catch records that had fork lengths recorded. These cutoffs are `fork_length < 45 = fry`, `fork_length > 45 = smolt`, `fork_length > yearling_cutoff = yearling`.

```{r, echo = F}
# create lifestage field based on fork length
standard_catch_unmarked <- standard_catch_unmarked_w_yearling |> 
  filter(species == "chinook") |>  # filter for only chinook
  mutate(month = month(date), # add to join with lad and yearling
         day = day(date)) |> 
  left_join(daily_yearling_ruleset) |> 
  mutate(life_stage = case_when(fork_length > cutoff ~ "yearling",
                                fork_length <= cutoff & fork_length > 45 ~ "smolt",
                                fork_length > 45  ~ "smolt",
                                fork_length <= 45 ~ "fry", # logic from flora includes week (all weeks but 7, 8, 9 had this threshold) but I am not sure this is necessary, worth talking through
                                T ~ NA)) |> 
  select(-species, -month, -day, -cutoff, -actual_count, -is_yearling) 

knitr::kable(head(standard_catch_unmarked, 10))
```

### Determine year specific lifestage proportions

There are `r standard_catch_unmarked |> filter(count != 0, is.na(life_stage)) |> tally()` entries with missing lifestage due to missing fork length data. 

```{r, echo = F}
standard_catch_unmarked |> 
  filter(count != 0, is.na(life_stage)) |> 
  head() |> 
  knitr::kable()
```

The first step in filling in these missing lifestages was to find the proportion for each lifestage category for a given stream, site, week, and year. This information could then be used to fill in the lifestage for missing rows within a week.

```{r, echo = F}
# FL-based lifestage logic ------------------------------------------------
# add logic to assign lifestage
# extrapolate lifestage for model for plus count fish/fish without fork lenghts based on weekly fl probabilities
# Create table with prob fry, smolt, and yearlings for each stream, site, week, year
weekly_lifestage_bins <- standard_catch_unmarked |> 
  filter(!is.na(fork_length), count != 0) |> # TODO consider removing this count != 0
  mutate(year = year(date), week = week(date)) |> 
  group_by(year, week, stream, site) |> 
  summarise(percent_fry = sum(life_stage == "fry")/n(),
            percent_smolt = sum(life_stage == "smolt")/n(),
            percent_yearling = sum(life_stage == "yearling")/n()) |> 
  ungroup() 

knitr::kable(head(weekly_lifestage_bins, 10))
```

### Determine general weekly lifestage proportions

For weeks that had no fork length data in a given week, we calculated a general lifestage proportion across years. Calculating the proportion for each lifestage category for a given stream, site, and week. 

```{r, echo = F}
# Use when no FL data for a year 
proxy_weekly_fl <- standard_catch_unmarked |> 
  mutate(year = year(date), week = week(date)) |> 
  filter(!is.na(life_stage)) |> 
  group_by(week, stream) |> 
  summarise(percent_fry = sum(life_stage == "fry")/n(),
            percent_smolt = sum(life_stage == "smolt")/n(),
            percent_yearling = sum(life_stage == "yearling")/n()) |> 
  ungroup()  

# Years without FL data 
# adds on stream and site variables
proxy_lifestage_bins_for_weeks_without_fl <- standard_catch_unmarked |> 
  group_by(year = year(date), week = week(date), stream, site) |> 
  summarise(fork_length = mean(fork_length, na.rm = TRUE)) |> 
  filter(is.na(fork_length)) |> #find the missing fork lengths
  left_join(proxy_weekly_fl, by = c("week" = "week", "stream" = "stream")) |> # apply the weekly proportions
  select(-fork_length) 

# bind together the specific and general lifestage proportions
all_lifestage_bins <- bind_rows(weekly_lifestage_bins, proxy_lifestage_bins_for_weeks_without_fl)


```

### Apply proportions to fill in missing values

We used these proportions to fill in missing lifestage values. See the final lifestage designations below. 

```{r, echo = F}
# create table of all na values that need to be filled
# total of 42,723, started with 25964, increases number of rows because add row for different lifestage as lifestage was previously not assigned
na_filled_lifestage <- standard_catch_unmarked |> 
  mutate(week = week(date), year = year(date)) |> 
  filter(is.na(fork_length) & count > 0) |> 
  left_join(all_lifestage_bins, by = c("week" = "week", "year" = "year", "stream" = "stream", "site" = "site")) |> 
  mutate(fry = round(count * percent_fry), 
         smolt = round(count * percent_smolt), 
         yearling = round(count * percent_yearling)) |> 
  select(-life_stage, -count, -week, -year, -c(percent_fry, percent_smolt, percent_yearling)) |> # remove because all na, assigning in next line
  pivot_longer(fry:yearling, names_to = 'life_stage', values_to = 'count') |> 
  filter(count != 0) |> # remove 0 values introduced when 0 prop of a lifestage, significantly decreases size of DF, note that this may need to be added back for modeling purposes
  mutate(model_lifestage_method = "assign count based on weekly distribution",
         week = week(date), 
         year = year(date)) 

# add filled values back into combined_rst 
# first filter combined rst to exclude rows in na_to_fill
# total of 460,461 rows
combined_rst_wo_na_fl <- standard_catch_unmarked |> 
  mutate(week = week(date), year = year(date)) |> 
  filter(!is.na(fork_length) & count > 0) |> # there are a few with 0 catch and fork lengths
  mutate(model_lifestage_method = "assigned from fl cutoffs") 

# weeks we cannot predict lifestage
gap_weeks <- proxy_lifestage_bins_for_weeks_without_fl |> 
  filter(is.na(percent_fry) & is.na(percent_smolt) & is.na(percent_yearling)) |> 
  select(year, week, stream, site)

formatted_standard_catch <- standard_catch_unmarked |> 
  mutate(week = week(date), year = year(date)) 

# only 19 of these: all feather river 1999
weeks_wo_lifestage <- gap_weeks |> 
  left_join(formatted_standard_catch, by = c("year" = "year", "stream" = "stream", "week" = "week", "site" = "site")) |> 
  filter(!is.na(count), count > 0) |> 
  mutate(model_lifestage_method = "Not able to determine, no weekly fl data ever") 

# 25,393 with no catch
no_catch <- standard_catch_unmarked |> 
  mutate(week = week(date), year = year(date)) |>
  filter(is.na(fork_length) & count == 0) 

# total rows of 528,596
# expected to be larger than original because add lifestage rows
updated_standard_catch <- bind_rows(combined_rst_wo_na_fl, na_filled_lifestage, no_catch, weeks_wo_lifestage) 

knitr::kable(head(updated_standard_catch, 10))
```

### Review lifestage

The following plot shows the general patten in the lifestage field where fry are caught earlier in the year and smolt are caught later in the year.

**Battle Creek: 2011**

```{r, echo = F}
updated_standard_catch |> 
  filter(stream == "battle creek", year == 2011) |> 
  mutate(fake_year = ifelse(month(date) %in% 10:12, 2000, 2001),
         fake_date = as_date(paste0(fake_year, "-", month(date), "-", day(date)))) |> 
  ggplot(aes(x = fake_date, y = count, fill = life_stage)) +
  geom_col() +
  scale_fill_manual(values = c("#440154FF", "#FDE725FF", "#21908CFF")) +
  labs(x = "",
       y = "count",
       fill = "") +
  theme_minimal()
```

## Save resulting data to package

```{r}
# Save to package
usethis::use_data(updated_standard_catch, overwrite = TRUE)
```
