---
title: "Run Assignment"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Run Assignment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6, 
  fig.height = 4,
  fig.align = "center"
)
```

```{r setup, include = FALSE}
library(SRJPEdata)
library(waterYearType)
source("data-raw/build_model_datasets.R")
```
The run variable in RST data is incomplete for several streams. Run assignment in the JPE will eventually be assigned using a Probability of Length-at-Date (PLAD) model, but for initial model development run was assigned using weekly averages within streams and sites.  

This file explores variability in historical run type as assigned by stream teams, and describes the methods used to calculated proportion of spring run by week, stream, and site using historical averages.

Data were checked and cleaned in scripts available [here](https://github.com/FlowWest/JPE-datasets/tree/main/scripts/rst).
Cleaned data were saved on the JPE cloud database and pulled using the `pull_tables_from_database.R` script. Data tables were initially processed in the `build_model_datasets.R` script.


## Variability in Spring Run Proportions

Initial inspection of RST data available showed that, for several tributaries, run data were incomplete. Deer and Mill Creek do not record run information, `28.9%` of Yuba River's run records were `NA`, and Battle Creek and Sacramento River records were `14.9%` and `13.7% NA`, respectively. Butte Creek, Clear Creek, and Feather River had the most complete run data out of all tributaries:

```{r}
updated_standard_catch |> 
  mutate(run = ifelse(run %in% c("not recorded", "unknown"), NA, run)) |> 
  group_by(stream) |> 
  summarise(proportion_NA = sum(is.na(run))/n()) |> 
  arrange(desc(proportion_NA))
```

Here we analyze how run proportions vary by `year`, `water year type`, `stream`, and `site`.

### Data preparation

First, we standardize run type assignment across the entire dataset:
```{r}
standardized_run <- updated_standard_catch |> 
  mutate(run = case_when(run %in% c("not recorded", "unknown") ~ NA,
                         run %in% c("late fall", "fall", "winter") ~ "not_spring",
                         TRUE ~ run))

standardized_run |> 
  distinct(run)

```

### Stream {.tabset}

By proportion of total count, Butte and Battle Creek show the most spring run Chinook salmon. Deer Creek and Mill Creek do not collect run data, and so all their run type assignments are `NA`. Feather River, Sacramento River, Yuba River, and Clear Creek show higher proportions of other runs. 

```{r}
standardized_run |> 
  group_by(stream, year) |> 
  count(run) |> 
  mutate(freq = n / sum(n)) |> 
  ggplot(aes(x = year, y = freq, fill = run)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~stream, scales = "free_y") +
  xlab("Year") +
  ylab("Proportion of count by run") +
  ggtitle("Proportions of spring run Chinook by stream and year") + 
  theme_minimal()
```

## Year {.tabset}

In earlier years, few records were assigned run type and so `NA` dominates. However, in later years the proportion of records assigned `NA` decrease and the ratio of `spring` to `not_spring` run records appears to stabilize.

```{r}
standardized_run |> 
  group_by(year) |> 
  count(run) |> 
  mutate(freq = n / sum(n)) |> 
  ggplot(aes(x = year, y = freq, fill = run)) + 
  geom_bar(stat = "identity") +
  xlab("Year") +
  ylab("Proportion of count by run") +
  ggtitle("Proportions of spring run Chinook by year") + 
  theme_minimal()
```

### Water Year Type {.tabset}

We used the package `waterYearType` to assign water year types to the years in the dataset. We simplified water year types into `Wet` and `Dry` using the following logic: 

- `Wet`, `Above Normal` = `Wet`
- `Below Normal`, `Dry`, `Critical` = `Dry`

Yuba River and Feather River are the only streams where the water year type appears to show a difference in proportion of catch that is spring run.

```{r}
standardized_run_wy <- standardized_run |> 
  left_join(water_year_indices |> 
              filter(location == "Sacramento Valley") |> 
              mutate(Yr_type = ifelse(Yr_type %in% c("Wet", "Above Normal"), "Wet", "Dry")) |> 
              select(year = WY, wy_type = Yr_type),
            by = "year") |> 
  mutate(wy_type = ifelse(year == 2022, "Dry", wy_type))

standardized_run_wy |> 
  group_by(wy_type, stream) |> 
  count(run) |> 
  mutate(freq = n / sum(n)) |> 
  ggplot(aes(x = wy_type, y = freq, fill = run)) + 
  geom_bar(stat = "identity") +
  xlab("Year") +
  ylab("Proportion of count by run") +
  ggtitle("Proportions of spring run Chinook by water year type and stream") + 
  theme_minimal() +
  facet_wrap(~stream)
```

## Site {.tabset}

Proportion spring run appeared to vary by site and year. For example, Clear Creek showed a distinct difference in proportion spring run by its two sampling sites.

```{r}
standardized_run |> 
  group_by(stream, site, year) |> 
  count(run) |> 
  mutate(freq = n / sum(n)) |> 
  filter(run == "spring") |> 
  ggplot(aes(x = year, y = freq, color = site)) + 
  geom_line() +
  facet_wrap(~stream) +
  xlab("Year") +
  ylab("Proportion spring run") +
  ggtitle("Proportions of spring run by site and stream") + 
  theme_minimal()
```
Clear Creek:

```{r}
standardized_run |> 
  group_by(stream, site, year) |> 
  count(run) |> 
  mutate(freq = n / sum(n)) |> 
  filter(stream == "clear creek") |> 
  ggplot(aes(x = year, y = freq, fill = run)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~site) +
  xlab("Year") +
  ylab("Proportion of count") +
  ggtitle("Run proportions in Clear Creek by site") + 
  theme_minimal()
```

## Assigning Run Type Using Historical Averages


### Methods

Prior to initializing stock-recruit analysis, estimates of annual chinook outmigrant abundance needed to be converted to annual *spring run* chinook outmigrant abundance. One approach would be to multiply weekly abundance estimates by proportion of spring run fish by week. Ultimately, the PLAD model will be applied to historical data. In the interim, existing run type assignments were used to interpolate missing weeks to get proportion of spring run.

Our methodology (performed for every tributary except Deer Creek and Mill Creek): 

1) To get the correct weeks where data were missing, we assigned `NA` values to all run type assignments that provided no information (i.e. `not recorded` and `unknown` records)

```{r}
updated_standard_catch_na_run <- updated_standard_catch |> 
  mutate(run = ifelse(run %in% c("not recorded", "unknown", NA_character_), NA_character_, run)) |> glimpse()
```

2) For each week where we had run type assignment data and counts above 0,  we calculated the `proportion_spring_run` and `proportion_not_spring_run` within years and sites.
```{r}
updated_standard_catch_na_run_no_deer_mill <- updated_standard_catch_na_run |> 
  filter(!stream %in% c("deer creek", "mill creek"))

weekly_run_bins <- updated_standard_catch_na_run_no_deer_mill |> 
  filter(!is.na(run), count != 0) |> 
  mutate(year = year(date), week = week(date)) |> 
  group_by(year, week, stream, site) |>
  summarize(percent_spring = sum(run == "spring", na.rm = T)/n(),
            percent_not_spring = sum(run != "spring", na.rm = T) / n()) |> 
  ungroup()

weekly_run_bins |> arrange(stream, year, week) |> head()
```


3) There were some years where there weren't weekly data to interpolate. For these weeks, we created proxy weekly bins (summarizing across years, but still within sites). We added these proxy bins to the weeks where data were missing

```{r}
# Use when no run data for a year 
proxy_weekly_run <- updated_standard_catch_na_run_no_deer_mill |>
  mutate(year = year(date), week = week(date)) |>
  filter(count > 0, !is.na(run)) |> 
  # do not group by year
  group_by(week, stream, site) |>
  summarise(percent_spring = sum(run == "spring", na.rm = T)/n(),
            percent_not_spring = sum(run != "spring", na.rm = T)/n()) |>
  ungroup()

proxy_run_bins_for_weeks_without_run <- updated_standard_catch_na_run_no_deer_mill |>
  filter(count > 0) |> 
  group_by(year, week, stream, site) |>
  summarise(all_na = sum(is.na(run)) == n()) |> 
  ungroup() |> 
  filter(all_na) |> 
  left_join(proxy_weekly_run, by = c("week", "stream", "site")) |>
  select(-all_na)

proxy_run_bins_for_weeks_without_run |> arrange(stream, year, site) |> head()
```

4) We added these two datasets together to get full coverage of weekly proportions of spring/not spring run Chinook.

```{r}
all_run_bins <- bind_rows(weekly_run_bins, proxy_run_bins_for_weeks_without_run) |>
  glimpse()
```

5) We took the original dataset and filtered to those records where count was above 0 and run type was not assigned. For these records, we used the weekly proportions to fill in counts of spring run and not spring run Chinook:

```{r}
na_filled_run <- updated_standard_catch_na_run_no_deer_mill |> 
  filter(is.na(run) & count > 0) |> 
  left_join(all_run_bins, by = c("year", "week", "stream", "site")) |> 
  mutate(spring_run = round(count * percent_spring),
         not_spring_run = round(count * percent_not_spring)) |> 
  select(-c(count, week, year, run)) |> # remove bc all NA, assigning in next line
  pivot_longer(spring_run:not_spring_run, names_to = 'run_for_model', values_to = 'count') |> 
  select(-c(percent_spring, percent_not_spring)) |>  
  filter(count != 0) |> # remove 0 values introduced when 0 prop of a lifestage, significantly decreases size of DF 
  mutate(model_run_method = "assign run based on weekly distribution",
         week = week(date), 
         year = year(date)) |> 
  select(-run_method) |> 
  glimpse()
```

6) We then combined the records filled in with historical averages with the rest of the original dataset, which included: 

* all original records where count was above 0 and run type was assigned,
* all original records where count was 0, and
* all records for Deer Creek and Mill Creek, which did not have any run type assigned: 

```{r}
combined_rst_wo_na_run <- updated_standard_catch_na_run_no_deer_mill |> 
  filter(!is.na(run) & count > 0) |> 
  mutate(run_for_model = if_else(run == "spring", "spring_run", "not_spring_run")) |> 
  mutate(model_run_method = ifelse(is.na(run_method), "not recorded", run_method)) |> 
  select(-run_method)

mill_and_deer <- updated_standard_catch_na_run |> 
  filter(stream %in% c("mill creek", "deer creek")) |> 
  mutate(run_for_model = NA,
         model_run_method = "mill and deer - no data to interpolate") |> 
  select(-run_method)

no_catch_run <- updated_standard_catch_na_run_no_deer_mill |> 
  filter(count == 0) |> 
  mutate(run_for_model = NA,
         model_run_method = "count is 0") |> 
  select(-run_method)

updated_standard_catch_with_run <- bind_rows(combined_rst_wo_na_run, na_filled_run, no_catch_run, mill_and_deer) |> glimpse()

```

### Results

For each tributary, site, and week, the catch dataset now has a run type assignemnt for the model that is either `spring_run`, `not_spring_run`, or `NA`:

```{r}
updated_standard_catch_with_run |> 
  mutate(run_for_model = case_when(run_for_model == "spring_run" ~ "Spring run", 
                                   run_for_model == "not_spring_run" ~ "Fall, Late Fall, Winter run",
                                   is.na(run_for_model) ~ NA_character_)) |> 
  ggplot() + 
  geom_line(aes(x = date, y = count, color = run_for_model), alpha = 0.8) +
  facet_wrap(~stream, scales = "free") + 
  theme_minimal() + 
  xlab("Year") + ylab("Count of fish") +
  ggtitle("Run type assignment supplemented with historic averages")

```


### Key Assumptions

- We assumed that taking the weekly averages across years and sites (within each stream) would be representative. Run type assignment did vary by site:

```{r}
updated_standard_catch_na_run |> 
  filter(stream == "clear creek") |> 
  group_by(site, week) |> 
  summarise(n = n(),
            prop_spring = sum(run == "spring", na.rm = T) / n,
            prop_other = sum(run %in% c("fall", "late fall", "winter", NA_character_) / n)) |> 
  ggplot(aes(x = week, y = prop_spring)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~site, scales = "free_y") + 
  xlab("Week") + ylab("Proportion spring run") +
  ggtitle("Proportion spring run by week and site for Clear Creek") + 
  theme_minimal()
```

We did not include subsite in this grouping due to data limitations; however, initial analyses did show that run type assignment varied by subsite:

```{r}
updated_standard_catch_na_run |> 
  filter(stream == "butte creek", !is.na(subsite)) |>
  group_by(subsite, site, week) |> 
  summarise(n = n(),
            prop_spring = sum(run == "spring", na.rm = T) / n,
            prop_other = sum(run %in% c("fall", "late fall", "winter", NA_character_) / n)) |> 
  ggplot(aes(x = week, y = prop_spring)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~subsite, scales = "free_y") + 
  xlab("Week") + ylab("Proportion spring run") + 
  ggtitle("Proportion spring run by week and subsite for Battle Creek") +
  theme_minimal()
```

- We converted all run type records encoded `not recorded` and `unknown` to `NA` for the purpose of this analysis, though this method is not used across other datasets within the JPE.

- Deer Creek and Mill Creek did not contain any run type assignment data to interpolate, and so we left these streams out and they will need another method to be useful in the JPE.

- There were some weeks within years with no historical data with which to calculate averages, and for those we used the average for those weeks across all years within the dataset:

```{r}
updated_standard_catch_na_run_no_deer_mill |>
  filter(count > 0) |> 
  group_by(year, week, stream, site) |>
  summarise(all_na = sum(is.na(run)) == n()) |> 
  ungroup() |> 
  filter(all_na) |> 
  group_by(stream, year) |> 
  tally() |> 
  rename(number_weeks_with_no_historical_data = n) |> 
  arrange(desc(number_weeks_with_no_historical_data), stream)
```

