---
title: "Draft of RST Sites Used in JPE"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Draft of RST Sites Used in JPE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(tidyverse)
library(lubridate)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  fig.width=8, fig.height=6
)
```

```{r setup, echo = FALSE}
library(SRJPEdata)
# source(here::here("data-raw", "pull_tables_from_database.R"))

colors_small <-  c("#9A8822", "#F5CDB4", "#F8AFA8", "#FDDDA0", "#74A089", #Royal 2
                            "#899DA4", "#C93312", "#DC863B" # royal 1 (- 3)
)
colors_full <-  c("#9986A5", "#C18748", #Royal 2
                  "#899DA4", "#9A8822", # Grand Budapest 1 (-4)
                  "#02401B", "#A2A475", # Cavalcanti 1
                  "#E6A0C4", "#C6CDF7", "#D8A499", "#7294D4", #Grand Budapest 2
                  "#9986A5", "#EAD3BF", "#AA9486", "#B6854D", "#798E87", # Isle of dogs 2 altered slightly
                  "#AC9765", "#b67c3b", "#175149", "#AF4E24", "#CEB175",
                  "#E54E21", "#6C8645", "#C18748", "#FBA72A", "#D3D4D8", "#CB7A5C", "#5785C1")

```

### Overview

The spring run juvenile production estimate (SR JPE) contains a suite of submodels that rely on different datasets. This document is currently focused on Chinook monitoring data that will be used to fit the stock recruit model. 

## Stock Recruit Rotary Screw Trap Site Selection Overview

The Spring Run JPE model considers data from 7 tributaries and the mainstem Sacramento river, including: 

* Battle Creek 
* Butte Creek 
* Clear Creek
* Deer Creek 
* Feather River 
* Mill Creek 
* Yuba River

Each of these tributaries has a unique monitoring setup and number of Rotary Screw Traps that we can use in the SR JPE model. This vignette describes the trapping sites that we use for each tributary. 

### Map
```{r}

```


## Choosing Sites 

### Battle Creek 
Use UBC (LBC older site, has not been fished for a long time) 


### Butte Creek 


### Clear Creek
Clear Site Decision  
Use LCC for total JPE - captures total spawning ground 
Use UCC for supplementary analysis  
Use for understanding of run & potential survival between traps  
While model is still in development compare UCC (all spring) to LCC PLAD results  
If results surprising, consider using both traps  
IF for some reason LCC (SR based on plad) and UCC counts do not align, we can reassess our decision and do an adult data analysis to come up with an adult scaling factor (what % of redds are below the UCC trap)  

### Deer Creek 
Only one site. 

### Feather River 
Feather Site Decision  
Use a similar approach as clear  
Use lower trap for total JPE (HFC) - captures total spawning ground 
Use LFC for supplementary analysis  
First round do analysis to compare if we need to use LFC as well 

### Mill Creek 
Only one site. 

### Yuba River
Yuba Site Decision  
Use hallwood, site that they are currently trapping, site with most historical data, and lowest site in system  







