---
title: "Calculating Trap Effort"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{trap_effort}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(tidyverse)
library(lubridate)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  fig.width=10, fig.height=8
)
```

```{r setup}
library(SRJPEdata)
source("data-raw/pull_tables_from_database.R")
```

Hours fished is not included as a variable in the data and needs to be calculated.
This file calculates hours fished by date, stream and site.

Data were checked and cleaned in scripts available [here](https://github.com/FlowWest/JPE-datasets/tree/main/scripts/rst).
Cleaned data were saved on the JPE cloud database and pulled using the `pull_tables_from_database.R` script. 

## Method

We explored three primary methods to calculate hours fished. Methods were chosen based on data collected at the RST site. When no times, trap revolutions, or trap RPMs were collected we assume 24 hours for each day that trap is fishing. The trap revolution method is commonly used by RST programs but we were not able to get reasonable results for most programs we tried it on so we ended up relying on trap times instead of revolutions. This may be due to different methods for determining cone rotation rate. For instance, Battle measures the average time per rotation whereas Feather River and Knights Landing measure the revolutions per minute. Using total revolutions and the cone rotation rate to calculate sampling period may be useful in the future when methods are more similar.  

1) Use start and stop time. 
2) Use rpms start, rpms end, and total revolutions (Method 2 was used as a comparison method but was ultimatly not picked as the final method used for any traps.)
3) Use start time only 

Below is the code for each of the methods described above: 

1) If the start and stop date/time of the sampling period are available, simply subtract the difference in time.

```{r, helpers}
# calculating hours fished when have start and stop datetime
hours_fished <- function(dat){
  dat %>%
    filter(!is.na(trap_stop_time), !is.na(trap_start_time)) %>%
    mutate(start_datetime = ymd_hms(paste(trap_start_date, trap_start_time)),
           stop_datetime = ymd_hms(paste(trap_stop_date, trap_stop_time)),
           hours_fished = round(difftime(stop_datetime, start_datetime, units = "hours"), 2))
}
```

2) If the RPMs start, RPMs end, and total revolutions are available, 
take the average cone RPMs and divide `total_revolutions`/average RMP/60 to get total hours fished.

```{r, helpers2}
# calculating hours fished when have start and stop datetime
revolution_calculated_hours_fished <- function(dat){
  dat %>%
    filter(!is.na(rpm_start) | !is.na(rpm_end), !is.na(total_revolutions)) %>% 
    mutate(prior_day_rpm = ifelse(is.na(rpm_end), NA, lag(rpm_end)),
           sum_rpms = ifelse(is.na(prior_day_rpm), rpm_start, rpm_start + prior_day_rpm), 
           cone_rpms = sum_rpms / 2,
           hours_fished = case_when(is.na(rpm_start) ~ round(total_revolutions/rpm_end/60, 2),
                                    is.na(rpm_end) ~ round(total_revolutions/rpm_start/60, 2), 
                                    TRUE ~ round(total_revolutions/cone_rpms/60, 2)))
}
```

3) If only one date is available, assume that the end date is the date sampled on the following day.

```{r, helper3}
# calculating hours fished when have only date and time
hours_fished_one_date <- function(dat) {
  dat %>%
  arrange(site, subsite, start_datetime) %>%
  mutate(end_datetime = lead(start_datetime),
         end_datetime = case_when(difftime(end_datetime, start_datetime, units = "hours") > 120 ~ start_datetime + hours(24), 
                                  T ~ end_datetime),
         hours_fished = round(difftime(end_datetime, start_datetime, units = "hours"), 2))
}
```

### Other Assumptions

Based on protocols and information from monitoring programs, sampling periods are typically 24 hours,
unless flows are very high in which traps are checked more frequently. Calculating hours fished based
on time and date provided in data provides more specificity, however, there are typos and errors resulting
in nonsensical hours fished. In these cases, we assume 24 hour sampling periods. 

- If only a date is available and no time, assume 24 hour sampling periods for each date.
- If information is missing or there are typos, assume 24 hour sampling periods.

## Hours fished {.tabset}

## Method 1) Battle creek, Clear Creek, Feather, Kinghts, Tisdale

The above creeks have data available to calculate the hours fished. There are a few
cases when start date is NA. For these we fill in the start date assuming
it is the sample date of the previous record.

```{r, battle}
method_one_hours_fished <- filter(trap, stream %in% 
                                c("battle creek", "clear creek", 
                                  "feather river", "sacramento river")) %>% 
  select(trap_start_date, trap_start_time, trap_stop_date, trap_stop_time, 
         stream, site, subsite, ) %>% 
  arrange(site, subsite, trap_stop_date, trap_stop_time) %>% 
  mutate(hours_fished_methodology = "using start time and stop time") %>% 
  hours_fished()

knitr::kable(head(method_one_hours_fished, 5))
```

### QC

```{r}
method_one_hours_fished %>% 
  ggplot(aes(x = stop_datetime, y = hours_fished, color = stream)) +
  geom_point()+
  theme_minimal()

method_one_hours_fished %>% 
  ggplot(aes(x = hours_fished, fill = stream)) +
  geom_histogram(position = "dodge") +
  theme_minimal()
```


## Method 3: Butte creek & Yuba 

Hours fished are calculated by taking the difference between the date/time at each 
observation. 

```{r, butte, yuba}
## Butte ####
method_three_hours_fished <- filter(trap, stream %in% c("butte creek", "yuba river")) %>%
  distinct(trap_stop_date, trap_stop_time, site, subsite, stream) %>%
  mutate(start_datetime = ymd_hms(paste(trap_stop_date, trap_stop_time))) %>%
  mutate(hours_fished_methodology = "using only start time") %>% 
  hours_fished_one_date() %>% 
  mutate(end_datetime = case_when(trap_stop_date == "2007-05-03" ~ ymd_hms("2007-05-04 08:00:00"),
                                  trap_stop_date == "2008-04-23" ~ ymd_hms("2008-04-24 07:15:00"),
                                  trap_stop_date == "2015-06-03" ~ ymd_hms("2015-06-04 08:00:00"),
                                  T ~ end_datetime),
         hours_fished = round(difftime(end_datetime, start_datetime, units = "hours"), 2),
         hours_fished = ifelse(hours_fished < 0, 24, hours_fished))
# tried treating post 2015 differnt since it is cAMP data and has start and end dates but caused a lot of huge outliers
knitr::kable(head(method_three_hours_fished, 5))
```

### QC

```{r}
method_three_hours_fished %>% 
  ggplot(aes(x = trap_stop_time, y = hours_fished, color = stream)) +
  geom_point()
# trying to figure out where to set cut of for effort
# decided to set as 5 days - 120 hours
method_three_hours_fished %>% 
  filter(hours_fished < 200) %>% 
  ggplot(aes(x = hours_fished, fill = stream)) +
  geom_histogram(position = "dodge") + theme_minimal()
# these are typos and will be fixed by assuming 24 hours fished
# butte_hours_fished %>% 
#   filter(hours_fished > 5000) %>% tally()
```

## Mill & Deer creek

Mill & Deer Creek do not have data available (no time variable) to calculate hours
Mill & Deer Creek also does not have sample period revolutions so we cannot use trap revolutions 
fished so assume 24 hours for every day with data. 

```{r, deer & mill}
## Deer & mill #####
deer_mill_hours_fished <- filter(trap, stream %in% c("deer creek", "mill creek")) %>% 
  distinct(trap_stop_date, stream, site, subsite) %>% 
  mutate(hours_fished_methodology = "24 hour assumption") %>% 
  mutate(hours_fished = 24)

knitr::kable(head(deer_mill_hours_fished, 5))
```


# Combine data

```{r}
# all wee need is stop_date, hours_fished, stream, site
# do not sum by subsites. this can be done in further analysis

method_one <- method_one_hours_fished %>% 
  select(stream, site, subsite, trap_stop_date, trap_stop_time, hours_fished) %>% 
  mutate(hours_fished = as.numeric(hours_fished))

method_three <- method_three_hours_fished %>% 
  select(stream, site, subsite, trap_stop_date, trap_stop_time, hours_fished) %>% 
  mutate(hours_fished = as.numeric(hours_fished))

mill_and_deer <- deer_mill_hours_fished %>% 
  select(stream, site, subsite, trap_stop_date, hours_fished) %>% 
  mutate(hours_fished = as.numeric(hours_fished))

hours_fished_combined <- bind_rows(method_one,
                                   method_three,
                                   mill_and_deer) %>% 
  # if hours fished NA assumed 24 hours
  mutate(hours_fished = ifelse(is.na(hours_fished), 24, hours_fished)) %>% 
  rename(date = trap_stop_date) %>% 
  group_by(stream, site, subsite, date) %>% 
  summarize(hours_fished = sum(hours_fished)) 
```

## Summarized by week

```{r}
hours_fished_weekly <- hours_fished_combined %>% 
  mutate(week = week(date),
         year = year(date)) %>% 
  group_by(stream, site, subsite, week, year) %>% 
  summarize(hours_fished = sum(hours_fished)) 
```

### QC

```{r}
hours_fished_weekly %>% 
  ggplot(aes(x = week, y = hours_fished, color = year)) +
  geom_point() + 
  facet_wrap(~stream)+ 
  theme_minimal()

hours_fished_weekly %>% 
  ggplot(aes(x = hours_fished, fill = stream)) +
  geom_histogram(alpha = .75, position = 'dodge') +
  theme_minimal()
```
