---
title: "Years to Include"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Years to Include}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE, 
  message = FALSE,
  comment = "#>", 
  fig.width=8, fig.height=5)

if (requireNamespace("devtools", quietly = TRUE)) devtools::load_all(".", quiet = TRUE)
library(SRJPEdata)

library(tidyverse)
library(lubridate)

colors_full <-  c("#9A8822", "#F5CDB4", "#F8AFA8", "#FDDDA0", "#74A089", #Royal 2
                  "#899DA4", "#C93312", "#DC863B", # royal 1 (- 3)
                  "#F1BB7B", "#FD6467", "#5B1A18", # Grand Budapest 1 (-4)
                  "#D8B70A", "#02401B", "#A2A475", # Cavalcanti 1
                  "#E6A0C4", "#C6CDF7", "#D8A499", "#7294D4", #Grand Budapest 2
                  "#9986A5", "#EAD3BF", "#AA9486", "#B6854D", "#798E87" # Isle of dogs 2 altered slightly
)
years_to_exclude_adult <- read_csv(here::here("data-raw", "helper-tables", "years_to_exclude_adult_datasets.csv"))  
```

The Spring Run Juvenile Production Estimate uses historical rotary screw trap and adult survey data. In order to improve model performance the modeling team worked with the stream teams to filter out data that is too incomplete to use for modeling. This article describes the process for selecting RST years to include and adult years to include.

Historically this process differed for each tributary and was based on the rules below. This ruleset was complicated and difficult to maintain. In April 2026 we decided to transition to a consistent ruleset to apply across all years and streams. We compared a number of thresholds and determined that a rule where at least 60% of the weeks (45:53, 1:22) were sampled was appropriate - it was not too stringent and did not include too many gaps.

## RST Data - Historical Ruleset

The modeling team hosted a modeling windows workshop where we discussed approaches to defining the time window that should be included in the juvenile abundance model. At the workshop we decided to use the full sampling season of data for each tributary, site, and year but to exclude years where we were concerned about data completeness. We came up with the following approach to determine excluded years:

1)  First exclude years with lots of missing data from cumulative catch curves. See `data-raw/years_to_exclude.csv` (exclusion_type = "really low sampling").
2)  Exclude yearlings from cumulative catch curves.
3)  Use updated cumulative catch curves to determine the critical window (average window over all historical years that captures 75% percent of catch).
4)  Remove additional years where there is no sampling for 4 consecutive weeks within the critical window. See `data-raw/years_to_exclude.csv` (exclusion_type = "missing four consecutive weeks in critical window").

### Heat map of all tributaries

The heat map below shows when sampling occurs for all streams. It shows that sampling is more complete across traps from 2004 - 2009 and from mid November to July. Some traps have continued sampling throughout the season. Some years there are gaps within season for specific traps or across multiple traps.

```{r, echo = FALSE}
# Need to join catch_data_week to make sure all weeks get included in the 
# sampling window heatmap
catch_date_week <- SRJPEdata::rst_catch %>% 
  mutate(day = day(date),
         month = month(date),
         fake_date = as_date(paste0("2000-", month, "-", day))) %>%
  select(fake_date) %>%
  distinct() %>%
  mutate(week = week(fake_date))
# This helper table provides a date range for each numeric week
# This is used as hover text label in the heatmap
date_range <- catch_date_week %>%
  group_by(week) %>%
  slice_min(fake_date) %>%
  rename(min_date = fake_date) %>%
  left_join(catch_date_week %>%
              group_by(week) %>%
              slice_max(fake_date) %>%
              rename(max_date = fake_date)) %>%
  mutate(min_date = format(min_date, "%b %d"),
         max_date = format(max_date, "%b %d"),
         date_range = paste(min_date,"-",max_date))

catch_summary <- SRJPEdata::rst_catch %>%
  select(date, stream) %>%
  distinct() %>% 
  mutate(week = week(date),
         year = year(date)) %>%
  group_by(stream, week, year) %>%
  tally()

catch_weekly <- catch_summary %>% 
  # add catch_date_week to make sure all weeks are included in the data
  right_join(select(catch_date_week, week) %>% 
               distinct()) %>% 
  pivot_wider(id_cols = c("stream", "year"), 
              names_from = week, 
              values_from = n, 
              values_fill = 0) %>%
  pivot_longer(cols = -c("stream","year"), names_to = "week", values_to = "number_days_sampled") %>%
  mutate(week = as.numeric(week)) %>%
  left_join(date_range) %>%
  mutate(fake_date = case_when(week %in% c(40:53) ~ ymd("1999-01-01") + weeks(week - 1),
                               T ~ ymd("2000-01-01") + weeks(week - 1)))

plot_data <- catch_weekly %>%
      filter(number_days_sampled > 0) %>%
      mutate(wy = ifelse(week >= 37, year + 1, year)) %>%
      group_by(week, wy, date_range, fake_date) %>%
      summarize(streams = paste(unique(stream), collapse = ", "),
                n_streams = length(unique(stream))) %>%
      ungroup() %>%
      mutate(week = factor(week, levels = c(37:53, 1:36)),
             wy = factor(wy))
year_labels <- c(1995, 2000, 2005, 2010, 2015, 2020)

heat_map <- ggplot(plot_data, aes(x = fake_date, y = wy, fill = n_streams, text = paste0(date_range, "<br>", streams))) +
  geom_tile() +
  #scale_y_continuous(n.breaks = 5, breaks = waiver()) +
  scale_y_discrete(breaks = year_labels) +
  scale_x_date(date_labels = "%b", breaks = "2 months") +
  theme_minimal() + 
  ylab('water year') +
  xlab("") + 
  # scale_fill_viridis_b(option = "A", n.breaks = 8, direction = -1) +
  labs(fill = "Number of streams")
plotly::ggplotly(heat_map) 
# heat_map
```

### RST Data - Current Ruleset

Ruleset defined as "at least 19 weeks included within a run year (week 45-53 and 1-22) for a given site" or 60% of weeks sampled.

```{r, echo = FALSE}
exclusion_catch <- SRJPEdata::rst_catch |>
  filter(!is.na(count)) |> # remove when trap is not fishing
  mutate(
    week = week(date),
    year = year(date),
    run_year = ifelse(week >= 45, year + 1, year)
  ) |>
  filter(week %in% c(45:53, 1:22)) |> # select to the week window used in BTSPASX
  select(stream, site, subsite, site_group, week, run_year) |>
  distinct() |>
  group_by(stream, site, subsite, site_group, run_year) |>
  summarise(number_weeks = n()) |>
  ungroup() |>
  group_by(stream, site, site_group, run_year) |>
  summarize(number_weeks = max(number_weeks)) |> # find max number of weeks when group by site
  # other exclusion rules are included in case this changes in the future
  mutate(
    exclude_20 = ifelse(number_weeks < 6, TRUE, FALSE),
    exclude_40 = ifelse(number_weeks < 13, TRUE, FALSE),
    exclude_60 = ifelse(number_weeks < 19, TRUE, FALSE),
    exclude_65 = ifelse(number_weeks < 21, TRUE, FALSE),
    exclude_70 = ifelse(number_weeks < 22, TRUE, FALSE),
    exclude_75 = ifelse(number_weeks < 24, TRUE, FALSE)
  ) 
years_to_exclude_rst_data <- exclusion_catch |> 
  filter(exclude_60 == T) |> 
  select(stream, site, run_year, number_weeks) |> 
  mutate(exclusion_type = "60% of weeks sampled")

usethis::use_data(years_to_exclude_rst_data, overwrite = TRUE)
years_exclude_nice_names <- years_to_exclude_rst_data |> select("Stream" = stream, 
                                                                "Site" = site, 
                                                    "Run Year" = run_year, 
                                                    "Exclusion Type" = exclusion_type)
knitr::kable(head(years_exclude_nice_names, 10))
```

*... with `r nrow(years_to_exclude_rst_data) - 10` more rows*


### Applying to Modeling Datasets

In order to apply the years to exclude information to the modeling datasets we did some additional processing to create a table describing the stream, site, year that should be included in the SR JPE modeling.

The table below shows a section of this table:

```{r, echo = FALSE}

years_to_include_rst_data <- exclusion_catch |> 
  filter(exclude_60 == F) |> 
  select(stream, site, run_year)

knitr::kable(head(years_to_include_rst_data, 10))

usethis::use_data(years_to_include_rst_data, overwrite = TRUE)
```

*... with `r nrow(years_to_include_rst_data) - 10` more rows*

## Adult Data - Years to Include in Model

We treated adult data a little differently to account for the two main types of adult data - adult survey data (holding, redd, carcass) and adult passage data (video passage).

Below are the methods we used for excluding years by adult data type:

**Survey data**

-   Exclude year if survey does not cover core reaches in a year or if less than 50% of reaches are sampled. 

    | Stream         | Core Reaches |
    |----------------|-------------------------------------------------------------------------------------------|
    | Battle Creek   | R1, R2, R3, R4 (R5, R6) |
    | Butte Creek    | A1, B1, C1, D1, E3          |
    | Clear Creek    | R1, R2, R3, R4, R5 (R5A, R5B, R5C) |
    | Deer Creek     | Lower Falls to A line, A line to Wilson Cove, Polk Springs to Murphy Trail, Murphy Trail to Ponderosa Way, Ponderosa Way to Trail 2E17, Trail 2E17 to Dillon Cove, Uper Falls to Potato Patch Camp, Potato Patch Camp to Highway 32 (Red Bridge)         |
    | Feather River  | 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38  |
    | Mill Creek     | Mccarthy Place To Savercool Place, Savercool Place To Black Rock, Black Rock To Below Ranch House, Below Ranch House To Above Avery, Above Avery To Pape Place, Pape Place to Buckhorn Gulch, Buckhorn Gulch To Upper Dam, Above Hwy 36,Hwy 36 To Little Hole In Ground, Hole In Ground To Ishi Trail Head, Ishi Trail Head To Big Bend, Big Bend to Canyon Camp, Canyon Camp To Sooner Place  |
    | Yuba River     | Yuba does not have core reaches with historical survyes, only samples section of river, not appropriate for use in JPE|

**Video data**

-   Exclude year if video is out for more than 4 weeks in the sampling season
-   Exclude year if flows exceed a threshold value on each tributary (overpass weir etc..). Threshold determined by monitoring program. 

We used these rules as preliminary way to identify potentially incomplete years and then conducted outreach to stream teams to review our list and highlight any other years that should be excluded or not.

### Years to exclude

We utilized above cumulative catch curves and heatmaps to come up with a list of years to exclude from modeling for adult data. See a section of the "years to exclude" table below.

```{r, echo = FALSE}
usethis::use_data(years_to_exclude_adult, overwrite = TRUE)
years_exclude_nice_names_adult <- years_to_exclude_adult |> select("Stream" = stream, 
                                                    "Year" = year, 
                                                    "Data Type" = data_type,
                                                    "Exclusion Type" = reason_for_exclusion)
knitr::kable(head(years_exclude_nice_names_adult, 10))
```

*... with `r nrow(years_exclude_nice_names_adult) - 10` more rows*

### Applying to Modeling Datasets

In order to apply the years to exclude information to the modeling datasets we did some additional processing to create a table describing the stream, year by data type that should be included in the SR JPE modeling.

The table below shows a section of this table:

```{r, echo = FALSE}
# years to exclude already applied to pull_adult_data.R
years_to_include_adult <- annual_adult |> 
  mutate(data_type = case_when(data_type == "upstream_estimate" ~ "upstream passage",
                               data_type == "carcass_estimate" ~ "carcass",
                               T ~ data_type))

knitr::kable(head(years_to_include_adult, 10))

usethis::use_data(years_to_include_adult, overwrite = TRUE)
```

*... with `r nrow(years_to_include_adult) - 10` more rows*
