---
title: "Years to Include"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Years to Include}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE, 
  message = FALSE,
  comment = "#>", 
  fig.width=8, fig.height=5)

if (requireNamespace("devtools", quietly = TRUE)) devtools::load_all(".", quiet = TRUE)
library(SRJPEdata)

library(tidyverse)
library(lubridate)

colors_full <-  c("#9A8822", "#F5CDB4", "#F8AFA8", "#FDDDA0", "#74A089", #Royal 2
                  "#899DA4", "#C93312", "#DC863B", # royal 1 (- 3)
                  "#F1BB7B", "#FD6467", "#5B1A18", # Grand Budapest 1 (-4)
                  "#D8B70A", "#02401B", "#A2A475", # Cavalcanti 1
                  "#E6A0C4", "#C6CDF7", "#D8A499", "#7294D4", #Grand Budapest 2
                  "#9986A5", "#EAD3BF", "#AA9486", "#B6854D", "#798E87" # Isle of dogs 2 altered slightly
)
years_to_exclude_adult <- read_csv(here::here("data-raw", "helper-tables", "years_to_exclude_adult_datasets.csv"))  
```

The Spring Run Juvenile Production Estimate uses historical rotary screw trap and adult survey data. In order to improve model performance the modeling team worked with the stream teams to filter out data that is too incomplete to use for modeling. This article describes the process for selecting RST years to include and adult years to include.

Historically this process differed for each tributary and was based on the rules below. This ruleset was complicated and difficult to maintain. In April 2026 we decided to transition to a consistent ruleset to apply across all years and streams. We compared a number of thresholds and determined that a rule where at least 60% of the weeks (45:53, 1:22) were sampled was appropriate - it was not too stringent and did not include too many gaps. In addition, any sites that do not have any mark-recapture efficiency trials are removed.

## RST Data - Historical Ruleset

The modeling team hosted a modeling windows workshop where we discussed approaches to defining the time window that should be included in the juvenile abundance model. At the workshop we decided to use the full sampling season of data for each tributary, site, and year but to exclude years where we were concerned about data completeness. We came up with the following approach to determine excluded years:

1)  First exclude years with lots of missing data from cumulative catch curves. See `data-raw/years_to_exclude.csv` (exclusion_type = "really low sampling").
2)  Exclude yearlings from cumulative catch curves.
3)  Use updated cumulative catch curves to determine the critical window (average window over all historical years that captures 75% percent of catch).
4)  Remove additional years where there is no sampling for 4 consecutive weeks within the critical window. See `data-raw/years_to_exclude.csv` (exclusion_type = "missing four consecutive weeks in critical window").

### Heat map of all tributaries

The heat map below shows when sampling occurs for all streams. It shows that sampling is more complete across traps from 2004 - 2009 and from mid November to July. Some traps have continued sampling throughout the season. Some years there are gaps within season for specific traps or across multiple traps.

```{r, echo = FALSE}
# Need to join catch_data_week to make sure all weeks get included in the 
# sampling window heatmap
catch_date_week <- SRJPEdata::rst_catch %>% 
  mutate(day = day(date),
         month = month(date),
         fake_date = as_date(paste0("2000-", month, "-", day))) %>%
  select(fake_date) %>%
  distinct() %>%
  mutate(week = week(fake_date))
# This helper table provides a date range for each numeric week
# This is used as hover text label in the heatmap
date_range <- catch_date_week %>%
  group_by(week) %>%
  slice_min(fake_date) %>%
  rename(min_date = fake_date) %>%
  left_join(catch_date_week %>%
              group_by(week) %>%
              slice_max(fake_date) %>%
              rename(max_date = fake_date)) %>%
  mutate(min_date = format(min_date, "%b %d"),
         max_date = format(max_date, "%b %d"),
         date_range = paste(min_date,"-",max_date))

catch_summary <- SRJPEdata::rst_catch %>%
  select(date, stream) %>%
  distinct() %>% 
  mutate(week = week(date),
         year = year(date)) %>%
  group_by(stream, week, year) %>%
  tally()

catch_weekly <- catch_summary %>% 
  # add catch_date_week to make sure all weeks are included in the data
  right_join(select(catch_date_week, week) %>% 
               distinct()) %>% 
  pivot_wider(id_cols = c("stream", "year"), 
              names_from = week, 
              values_from = n, 
              values_fill = 0) %>%
  pivot_longer(cols = -c("stream","year"), names_to = "week", values_to = "number_days_sampled") %>%
  mutate(week = as.numeric(week)) %>%
  left_join(date_range) %>%
  mutate(fake_date = case_when(week %in% c(40:53) ~ ymd("1999-01-01") + weeks(week - 1),
                               T ~ ymd("2000-01-01") + weeks(week - 1)))

plot_data <- catch_weekly %>%
      filter(number_days_sampled > 0) %>%
      mutate(wy = ifelse(week >= 37, year + 1, year)) %>%
      group_by(week, wy, date_range, fake_date) %>%
      summarize(streams = paste(unique(stream), collapse = ", "),
                n_streams = length(unique(stream))) %>%
      ungroup() %>%
      mutate(week = factor(week, levels = c(37:53, 1:36)),
             wy = factor(wy))
year_labels <- c(1995, 2000, 2005, 2010, 2015, 2020)

heat_map <- ggplot(plot_data, aes(x = fake_date, y = wy, fill = n_streams, text = paste0(date_range, "<br>", streams))) +
  geom_tile() +
  #scale_y_continuous(n.breaks = 5, breaks = waiver()) +
  scale_y_discrete(breaks = year_labels) +
  scale_x_date(date_labels = "%b", breaks = "2 months") +
  theme_minimal() + 
  ylab('water year') +
  xlab("") + 
  # scale_fill_viridis_b(option = "A", n.breaks = 8, direction = -1) +
  labs(fill = "Number of streams")
plotly::ggplotly(heat_map) 
# heat_map
```

## RST Data - Current Ruleset

Ruleset defined as "at least 19 weeks included within a run year (week 45-53 and 1-22) for a given site" or 60% of weeks sampled. Sites that do not have any (across all years) mark-repcature efficiency trials are also excluded.

```{r, echo = FALSE}
# Find sites with efficiency data. Any site that has NO efficiency trials will be excluded
stream_site_with_efficiency_data <- SRJPEdata::weekly_efficiency |>
  distinct(stream, site) |>
  mutate(if_efficiency_data = T)

# Assign stream/site/year as exclude or not based on criteria
exclusion_catch <- SRJPEdata::weekly_juvenile_abundance_catch_data |>
  filter(!is.na(count)) |> # remove when trap is not fishing
  filter(week %in% c(45:53, 1:22)) |> # select to the week window used in BTSPASX
  select(stream, site, week, run_year) |>
  distinct() |>
  group_by(stream, site, run_year) |>
  summarise(number_weeks = n()) |>
  # other exclusion rules are included in case this changes in the future
  mutate(
    exclude_20 = ifelse(number_weeks < 6, TRUE, FALSE),
    exclude_40 = ifelse(number_weeks < 13, TRUE, FALSE),
    exclude_60 = ifelse(number_weeks < 19, TRUE, FALSE),
    exclude_65 = ifelse(number_weeks < 21, TRUE, FALSE),
    exclude_70 = ifelse(number_weeks < 22, TRUE, FALSE),
    exclude_75 = ifelse(number_weeks < 24, TRUE, FALSE)
  ) |> 
  left_join(stream_site_with_efficiency_data) |> 
  mutate(if_efficiency_data = if_else(is.na(if_efficiency_data), F, T)) 

years_to_exclude_rst_data_all <- exclusion_catch |>
  filter(exclude_60 == T | if_efficiency_data == F) |> 
  select(stream, site, run_year, number_weeks) |>
  mutate(exclusion_type = "60% of weeks sampled", apply_to = "all runs")

years_exclude_nice_names <- years_to_exclude_rst_data_all |>
  select(
    "Stream" = stream,
    "Site" = site,
    "Run Year" = run_year,
    "Exclusion Type" = exclusion_type
  )
knitr::kable(head(years_exclude_nice_names, 10))
```

*... with `r nrow(years_to_exclude_rst_data) - 10` more rows*

### Spring run specific ruleset

Although the focus of this effort is on spring run Chinook salmon, the SR JPE modeling was designed to be applicable for modeling all runs as PLAD run designations are available for all runs. The "years to exclude" ruleset is broadly applicable, however, there are some specific cases where there may be site/years beyond those identified in the "years to exclude ruleset" that are not expected for perform well, specifically for spring run. This may happen when the number of weeks sampling is met but there is a large gap in sampling during the peak of spring run outmigration. We developed an additional list to track these cases and apply to modeling.

```{r, echo = FALSE}
years_to_exclude_rst_data <- years_to_exclude_rst_data_all |>
  ungroup() |>
  add_row(
    stream = "sacramento river",
    site = "tisdale",
    run_year = 2013,
    number_weeks = 20,
    exclusion_type = "spring run specific - 11 consecutive weeks not sampled Dec 16 to Feb 19",
    apply_to = "spring"
  ) |>
  add_row(
    stream = "feather river",
    site = "herringer riffle",
    run_year = 2008,
    number_weeks = 21,
    exclusion_type = "spring run specific - 9 weeks missing Nov 04 to Jan 01",
    apply_to = "spring"
  )

knitr::kable(years_to_exclude_rst_data |> filter(grepl("spring run specific", exclusion_type)))
```

### years_to_exclude_rst_data object

We combine the "years to exclude" ruleset and the spring run specific ruleset into one data object `years_to_exclude_rst_data` and add a column called `apply_to`. This column will be `all runs` for the broad ruleset or could be specific to a particular run (e.g. will be `spring` for spring run specific cases)

```{r}
usethis::use_data(years_to_exclude_rst_data, overwrite = TRUE)
```

### years_to_include_rst_data object

In order to apply the years to exclude information to the modeling datasets we did some additional processing to create a table describing the stream, site, year that should be included in the SR JPE modeling. There are two versions of this table depending on if a list of site/years for all runs or spring run specific are needed.

- `years_to_include_rst_data` use when want to apply ruleset broadly, applicable for all runs
- `years_to_include_rst_data_spring` use when want to apply ruleset specifically to spring run

The table below shows a section of this table applicable to all runs:

```{r, echo = FALSE}
years_to_include_rst_data <- exclusion_catch |>
  filter(exclude_60 == F & if_efficiency_data == T) |>
  select(stream, site, run_year)

knitr::kable(head(years_to_include_rst_data, 10))

usethis::use_data(years_to_include_rst_data, overwrite = TRUE)

years_to_include_rst_data_spring <- years_to_include_rst_data |>
  left_join(
    years_to_exclude_rst_data |> 
      filter(apply_to == "spring") # flag the site/years that need to be removed
  ) |>
  filter(is.na(apply_to)) |> # remove the spring specific site years
  select(stream, site, run_year)

usethis::use_data(years_to_include_rst_data_spring, overwrite = TRUE)
```

*... with `r nrow(years_to_include_rst_data_spring) - 10` more rows*

## Adult Data - Years to Include in Model

We treated adult data a little differently to account for the two main types of adult data - adult survey data (holding, redd, carcass) and adult passage data (video passage).

Below are the methods we used for excluding years by adult data type:

**Survey data**

-   Exclude year if survey does not cover core reaches in a year or if less than 50% of reaches are sampled. 

    | Stream         | Core Reaches |
    |----------------|-------------------------------------------------------------------------------------------|
    | Battle Creek   | R1, R2, R3, R4 (R5, R6) |
    | Butte Creek    | A1, B1, C1, D1, E3          |
    | Clear Creek    | R1, R2, R3, R4, R5 (R5A, R5B, R5C) |
    | Deer Creek     | Lower Falls to A line, A line to Wilson Cove, Polk Springs to Murphy Trail, Murphy Trail to Ponderosa Way, Ponderosa Way to Trail 2E17, Trail 2E17 to Dillon Cove, Uper Falls to Potato Patch Camp, Potato Patch Camp to Highway 32 (Red Bridge)         |
    | Feather River  | 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38  |
    | Mill Creek     | Mccarthy Place To Savercool Place, Savercool Place To Black Rock, Black Rock To Below Ranch House, Below Ranch House To Above Avery, Above Avery To Pape Place, Pape Place to Buckhorn Gulch, Buckhorn Gulch To Upper Dam, Above Hwy 36,Hwy 36 To Little Hole In Ground, Hole In Ground To Ishi Trail Head, Ishi Trail Head To Big Bend, Big Bend to Canyon Camp, Canyon Camp To Sooner Place  |
    | Yuba River     | Yuba does not have core reaches with historical survyes, only samples section of river, not appropriate for use in JPE|

**Video data**

-   Exclude year if video is out for more than 4 weeks in the sampling season
-   Exclude year if flows exceed a threshold value on each tributary (overpass weir etc..). Threshold determined by monitoring program. 

We used these rules as preliminary way to identify potentially incomplete years and then conducted outreach to stream teams to review our list and highlight any other years that should be excluded or not.

### Years to exclude

We utilized above cumulative catch curves and heatmaps to come up with a list of years to exclude from modeling for adult data. See a section of the "years to exclude" table below.

```{r, echo = FALSE}
usethis::use_data(years_to_exclude_adult, overwrite = TRUE)
years_exclude_nice_names_adult <- years_to_exclude_adult |> select("Stream" = stream, 
                                                    "Year" = year, 
                                                    "Data Type" = data_type,
                                                    "Exclusion Type" = reason_for_exclusion)
knitr::kable(head(years_exclude_nice_names_adult, 10))
```

*... with `r nrow(years_exclude_nice_names_adult) - 10` more rows*

### Applying to Modeling Datasets

In order to apply the years to exclude information to the modeling datasets we did some additional processing to create a table describing the stream, year by data type that should be included in the SR JPE modeling.

The table below shows a section of this table:

```{r, echo = FALSE}
# years to exclude already applied to pull_adult_data.R
years_to_include_adult <- annual_adult |> 
  mutate(data_type = case_when(data_type == "upstream_estimate" ~ "upstream passage",
                               data_type == "carcass_estimate" ~ "carcass",
                               T ~ data_type))

knitr::kable(head(years_to_include_adult, 10))

usethis::use_data(years_to_include_adult, overwrite = TRUE)
```

*... with `r nrow(years_to_include_adult) - 10` more rows*
