---
title: "Water Temperature Regression Analysis"
output: 
  html_document:
  theme: flatly
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include=FALSE}
library(tidyverse)
library(googleCloudStorageR)
library(zoo)
library(CDECRetrieve)
library(lubridate)
library(weathermetrics)

root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
```


# Regression Analysis for Water Temperature Data

The goal of this analysis is to develop a regression to predict water temperature when data are missing or the time series is incomplete.

Currently this analysis is performed for Feather River and Yuba River. The resulting dataset with predicted values are saved and integrated in the development of a water temperature dataset.


## Data used to build models: Butte Creek

Butte Creek is used to build the regression models because the time series is complete and the data are high quality.
  * Date range covered is 1999 - 2024

```{r, include = F}
butte <- cdec_query(station = "BCK", dur_code = "H", sensor_num = "25", start_date = "2000-01-01")
butte_format <- butte |> 
    mutate(date = as_date(datetime),
           temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) |>
    filter(temp_degC < 40, temp_degC > 0.5) |>
    group_by(date) |> 
    summarise(mean = mean(temp_degC, na.rm = TRUE),
              max = max(temp_degC, na.rm = TRUE),
              min = min(temp_degC, na.rm = TRUE)) |> 
    pivot_longer(mean:min, names_to = "statistic", values_to = "value") |>
    mutate(stream = "butte creek",
           gage_agency = "CDEC",
           gage_number = "BCK",
           parameter = "temperature")

```


## Feather River

### Data Preparation and Approach

1. Pull in gage data from CDEC (GRL will represent the High Flow Channel (HFC) and FRA will represent the Low Flow Channel (LFC):
* GRL (2003-03-05 to 2007-06-01 H; 2020-01-04 to present E): located after Thermalito Afterbay
* FRA (2002-01-01 to present): located between Lake Oroville and Thermalito Afterbay

2. Prepare datasets for regression analysis:
* Dataset with no missing data to train (Butte Creek)
* Dataset with missing data to predict (Feather River)

3. Combine datasets with no missing data, and missing data

4. Identify gaps to predict

5. Use data where there are no missing data for either dataset for regression modeling
 

```{r, include = F}
# Pull in Gage Data that were extracted in the "data-raw/qc-markdowns/temperature_data_prep.Rmd"
hfc_feather <- cdec_query(station = "GRL", dur_code = "H", sensor_num = "25", start_date = "2003-03-05", end_date = "2007-06-01")
feather_hfc_format <- hfc_feather |> 
         mutate(date = as_date(datetime),
                temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) |>
         filter(temp_degC < 40, temp_degC > 0) |>
         group_by(date) |> 
         summarise(mean = mean(temp_degC, na.rm = TRUE),
                   max = max(temp_degC, na.rm = TRUE),
                   min = min(temp_degC, na.rm = TRUE)) |> 
         pivot_longer(mean:min, names_to = "statistic", values_to = "value") |>
         mutate(stream = "feather river",
                gage_agency = "CDEC",
                gage_number = "GRL",
                site_group = "upper feather hfc",
                parameter = "temperature") |> 
  glimpse()

lfc_feather <- cdec_query(station = "FRA", dur_code = "H", sensor_num = "25", start_date = "1996-01-01")
feather_lfc_format <- lfc_feather |> 
         mutate(date = as_date(datetime),
                temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) |>
         filter(temp_degC < 40, temp_degC > 0) |>
         group_by(date) |> 
         summarise(mean = mean(temp_degC, na.rm = TRUE),
                   max = max(temp_degC, na.rm = TRUE),
                   min = min(temp_degC, na.rm = TRUE)) |> 
         pivot_longer(mean:min, names_to = "statistic", values_to = "value") |>
         mutate(stream = "feather river",
                gage_agency = "CDEC",
                gage_number = "FRA",
                site_group = "upper feather lfc",
                parameter = "temperature")
```

### LFC {.tabset}

```{r, include = F}
# Prepare data for regression
feather_lfc_format_wide <- feather_lfc_format |> 
  ungroup() |> 
  select(-gage_agency, -gage_number) |> 
  pivot_wider(id_cols = c(stream, date, site_group), names_from = "statistic", values_from = "value", values_fill = NA)

butte_format_wide <- butte_format |> 
  ungroup() |> 
  select(-gage_agency, -gage_number) |> 
  pivot_wider(id_cols = c(stream, date), names_from = "statistic", values_from = "value", values_fill = NA)
# Combine butte and feather together for regression modeling
feather_lfc_regression_data_full <- feather_lfc_format_wide |> 
  ungroup() |> 
  select(date, mean, max, min) |> 
  rename(feather_temp_mean = mean,
         feather_temp_min = min,
         feather_temp_max = max) |> 
  full_join(butte_format_wide |> 
              select(date, mean, max, min) |> 
              rename(butte_temp_mean = mean,
                     butte_temp_min = min,
                     butte_temp_max = max))

# FOR PREDICTIONS identify gaps to predict data
feather_gap_mean <- feather_lfc_regression_data_full |> 
  filter(is.na(feather_temp_mean), !is.na(butte_temp_mean)) |> 
  select(-c(feather_temp_mean, feather_temp_min, feather_temp_max)) |> 
  rename(butte_temp = butte_temp_mean)
feather_gap_min <- feather_lfc_regression_data_full |> 
  filter(is.na(feather_temp_min), !is.na(butte_temp_min)) |> 
 select(-c(feather_temp_mean, feather_temp_min, feather_temp_max)) |> 
  rename(butte_temp = butte_temp_min)
feather_gap_max <- feather_lfc_regression_data_full |> 
  filter(is.na(feather_temp_max), !is.na(butte_temp_max)) |> 
 select(-c(feather_temp_mean, feather_temp_min, feather_temp_max)) |> 
  rename(butte_temp = butte_temp_max)

# FOR MODEL use data where there are no missing data for either butte or feather for regression modeling
feather_lfc_regression_data_mean <- feather_lfc_regression_data_full |> 
  filter(!is.na(feather_temp_mean), !is.na(butte_temp_mean)) |> 
  rename(butte_temp = butte_temp_mean,
         temp = feather_temp_mean)
feather_lfc_regression_data_min <- feather_lfc_regression_data_full |> 
  filter(!is.na(feather_temp_min), !is.na(butte_temp_min)) |> 
  rename(butte_temp = butte_temp_min,
         temp = feather_temp_min)
feather_lfc_regression_data_max <- feather_lfc_regression_data_full |> 
  filter(!is.na(feather_temp_max), !is.na(butte_temp_max)) |> 
  rename(butte_temp = butte_temp_max,
         temp = feather_temp_max)
```

*Plot of mean temp for Feather River LFC and Butte Creek*
```{r, echo = F}
# plot for mean
ggplot(data = feather_lfc_regression_data_mean, aes(x = butte_temp, y = temp)) +
  geom_point() +
  stat_smooth(method = "lm")
```

*Plot of min temp for Feather River LFC and Butte Creek*
```{r, echo = F}
# plot for min
ggplot(data = feather_lfc_regression_data_min, aes(x = butte_temp, y = temp)) +
  geom_point() +
  stat_smooth(method = "lm")
```

*Plot of max temp for Feather River LFC and Butte Creek*
```{r, echo = F}
# plot for max
ggplot(data = feather_lfc_regression_data_max, aes(x = butte_temp, y = temp)) +
  geom_point() +
  stat_smooth(method = "lm")
```

#### Building Mean Regression
```{r}
# LFC regression and predictions

# MEAN Regression
split <-rsample::initial_split(feather_lfc_regression_data_mean, prop = 0.8)
train <- rsample::training(split)
test <- rsample::testing(split)
feather_lfc_mod_mean <- lm(temp ~ date + butte_temp, data = train)
summary(feather_lfc_mod_mean)
test_predict <- predict(feather_lfc_mod_mean, test)
test_predict_df <- test |>
  mutate(predicted = test_predict)
# evaluate predictions - MAPE of 10% is not bad
mean(abs((
  test_predict_df$predicted - test_predict_df$temp
)) / test_predict_df$temp)
```

#### Building Min Regression
```{r}
# MIN Regression
split <-rsample::initial_split(feather_lfc_regression_data_min, prop = 0.8)
train <- rsample::training(split)
test <- rsample::testing(split)
feather_lfc_mod_min <- lm(temp ~ date + butte_temp, data = train)
summary(feather_lfc_mod_min)
test_predict <- predict(feather_lfc_mod_min, test)
test_predict_df <- test |>
  mutate(predicted = test_predict)
# evaluate predictions - MAPE of 10% is not bad
mean(abs((
  test_predict_df$predicted - test_predict_df$temp
)) / test_predict_df$temp)
```

#### Building Max Regression
```{r}
# MAX Regression
split <-rsample::initial_split(feather_lfc_regression_data_max, prop = 0.8)
train <- rsample::training(split)
test <- rsample::testing(split)
feather_lfc_mod_max <- lm(temp ~ date + butte_temp, data = train)
summary(feather_lfc_mod_max)
test_predict <- predict(feather_lfc_mod_max, test)
test_predict_df <- test |>
  mutate(predicted = test_predict)
# evaluate predictions - MAPE of 10% is not bad
mean(abs((
  test_predict_df$predicted - test_predict_df$temp
)) / test_predict_df$temp)
```

#### Predictions

```{r}
# Predictions
feather_gap_predicted_mean_lfc <- predict(feather_lfc_mod_mean, feather_gap_mean)
feather_lfc_mean_predicted <- feather_gap_mean |> 
  mutate(value = feather_gap_predicted_mean_lfc,
         statistic = "mean") |> 
  select(date, value, statistic)
ggplot(feather_lfc_mean_predicted, aes(x = date, y = value)) +
  geom_line()
feather_gap_predicted_min_lfc <- predict(feather_lfc_mod_min, feather_gap_min)
feather_lfc_min_predicted <- feather_gap_min |> 
  mutate(value = feather_gap_predicted_min_lfc,
         statistic = "min") |> 
  select(date, value, statistic)
ggplot(feather_lfc_min_predicted, aes(x = date, y = value)) +
  geom_line()
feather_gap_predicted_max_lfc <- predict(feather_lfc_mod_max, feather_gap_max)
feather_lfc_max_predicted <- feather_gap_max |> 
  mutate(value = feather_gap_predicted_max_lfc,
         statistic = "max") |> 
  select(date, value, statistic)
ggplot(feather_lfc_max_predicted, aes(x = date, y = value)) +
  geom_line()
```

#### Full dataset

```{r}
feather_gap_lfc <- bind_rows(feather_lfc_max_predicted,
                             feather_lfc_mean_predicted,
                             feather_lfc_min_predicted) |> 
  mutate(stream = "feather river",
         site_group = "upper feather lfc") %>% 
  pivot_wider(id_cols = c(date, stream, site_group), values_from = "value", names_from = "statistic") |>
  rename(mean_i = mean,
         max_i = max,
         min_i = min) |> 
  full_join(feather_lfc_format |> 
              pivot_wider(id_cols = c(stream, date, gage_agency, gage_number, site_group), values_from = "value", names_from = "statistic")) |> 
  mutate(gage_agency = ifelse(is.na(mean), "interpolated", gage_agency),
         gage_number = ifelse(is.na(mean), "interpolated", gage_number),
         mean = ifelse(is.na(mean), mean_i, mean),
         min = ifelse(is.na(min), min_i, min),
         max = ifelse(is.na(max), max_i, max)) |>
  select(-c(min_i, mean_i, max_i)) |> 
  pivot_longer(cols = mean:min, values_to = "value", names_to = "statistic") |> 
  group_by(stream, date, statistic, gage_agency, gage_number, site_group) |> glimpse()

ggplot() +
  geom_line(data = feather_gap_lfc, aes(x = date, y = value, color = statistic)) +
  theme_minimal()
```
```{r, include = F}
#write csv
write_csv(feather_gap_lfc, here::here("data-raw", "temperature-data", "feather_lfc_temp_interpolation.csv"))
```


### HFC {.tabset}

```{r, include = F}
# need reformat the dataset so can find the dates when min/max are NA (e.g. otherwise just won't ahve data)
feather_hfc_format_wide <- feather_hfc_format |> 
  ungroup() |> 
  select(-gage_agency, -gage_number) |> 
pivot_wider(id_cols = c(stream, date, site_group), names_from = "statistic", values_from = "value", values_fill = NA)
# set up dataframes for HFC
feather_hfc_regression_data_full <- feather_hfc_format_wide |> 
  ungroup() |> 
  select(date, mean, max, min) |> 
  rename(feather_temp_mean = mean,
         feather_temp_min = min,
         feather_temp_max = max) |> 
  full_join(butte_format_wide |> 
              select(date, mean, max, min) |> 
              rename(butte_temp_mean = mean,
                     butte_temp_min = min,
                     butte_temp_max = max))
# identify gaps to predict data
feather_gap_mean <- feather_hfc_regression_data_full |> 
  filter(is.na(feather_temp_mean), !is.na(butte_temp_mean)) |> 
  select(-c(feather_temp_mean, feather_temp_min, feather_temp_max)) |> 
  rename(butte_temp = butte_temp_mean)
feather_gap_min <- feather_hfc_regression_data_full |> 
  filter(is.na(feather_temp_min), !is.na(butte_temp_min)) |> 
 select(-c(feather_temp_mean, feather_temp_min, feather_temp_max)) |> 
  rename(butte_temp = butte_temp_min)
feather_gap_max <- feather_hfc_regression_data_full |> 
  filter(is.na(feather_temp_max), !is.na(butte_temp_max)) |> 
 select(-c(feather_temp_mean, feather_temp_min, feather_temp_max)) |> 
  rename(butte_temp = butte_temp_max)

feather_hfc_regression_data_mean <- feather_hfc_regression_data_full |> 
  filter(!is.na(feather_temp_mean), !is.na(butte_temp_mean)) |> 
  rename(butte_temp = butte_temp_mean,
         temp = feather_temp_mean)
feather_hfc_regression_data_min <- feather_hfc_regression_data_full |> 
  filter(!is.na(feather_temp_min), !is.na(butte_temp_min)) |> 
  rename(butte_temp = butte_temp_min,
         temp = feather_temp_min)
feather_hfc_regression_data_max <- feather_hfc_regression_data_full |> 
  filter(!is.na(feather_temp_max), !is.na(butte_temp_max)) |> 
  rename(butte_temp = butte_temp_max,
         temp = feather_temp_max)
```

*Plot of mean temp for Feather River HFC and Butte Creek*
```{r}
# plot for mean
ggplot(data = feather_hfc_regression_data_mean, aes(x = butte_temp, y = temp)) +
  geom_point() +
  stat_smooth(method = "lm")
```

*Plot of min temp for Feather River HFC and Butte Creek*
```{r}
# plot for mean
ggplot(data = feather_hfc_regression_data_min, aes(x = butte_temp, y = temp)) +
  geom_point() +
  stat_smooth(method = "lm")
```

*Plot of max temp for Feather River HFC and Butte Creek*
```{r}
# plot for mean
ggplot(data = feather_hfc_regression_data_max, aes(x = butte_temp, y = temp)) +
  geom_point() +
  stat_smooth(method = "lm")
```

#### Building Mean Regression
```{r}
# HFC regression and predictions

# MEAN Regression
split <-rsample::initial_split(feather_hfc_regression_data_mean, prop = 0.8)
train <- rsample::training(split)
test <- rsample::testing(split)
feather_hfc_mod_mean <- lm(temp ~ date + butte_temp, data = train)
summary(feather_hfc_mod_mean)
test_predict <- predict(feather_hfc_mod_mean, test)
test_predict_df <- test |>
  mutate(predicted = test_predict)
# evaluate predictions - MAPE of 10% is not bad
mean(abs((
  test_predict_df$predicted - test_predict_df$temp
)) / test_predict_df$temp)
```

#### Building Min Regression
```{r}
# MIN Regression
split <-rsample::initial_split(feather_hfc_regression_data_min, prop = 0.8)
train <- rsample::training(split)
test <- rsample::testing(split)
feather_hfc_mod_min <- lm(temp ~ date + butte_temp, data = train)
summary(feather_hfc_mod_min)
test_predict <- predict(feather_hfc_mod_min, test)
test_predict_df <- test |>
  mutate(predicted = test_predict)
# evaluate predictions - MAPE of 10% is not bad
mean(abs((
  test_predict_df$predicted - test_predict_df$temp
)) / test_predict_df$temp)
```

#### Building Max Regression
```{r}
# MAX Regression
split <-rsample::initial_split(feather_hfc_regression_data_max, prop = 0.8)
train <- rsample::training(split)
test <- rsample::testing(split)
feather_hfc_mod_max <- lm(temp ~ date + butte_temp, data = train)
summary(feather_hfc_mod_max)
test_predict <- predict(feather_hfc_mod_max, test)
test_predict_df <- test |>
  mutate(predicted = test_predict)
# evaluate predictions - MAPE of 10% is not bad
mean(abs((
  test_predict_df$predicted - test_predict_df$temp
)) / test_predict_df$temp)
```

#### Predictions
```{r}
# Predictions
feather_gap_predicted_mean_hfc <- predict(feather_hfc_mod_mean, feather_gap_mean)
feather_hfc_mean_predicted <- feather_gap_mean |> 
  mutate(value = feather_gap_predicted_mean_hfc,
         statistic = "mean") |> 
  select(date, value, statistic)
ggplot(feather_hfc_mean_predicted, aes(x = date, y = value)) +
  geom_line()
feather_gap_predicted_min_hfc <- predict(feather_hfc_mod_min, feather_gap_min)
feather_hfc_min_predicted <- feather_gap_min |> 
  mutate(value = feather_gap_predicted_min_hfc,
         statistic = "min") |> 
  select(date, value, statistic)
ggplot(feather_hfc_min_predicted, aes(x = date, y = value)) +
  geom_line()
feather_gap_predicted_max_hfc <- predict(feather_hfc_mod_max, feather_gap_max)
feather_hfc_max_predicted <- feather_gap_max |> 
  mutate(value = feather_gap_predicted_max_hfc,
         statistic = "max") |> 
  select(date, value, statistic)
ggplot(feather_hfc_max_predicted, aes(x = date, y = value)) +
  geom_line()
```

#### Full dataset

```{r}
feather_gap_hfc <- bind_rows(feather_hfc_max_predicted,
                             feather_hfc_mean_predicted,
                             feather_hfc_min_predicted) |> 
  mutate(stream = "feather river",
         site_group = "upper feather hfc") %>% 
  pivot_wider(id_cols = c(date, stream, site_group), values_from = "value", names_from = "statistic") |>
  rename(mean_i = mean,
         max_i = max,
         min_i = min) |> 
  full_join(feather_hfc_format |> 
              pivot_wider(id_cols = c(stream, date, gage_agency, gage_number, site_group), values_from = "value", names_from = "statistic")) |> 
  mutate(gage_agency = ifelse(is.na(mean), "interpolated", gage_agency),
         gage_number = ifelse(is.na(mean), "interpolated", gage_number),
         mean = ifelse(is.na(mean), mean_i, mean),
         min = ifelse(is.na(min), min_i, min),
         max = ifelse(is.na(max), max_i, max)) |> 
  select(-c(min_i, mean_i, max_i)) |> 
  pivot_longer(cols = mean:min, values_to = "value", names_to = "statistic") |> 
  group_by(stream, date, statistic, gage_agency, gage_number, site_group) |> glimpse()

ggplot() +
  geom_line(data = feather_gap_hfc, aes(x = date, y = value, color = statistic)) +
  theme_minimal()
```

```{r, include = F}
# write csv
write_csv(feather_gap_hfc, here::here("data-raw", "temperature-data", "feather_hfc_temp_interpolation.csv"))

```

## Yuba River

### Data Prepartion and Approach {.tabset}

1. Pull in gage data from YR7 CDEC gage, however, this only contains data from 2020 onwards. Temperature data collected along with RST data was originally used to fill the gap prior to 2020; however, due to inconsistencies in these data sources the resulting predicted mean values were lower than the min values as the RST data only has mean data. 

2. Prepare datasets for regression analysis
* Dataset with no missing data to train (Butte Creek)
* Dataset with missing data to predict (Yuba River)

3. Combine datasets with no missing data, and missing data

4. Identify gaps to predict

5. Use data where there are no missing data for either dataset for regression modeling

```{r, include = F}
# Gage Data
YR7_daily_temps <- cdec_query(station = "YR7", dur_code = "E", sensor_num = "146", start_date = "2019-01-01")

yuba_format <- YR7_daily_temps |> 
    mutate(date = as_date(datetime),
           year = year(datetime)) |> 
    filter(parameter_value < 40, parameter_value > 0, !is.na(date)) |> 
    group_by(date) |> 
    summarise(mean = mean(parameter_value, na.rm = TRUE),
             max = max(parameter_value, na.rm = TRUE),
             min = min(parameter_value, na.rm = TRUE)) |> 
  pivot_longer(mean:min, names_to = "statistic", values_to = "value") |>
  mutate(stream = "yuba river",
         gage_agency = "CDEC",
         gage_number = "YR7",
         parameter = "temperature") |> 
  glimpse()
```

```{r, include = F}
# combine butte and yuba together for regression modeling
yuba_format_wide <- yuba_format |> 
  ungroup() |> 
  select(-gage_agency, -gage_number) |> 
  pivot_wider(id_cols = c(stream, date), names_from = "statistic", values_from = "value", values_fill = NA)

yuba_regression_data_full <- yuba_format_wide |> 
  ungroup() |> 
  select(date, mean, max, min) |> 
  rename(yuba_temp_mean = mean,
         yuba_temp_min = min,
         yuba_temp_max = max) |> 
  full_join(butte_format_wide |> 
              select(date, mean, max, min) |> 
              rename(butte_temp_mean = mean,
                     butte_temp_min = min,
                     butte_temp_max = max))
# FOR PREDICTIONS identify gaps to predict data
yuba_gap_mean <- yuba_regression_data_full |> 
  filter(is.na(yuba_temp_mean), !is.na(butte_temp_mean)) |> 
  select(-c(yuba_temp_mean, yuba_temp_min, yuba_temp_max)) |> 
  rename(butte_temp = butte_temp_mean)
yuba_gap_min <- yuba_regression_data_full |> 
  filter(is.na(yuba_temp_min), !is.na(butte_temp_min)) |> 
 select(-c(yuba_temp_mean, yuba_temp_min, yuba_temp_max)) |> 
  rename(butte_temp = butte_temp_min)
yuba_gap_max <- yuba_regression_data_full |> 
  filter(is.na(yuba_temp_max), !is.na(butte_temp_max)) |> 
select(-c(yuba_temp_mean, yuba_temp_min, yuba_temp_max)) |> 
  rename(butte_temp = butte_temp_max)

# FOR MODEL use data where there are no missing data for either butte or feather for regression modeling
yuba_regression_data_mean <- yuba_regression_data_full |> 
  filter(!is.na(yuba_temp_mean), !is.na(butte_temp_mean)) |> 
  rename(butte_temp = butte_temp_mean,
         temp = yuba_temp_mean)
yuba_regression_data_min <- yuba_regression_data_full |> 
  filter(!is.na(yuba_temp_min), !is.na(butte_temp_min)) |> 
  rename(butte_temp = butte_temp_min,
         temp = yuba_temp_min)
yuba_regression_data_max <- yuba_regression_data_full |> 
  filter(!is.na(yuba_temp_max), !is.na(butte_temp_max)) |> 
  rename(butte_temp = butte_temp_max,
         temp = yuba_temp_max)
```

*Plot of mean temp for Yuba River and Butte Creek*
```{r}
# plot for mean
ggplot(data = yuba_regression_data_mean, aes(x = butte_temp, y = temp)) +
  geom_point() +
  stat_smooth(method = "lm")
```

*Plot of min temp for Yuba River and Butte Creek*
```{r}
# plot for mean
ggplot(data = yuba_regression_data_min, aes(x = butte_temp, y = temp)) +
  geom_point() +
  stat_smooth(method = "lm")
```

*Plot of max temp for Yuba River and Butte Creek*
```{r}
# plot for mean
ggplot(data = yuba_regression_data_max, aes(x = butte_temp, y = temp)) +
  geom_point() +
  stat_smooth(method = "lm")
```

#### Building Mean Regression
```{r}
# MEAN Regression
split <-rsample::initial_split(yuba_regression_data_mean, prop = 0.8)
train <- rsample::training(split)
test <- rsample::testing(split)
yuba_mod_mean <- lm(temp ~ date + butte_temp, data = train)
summary(yuba_mod_mean)
test_predict <- predict(yuba_mod_mean, test)
test_predict_df <- test |>
  mutate(predicted = test_predict)
# evaluate predictions - MAPE of 10% is not bad
mean(abs((
  test_predict_df$predicted - test_predict_df$temp
)) / test_predict_df$temp)
```

#### Building Min Regression
```{r}
# MIN Regression
split <-rsample::initial_split(yuba_regression_data_min, prop = 0.8)
train <- rsample::training(split)
test <- rsample::testing(split)
yuba_mod_min <- lm(temp ~ date + butte_temp, data = train)
summary(yuba_mod_min)
test_predict <- predict(yuba_mod_min, test)
test_predict_df <- test |>
  mutate(predicted = test_predict)
# evaluate predictions - MAPE of 10% is not bad
mean(abs((
  test_predict_df$predicted - test_predict_df$temp
)) / test_predict_df$temp)
```

#### Building Max Regression
```{r}
# MAX Regression
split <-rsample::initial_split(yuba_regression_data_max, prop = 0.8)
train <- rsample::training(split)
test <- rsample::testing(split)
yuba_mod_max <- lm(temp ~ date + butte_temp, data = train)
summary(yuba_mod_max)
test_predict <- predict(yuba_mod_max, test)
test_predict_df <- test |>
  mutate(predicted = test_predict)
# evaluate predictions - MAPE of 10% is not bad
mean(abs((
  test_predict_df$predicted - test_predict_df$temp
)) / test_predict_df$temp)
```

#### Predictions

```{r}
# Predictions
yuba_gap_predicted_mean <- predict(yuba_mod_mean, yuba_gap_mean)
yuba_mean_predicted <- yuba_gap_mean |> 
  mutate(value = yuba_gap_predicted_mean,
         statistic = "mean_predicted") |> 
  select(date, value, statistic)
ggplot(yuba_mean_predicted, aes(x = date, y = value)) +
  geom_line()
yuba_gap_predicted_min <- predict(yuba_mod_min, yuba_gap_min)
yuba_min_predicted <- yuba_gap_min |> 
  mutate(value = yuba_gap_predicted_min,
         statistic = "min_predicted") |> 
  select(date, value, statistic)
ggplot(yuba_min_predicted, aes(x = date, y = value)) +
  geom_line()
yuba_gap_predicted_max <- predict(yuba_mod_max, yuba_gap_max)
yuba_max_predicted <- yuba_gap_max |> 
  mutate(value = yuba_gap_predicted_max,
         statistic = "max_predicted") |> 
  select(date, value, statistic)
ggplot(yuba_max_predicted, aes(x = date, y = value)) +
  geom_line()
```

#### Full dataset
```{r}
yuba_gap <- bind_rows(yuba_max_predicted,
                      yuba_mean_predicted,
                      yuba_min_predicted) |> 
  mutate(stream = "yuba river") |> 
  pivot_wider(id_cols = c(date, stream), values_from = "value", names_from = "statistic") |> 
  full_join(yuba_format |> 
              pivot_wider(id_cols = c(stream, date, gage_agency, gage_number), values_from = "value", names_from = "statistic")) |> 
  mutate(gage_agency = ifelse(is.na(mean), "interpolated", gage_agency),
         gage_number = ifelse(is.na(mean), "interpolated", gage_number),
         mean = ifelse(is.na(mean), mean_predicted, mean),
         min = ifelse(is.na(min), min_predicted, min),
         max = ifelse(is.na(max), max_predicted, max)) |> 
  select(-c(mean_predicted, max_predicted, min_predicted)) |> 
    pivot_longer(cols = mean:min, values_to = "value", names_to = "statistic") |>
  group_by(stream, date, statistic, gage_agency, gage_number) |> glimpse()
  

ggplot() +
  geom_line(data = yuba_gap, aes(x = date, y = value, color = statistic)) +
  theme_minimal()

```

```{r, include = F}
# write csv
write_csv(yuba_gap, here::here("data-raw", "temperature-data", "yuba_temp_interpolation.csv"))
```