---
title: "Data Checks"
output: html_document
date: "2025-05-17"
---

```{r setup, include=FALSE}
library(tidyverse)
library(SRJPEdata)
library(lubridate)

colors_full <-  c("#9A8822", "#F5CDB4", "#F8AFA8", "#FDDDA0", "#74A089", #Royal 2
                  "#899DA4", "#C93312", "#DC863B", # royal 1 (- 3)
                  "#F1BB7B", "#FD6467", "#5B1A18", # Grand Budapest 1 (-4)
                  "#D8B70A", "#02401B", "#A2A475", # Cavalcanti 1
                  "#E6A0C4", "#C6CDF7", "#D8A499", "#7294D4", #Grand Budapest 2
                  "#9986A5", "#EAD3BF", "#AA9486", "#B6854D", "#798E87" # Isle of dogs 2 altered slightly
)
```

The following checks are performed:

- Find percent difference greater than 2%
- Find where data are missing

There are no longer any differences in the dataset for the following reasons:
- We found an issue where week was being defined differently by two different R packages resulting in different weekly summaries. This has been resolved. Week is defined as Julian Week (7-day period starting with Jan 1)
- We removed the lifestage processing as it is not currently being used.

# Juvenile

```{r, include = F}
# pull data from TEMP_pull_from_edi.R
#source("data-raw/pull_data_scripts/TEMP_pull_from_edi.R")
# This was generated from TEMP_pull_from_edi and loaded to save time
temp_catch <- read_csv("data-raw/data-checks/stream_team_review/temp_catch1.csv")

# run build_rst_model_datasets.R
#weekly_juv_srjpedata <- weekly_juvenile_abundance_catch_data
#write_csv(weekly_juv_srjpedata, "data-raw/data-checks/stream_team_review/weekly_juv_data_run_local.csv")
#local_data <- read_csv("data-raw/data-checks/stream_team_review/weekly_juv_data_run_local.csv")
# ck <- weekly_juvenile_abundance_catch_data_ck |>
#   select(year, week, stream, site, count) |>
#   rename(srjpedata = count) |>
#   full_join(weekly_juv_srjpedata |>
#               select(year, week, stream, site, count))
# filter(ck, srjpedata != count)

# filter to chinook
weekly_edi_catch1 <- temp_catch |> 
  mutate(week = week(date),
         year = year(date),
         species = tolower(species)) |> 
  filter(species %in% c("chinook", "chinook salmon") | is.na(species)) |> 
  group_by(stream, site, week, year) |> 
  summarize(count = sum(count, na.rm = T))


# remove adults and adipose clipped
weekly_edi_catch2_prep <- temp_catch |> 
   mutate(subsite = case_when(site == "okie dam" & is.na(subsite) ~ "okie dam 1", # fix missing subsites
                             stream == "battle creek" & is.na(subsite) & year(date) > 2004 ~ "ubc",
                             site == "yuba river" ~ "hal",
                             T ~ subsite),
         site = case_when(stream == "battle creek" & is.na(site) & year(date) > 2004 ~ "ubc",
                          stream == "yuba river" ~ "hallwood",
                          T ~ site),
         week = week(date),
         year = year(date),
         species = tolower(species),
         life_stage = ifelse(is.na(life_stage), "not recorded", tolower(life_stage)),
         remove = case_when(stream != "butte creek" & adipose_clipped == T ~ "remove",
                            T ~ "keep"))
okie_rst <- weekly_edi_catch2_prep |> 
  mutate(date = as_date(date)) |> 
  filter(site == "okie dam", subsite != "okie dam fyke trap") |> 
  pull(date)

weekly_edi_catch2 <- weekly_edi_catch2_prep |> 
  mutate(butte_fyke_filter = case_when(site == "okie dam" & as_date(date) %in% okie_rst ~ "rst & fyke",
                                       site == "okie dam" & !as_date(date) %in% okie_rst ~ "fyke only",
                                       T ~ "not butte")) |> 
  filter((species %in% c("chinook", "chinook salmon") | is.na(species)),
         life_stage != "adult",
         remove == "keep",
         butte_fyke_filter != "fyke only") |> 
  group_by(stream, site, week, year) |> 
  summarize(count = sum(count, na.rm = T))

weekly_rst_catch <- SRJPEdata::rst_catch |> 
  mutate(week = week(date),
         year = year(date)) |> 
  group_by(stream, site, week, year) |> 
  summarize(count = sum(count, na.rm = T))

catch_compare_1 <- full_join(weekly_rst_catch |> 
                               rename(srjpedata_count = count),
                             weekly_edi_catch1) |> 
   mutate(p_diff = ((abs(count - srjpedata_count))/((count + srjpedata_count)/2)) * 100)

catch_compare_2 <- full_join(SRJPEdata::weekly_juvenile_abundance_catch_data |> 
                               select(year, week, stream, site, count) |> 
                               rename(srjpedata_count = count),
                             weekly_edi_catch2) |> 
   mutate(p_diff = ((abs(count - srjpedata_count))/((count + srjpedata_count)/2)) * 100)
```


## Weekly counts {.tabset}

1. Summary of percent difference
2. Percent difference greater than 2%
3. Missing data

```{r, include = F}
# 1. rst_catch v EDI data
# 2. weekly_juvenile_abundance_catch_data v EDI data

p_diff_summary_plot <- function(data_select, stream_select) {
  data_select |> 
  filter(stream == stream_select) |> 
  ggplot(aes(x = week, y = p_diff, color = site)) +
  geom_point() +
  scale_color_manual(values = colors_full) +
  facet_wrap(~year) +
  theme_bw() +
  labs(y = "percent difference")
}

p_diff_greater_than_2 <- function(data_select, stream_select) {
  data_select |> 
  filter(stream == stream_select, p_diff > 2) |> 
  ggplot(aes(x = week, y = p_diff, color = site)) +
  geom_point() +
  scale_color_manual(values = colors_full) +
  facet_wrap(~year) +
  theme_bw() +
  labs(y = "percent difference")
}

na_plot <- function(data_select, site_select) {
  data_select |> 
    filter(site== site_select, (is.na(srjpedata_count) & !is.na(count)) | (!is.na(srjpedata_count) & is.na(count))) |> 
    pivot_longer(cols = c(srjpedata_count, count), names_to = "count_type", values_to = "value") |> 
  ggplot(aes(x = week, y = value, color = count_type)) +
  geom_point() +
  scale_color_manual(values = colors_full) +
  facet_wrap(~year) +
  theme_bw() +
  labs(y = "count")
}
```

### Battle/Clear

#### Summary of the percent difference between data on EDI and data used for juvenile abundance model

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(catch_compare_2, "battle creek")
```

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(catch_compare_2, "clear creek")
```

#### Highlight datapoints where the percent difference is greater than 2% different

- Battle and Clear Creek do not have any points with greater than 2% difference

#### Find missing data

The missing data in SRJPEdata should align with the years that have been excluded

##### UBC

UBC: run year 1998, 2001, 2003, 2007, 2015

```{r, echo = F, message = F, message = F, warning = F}
SRJPEdata::years_to_exclude_rst_data |> filter(site == "ubc")
```

```{r, echo = F, warning = F, message = F}
na_plot(catch_compare_2, "ubc")
```

##### UCC

UCC: run year 2003, 2009, 2023, 2024

```{r, echo = F, message = F, message = F, warning = F}
SRJPEdata::years_to_exclude_rst_data |> filter(site == "ucc")
```

```{r, echo = F, warning = F, message = F}
na_plot(catch_compare_2, "ucc")
```


### Butte

#### Summary of the percent difference between data on EDI and data used for juvenile abundance model

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(catch_compare_2, "butte creek")
```

#### Highlight datapoints where the percent difference is greater than 2% different

- Butte Creek does not have any differences greater than 2%

```{r, include = F}
# Dig into data points that are off
# Issues were coming from adults not being excluded and this is now fixed

# filter(catch_compare_2, stream == "butte creek", p_diff > 2)
# 
# filter(weekly_juvenile_abundance_catch_data, week == 44, year == 2022, stream == "butte creek") |> view()
# filter(butte_catch_edi, week(date) == 44, year(date) == 2022, stream == "butte creek") 
# filter(rst_catch, julian_week == 44, julian_year == 2022, stream == "butte creek")

```

#### Find missing data

okie dam: run year 1997, 1998, 2005, 2006, 2019

```{r, echo = F, message = F, message = F, warning = F}
SRJPEdata::years_to_exclude_rst_data |> filter(site == "okie dam")
```

```{r, echo = F, warning = F, message = F}
na_plot(catch_compare_2, "okie dam")
```

### Deer/Mill

#### Summary of the percent difference between data on EDI and data used for juvenile abundance model

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(catch_compare_2, "deer creek")
```

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(catch_compare_2, "mill creek")
```

#### Highlight datapoints where the percent difference is greater than 2% different

- Mill and Deer Creek do not have differences greater than 2%


#### Find missing data

deer creek: run year 1992, 1993, 1994, 1997, 1998, 1999, 2004, 2006, 2008, 2022

The EDI data missing is because for the comparison I did not pull from the most recent Deer/Mill EDI package.

```{r, echo = F, message = F, message = F, warning = F}
SRJPEdata::years_to_exclude_rst_data |> filter(site == "deer creek")
```

```{r, echo = F, warning = F, message = F}
na_plot(catch_compare_2, "deer creek")
```

mill creek: run year 1997, 1998, 1999, 2004, 2009, 2021

The EDI data missing is because for the comparison I did not pull from the most recent Deer/Mill EDI package.

```{r, echo = F, message = F, message = F, warning = F}
SRJPEdata::years_to_exclude_rst_data |> filter(site == "mill creek")
```

```{r, echo = F, warning = F, message = F}
na_plot(catch_compare_2, "mill creek")
```

### Feather

#### Summary of the percent difference between data on EDI and data used for juvenile abundance model

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(catch_compare_2, "feather river")
```

#### Highlight datapoints where the percent difference is greater than 2% different

- Feather Rive does not have any data with greater than 2% difference

#### Find missing data

eye riffle: run year 2021

```{r, echo = F, message = F, message = F, warning = F}
SRJPEdata::years_to_exclude_rst_data |> filter(site == "eye riffle")
```


```{r, echo = F, warning = F, message = F}
na_plot(catch_compare_2, "eye riffle")
```

herringer riffle: run year 2010, 2016, 2017, 2021

```{r, echo = F, message = F, message = F, warning = F}
SRJPEdata::years_to_exclude_rst_data |> filter(site == "herringer riffle")
```


```{r, echo = F, warning = F, message = F}
na_plot(catch_compare_2, "herringer riffle")
```

steep riffle: run year 2006, 2013, 2015, 2016

```{r, echo = F, message = F, message = F, warning = F}
SRJPEdata::years_to_exclude_rst_data |> filter(site == "steep riffle")
```


```{r, echo = F, warning = F, message = F}
na_plot(catch_compare_2, "steep riffle")
```

### Yuba

#### Summary of the percent difference between data on EDI and data used for juvenile abundance model

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(catch_compare_2, "yuba river")
```

#### Highlight datapoints where the percent difference is greater than 2% different

- Yuba River does not have any datapoints greater than 2% difference

#### Find missing data

halwood: run years 2000, 2003

```{r, echo = F, message = F, message = F, warning = F}
SRJPEdata::years_to_exclude_rst_data |> filter(site == "hallwood")
```

```{r, echo = F, warning = F, message = F}
na_plot(catch_compare_2, "hallwood")
```

### Knights/Tisdale

#### Summary of the percent difference between data on EDI and data used for juvenile abundance model

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(catch_compare_1, "sacramento river")
```

#### Highlight datapoints where the percent difference is greater than 2% different

- Knights Landing and Tisdale do not have datapoints greater than 2% difference

#### Find missing data

knights landing: run years 2002, 2013

Note than pre 2006 data are not on EDI

```{r, echo = F, message = F, message = F, warning = F}
SRJPEdata::years_to_exclude_rst_data |> filter(site == "knights landing")
```

```{r, echo = F, warning = F, message = F}
na_plot(catch_compare_2, "knights landing")
```

tisdale: run years 2010, 2012

```{r, echo = F, message = F, message = F, warning = F}
SRJPEdata::years_to_exclude_rst_data |> filter(site == "tisdale")
```

```{r, echo = F, warning = F, message = F}
na_plot(catch_compare_2, "tisdale")
```

## Weekly efficiency {.tabset}

- weekly_juvenile_abundance_efficiency_data v EDI data

Note that the data on EDI has a few NA years.

```{r, include = F}
temp_recapture <- read_csv("data-raw/data-checks/stream_team_review/temp_recapture1.csv")
temp_release <- read_csv("data-raw/data-checks/stream_team_review/temp_release1.csv")

weekly_efficiency_edi <- 
  left_join(temp_release, 
            temp_recapture |> # need to summarize first so you don't get duplicated release data when joining
              group_by(stream, site, site_group, release_id) |> 
              summarize(count = sum(count, na.rm = T)),
            by = c("release_id", "stream", "site", "site_group")) |> 
  group_by(stream, 
           site, 
           site_group, 
           #origin,
           #median_fork_length_released, # we add origin and fork length for figures
           week_released = week(date_released), 
           year_released = year(date_released)) |> 
  summarize(number_released = sum(number_released, na.rm = TRUE),
            number_recaptured = sum(count, na.rm = TRUE)) |> 
  ungroup() 
eff_compare <- full_join(SRJPEdata::weekly_juvenile_abundance_efficiency_data |> 
                               select(year, week, stream, site, number_released, number_recaptured) |> 
                               rename(srjpedata_recapture = number_recaptured,
                                      srjpedata_relase = number_released),
                             weekly_efficiency_edi |> 
                           rename(year = year_released,
                                  week = week_released)) |> 
   mutate(p_diff_recapture = ((abs(number_recaptured - srjpedata_recapture))/((number_recaptured + srjpedata_recapture)/2)) * 100,
          p_diff_release = ((abs(number_released - srjpedata_relase))/((number_released + srjpedata_relase)/2)) * 100) |> 
  pivot_longer(cols = c(p_diff_recapture, p_diff_release), names_to = "type", values_to = "p_diff")

eff_compare_release <- filter(eff_compare, type == "p_diff_release")
eff_compare_recapture <- filter(eff_compare, type == "p_diff_recapture")
```

### Battle/Clear

#### Summary of the percent difference between data on EDI and data used for juvenile abundance model

**Recaptures**

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(eff_compare_recapture, "battle creek")
```

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(eff_compare_recapture, "clear creek")
```

**Releases**

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(eff_compare_release, "battle creek")
```

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(eff_compare_release, "clear creek")
```

### Butte

#### Summary of the percent difference between data on EDI and data used for juvenile abundance model

**Recaptures**

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(eff_compare_recapture, "butte creek")
```

**Releases**

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(eff_compare_release, "butte creek")
```

### Deer/Mill

#### Summary of the percent difference between data on EDI and data used for juvenile abundance model

These data are pulled from DataTackle. Additional testing has been done on this data and we are confident they are correct.

### Feather

#### Summary of the percent difference between data on EDI and data used for juvenile abundance model

**Recaptures**

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(eff_compare_recapture, "feather river")
```

**Releases**

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(eff_compare_release, "feather river")
```

### Yuba

#### Summary of the percent difference between data on EDI and data used for juvenile abundance model

**Recaptures**

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(eff_compare_recapture, "yuba river")
```

**Releases**

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(eff_compare_release, "yuba river")
```

### Knights Landing/Tisdale

#### Summary of the percent difference between data on EDI and data used for juvenile abundance model

**Recaptures**

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(eff_compare_recapture, "sacramento river")
```

**Releases**

```{r, echo = F, warning = F, message = F}
p_diff_summary_plot(eff_compare_release, "sacramento river")
```

# Adult {.tabselect}

We do not have a good "source of truth" for the adult data so including plots of annual
counts and data table for review.

```{r, include = F}

adult_plot <- function(stream_select) {
SRJPEdata::observed_adult_input |> 
  filter(stream == stream_select) |> 
    ggplot(aes(x = year, y = count, color = data_type)) +
    geom_line() +
    scale_color_manual(values = colors_full) +
    theme_bw() +
    labs(y = "annual count")
}
```

## Battle/Clear

```{r, echo = F, warning = F, message = F}
adult_plot("battle creek")
```
```{r, echo = F, warning = F, message = F}
adult_plot("clear creek")
```

## Butte

```{r, echo = F, warning = F, message = F}
adult_plot("butte creek")
```

## Deer/Mill

```{r, echo = F, warning = F, message = F}
adult_plot("deer creek")
```
```{r, echo = F, warning = F, message = F}
adult_plot("mill creek")
```

## Feather

Note that currently Feather River adult data is not being used because we do not have
a method for identifying spring run.

```{r, echo = F, warning = F, message = F}
adult_plot("feather river")
```

## Yuba

```{r, echo = F, warning = F, message = F}
adult_plot("yuba river")
```
