---
title: "Data QC - Visual Inspection"
format:
    html:
        embed-resources: true
        toc: true
        toc-depth: 3
        toc-title: "Contents"
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(tidyverse)
library(lubridate)
library(plotly)
library(knitr)

# Helper: IQR-based outlier flag
flag_outliers <- function(x, multiplier = 3) {
  q <- quantile(x, c(0.25, 0.75), na.rm = TRUE)
  iqr <- q[2] - q[1]
  x < (q[1] - multiplier * iqr) | x > (q[2] + multiplier * iqr)
}
```

## Overview

This document provides visual QC checks for the primary SR JPE data objects. The goal is to identify potential **outliers**, **data gaps**, and **anomalies** that may warrant follow-up. Each section covers one data object.

---

# 1. rst_catch

RST catch monitoring data across all tributary and mainstem sites.

```{r rst-catch-prep}
catch <- SRJPEdata::rst_catch |>
  mutate(
    year = year(date),
    month = month(date),
    week = week(date),
    run_year = ifelse(week >= 45, year + 1, year)
  )

# Daily total catch per site
daily_catch <- catch |>
  filter(!is.na(count)) |>
  group_by(date, stream, site, run_year) |>
  summarise(total_count = sum(count, na.rm = TRUE), .groups = "drop") |>
  mutate(outlier = flag_outliers(total_count))

# Weekly catch for trend plots
weekly_catch <- catch |>
  filter(!is.na(count)) |>
  group_by(stream, site, run_year, week) |>
  summarise(total_count = sum(count, na.rm = TRUE), .groups = "drop")
```

## 1a. Total catch over time by site

```{r rst-catch-time}
p <- catch |>
  filter(!is.na(count)) |>
  group_by(date, stream, site) |>
  summarise(total_count = sum(count, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = date, y = total_count, color = site)) +
  geom_line(alpha = 0.6) +
  scale_y_log10() +
  theme_minimal() +
  facet_wrap(~stream, scales = "free_x") +
  theme(legend.position = "none") +
  labs(title = "Daily total catch by site (log scale)", x = NULL, y = "Total count (log10)", color = "Site")

ggplotly(p)
```

## 1b. Potential outlier catch days (>3x IQR above 75th percentile, by site)

```{r rst-catch-outliers}
outlier_days <- daily_catch |>
  filter(outlier) |>
  arrange(desc(total_count)) |>
  select(date, stream, site, run_year, total_count)

kable(outlier_days, caption = paste0("Days with outlier catch values (n = ", nrow(outlier_days), ")"))
```

## 1c. Annual catch by site

```{r rst-catch-annual}
annual_catch <- catch |>
  filter(!is.na(count)) |>
  group_by(stream, site, run_year) |>
  summarise(total_count = sum(count, na.rm = TRUE), .groups = "drop")

p2 <- annual_catch |>
  ggplot(aes(x = run_year, y = total_count, color = site)) +
  geom_line() +
  geom_point(size = 1) +
  theme_minimal() +
  facet_wrap(~stream, scales = "free_y") +
  theme(legend.position = "none") +
  labs(title = "Annual catch by site and stream", x = "Run year", y = "Total count")

ggplotly(p2)
```

## 1d. Data gaps — weeks with no catch data within the run year window (weeks 45–53 & 1–22)

```{r rst-catch-gaps}
# Expected weeks in window
window_weeks <- c(45:53, 1:22)

weeks_observed <- catch |>
  filter(!is.na(count), week %in% window_weeks) |>
  select(stream, site, run_year, week) |>
  distinct()

# Sites and run years present in data
site_run_years <- weeks_observed |>
  distinct(stream, site, run_year)

# All expected combinations
expected <- site_run_years |>
  crossing(week = window_weeks) |>
  left_join(weeks_observed |> mutate(sampled = TRUE), by = c("stream", "site", "run_year", "week")) |>
  mutate(sampled = replace_na(sampled, FALSE))

gap_summary <- expected |>
  group_by(stream, site, run_year) |>
  summarise(
    n_weeks_sampled = sum(sampled),
    n_weeks_missing = sum(!sampled),
    pct_missing = round(100 * mean(!sampled), 1),
    .groups = "drop"
  ) |>
  filter(n_weeks_missing > 0) |>
  arrange(desc(pct_missing))

p3 <- gap_summary |>
  ggplot(aes(x = run_year, y = n_weeks_sampled)) +
  geom_line(color = "steelblue", alpha = 0.7) +
  geom_point(color = "steelblue", size = 1.5) +
  geom_hline(yintercept = 19, linetype = "dashed", color = "red") +
  facet_wrap(~site, scales = "free_x") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 4)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Weeks sampled per run year (window: wks 45-53 & 1-22)",
    subtitle = "Red dashed line = 19-week threshold",
    x = "Run year", y = "Weeks sampled"
  )

ggplotly(p3)
```

```{r rst-catch-gaps-table}
kable(
  gap_summary |> filter(n_weeks_sampled < 19) |> arrange(stream, site, run_year),
  caption = "Site-years below the 19-week sampling threshold"
)
```

---

# 2. release

RST efficiency trial release data.

```{r release-prep}
rel <- SRJPEdata::release |>
  mutate(year = year(date_released))
```

## 3a. Number released over time by stream

```{r release-time}
p7 <- rel |>
  ggplot(aes(x = date_released, y = number_released, color = stream)) +
  geom_point(alpha = 0.5) +
  theme_minimal() +
  facet_wrap(~stream, scales = "free") +
  labs(title = "Number released per trial over time", x = NULL, y = "Number released")

ggplotly(p7)
```

## 3b. Number released — outliers

```{r release-outliers}
rel_outliers <- rel |>
  filter(!is.na(number_released)) |>
  mutate(outlier = flag_outliers(number_released)) |>
  filter(outlier) |>
  select(date_released, stream, site, number_released, run, life_stage) |>
  arrange(desc(number_released))

kable(rel_outliers, caption = paste0("Release trial outliers by number released (n = ", nrow(rel_outliers), ")"))
```

## 3c. Trials per year by stream

```{r release-annual}
p8 <- rel |>
  count(stream, site, year) |>
  ggplot(aes(x = year, y = n, color = site)) +
  geom_line() +
  geom_point(size = 1) +
  theme_minimal() +
  facet_wrap(~stream, scales = "free_y") +
  theme(legend.position = "none") +
  labs(title = "Release trials per year by site", x = "Year", y = "Number of trials")

ggplotly(p8)
```

## 3d. Years with no release trials

```{r release-gaps}
rel_coverage <- rel |>
  distinct(stream, site, year) |>
  group_by(stream, site) |>
  summarise(
    min_year = min(year),
    max_year = max(year),
    n_years = n(),
    expected_years = max_year - min_year + 1,
    gap_years = expected_years - n_years,
    .groups = "drop"
  ) |>
  filter(gap_years > 0) |>
  arrange(desc(gap_years))

kable(rel_coverage, caption = "Sites with gaps in release trial coverage")
```

---

# 4. recaptures

RST efficiency trial recapture data.

```{r recaptures-prep}
recap <- SRJPEdata::recaptures |>
  mutate(year = year(date))

# Join to release to compute efficiency
rel_summary <- SRJPEdata::release |>
  group_by(release_id, stream, site) |>
  summarise(number_released = sum(number_released, na.rm = TRUE), .groups = "drop")

recap_summary <- recap |>
  filter(!is.na(count)) |>
  group_by(release_id, stream, site) |>
  summarise(number_recaptured = sum(count, na.rm = TRUE), .groups = "drop")

efficiency_trials <- rel_summary |>
  left_join(recap_summary, by = c("release_id", "stream", "site")) |>
  mutate(
    number_recaptured = replace_na(number_recaptured, 0),
    efficiency = number_recaptured / number_released,
    date_released = SRJPEdata::release$date_released[match(release_id, SRJPEdata::release$release_id)]
  ) |>
  filter(!is.na(date_released))
```

## 4a. Efficiency over time by stream

```{r recaptures-efficiency-time}
p9 <- efficiency_trials |>
  filter(efficiency <= 1) |>
  ggplot(aes(x = date_released, y = efficiency, color = stream)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", se = FALSE, linetype = "dashed") +
  theme_minimal() +
  facet_wrap(~stream, scales = "free_x") +
  labs(title = "Trial efficiency over time by stream", x = NULL, y = "Efficiency (recaptured / released)")

ggplotly(p9)
```

## 4b. Suspiciously high efficiency (> 0.5)

```{r recaptures-high-efficiency}
high_eff <- efficiency_trials |>
  filter(efficiency > 0.5) |>
  arrange(desc(efficiency)) |>
  select(release_id, date_released, stream, site, number_released, number_recaptured, efficiency)

kable(high_eff, caption = paste0("Trials with efficiency > 0.5 (n = ", nrow(high_eff), ")"))
```

## 4c. Trials with zero recaptures

```{r recaptures-zero}
zero_recap <- efficiency_trials |>
  filter(number_recaptured == 0) |>
  count(stream, site) |>
  arrange(desc(n))

kable(zero_recap, caption = "Number of trials with zero recaptures by stream/site")
```

---

# 5. weekly_juvenile_abundance_catch_data

Weekly summarized catch for BT-SPAS-X model input.

```{r wja-catch-prep}
wja_catch <- SRJPEdata::weekly_juvenile_abundance_catch_data
```

## 5a. Weekly count over time by stream

```{r wja-catch-time}
p11 <- wja_catch |>
  filter(!is.na(count)) |>
  mutate(approx_date = make_date(run_year, 1, 1) + weeks(week - 1)) |>
  ggplot(aes(x = approx_date, y = count, color = stream)) +
  geom_line(alpha = 0.6) +
  scale_y_log10() +
  theme_minimal() +
  facet_wrap(~stream, scales = "free_x") +
  labs(title = "Weekly catch (log scale) by stream", x = NULL, y = "Count (log10)")

ggplotly(p11)
```

## 5b. Count outliers by site

```{r wja-catch-outliers}
wja_count_outliers <- wja_catch |>
  filter(!is.na(count)) |>
  group_by(stream, site) |>
  mutate(outlier = flag_outliers(count)) |>
  ungroup() |>
  filter(outlier) |>
  select(run_year, week, stream, site, count) |>
  arrange(desc(count))

kable(wja_count_outliers, caption = paste0("Weekly count outliers (n = ", nrow(wja_count_outliers), ")"))
```

## 5c. Annual hours fished by site

```{r wja-catch-hours}
annual_hours <- wja_catch |>
  filter(!is.na(hours_fished)) |>
  group_by(stream, site, run_year) |>
  summarise(total_hours = sum(hours_fished, na.rm = TRUE), .groups = "drop")

p12 <- annual_hours |>
  ggplot(aes(x = run_year, y = total_hours, color = site)) +
  geom_line(alpha = 0.7) +
  geom_point(size = 1.5) +
  facet_wrap(~stream, scales = "free") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 4)) +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Annual hours fished by site", x = "Run year", y = "Total hours fished")

ggplotly(p12)
```

## 5d. Flow (cfs) over time

```{r wja-catch-flow}
p13 <- wja_catch |>
  filter(!is.na(flow_cfs)) |>
  mutate(approx_date = make_date(run_year, 1, 1) + weeks(week - 1)) |>
  ggplot(aes(x = approx_date, y = flow_cfs, color = stream)) +
  geom_line(alpha = 0.5) +
  scale_y_log10() +
  theme_minimal() +
  facet_wrap(~stream, scales = "free_x") +
  labs(title = "Weekly flow (cfs, log scale) by stream", x = NULL, y = "Flow CFS (log10)")

ggplotly(p13)
```

## 5e. Weeks sampled heatmap by site

Each cell represents one week within a run year. Green = sampled (non-NA count), white = not sampled.

```{r wja-catch-heatmap, fig.height=20}
# Reindex weeks so run year window reads left to right (wk 45 = index 1)
heatmap_data <- wja_catch |>
  mutate(
    week_index = if_else(week >= 45, week - 44L, week + 9L)
  )

# All site x run_year x week_index combinations that exist in the data
all_combos <- heatmap_data |>
  distinct(stream, site, run_year) |>
  crossing(week_index = 1:31)

sampled <- heatmap_data |>
  filter(!is.na(count)) |>
  distinct(stream, site, run_year, week_index) |>
  mutate(sampled = TRUE)

plot_data <- all_combos |>
  left_join(sampled, by = c("stream", "site", "run_year", "week_index")) |>
  mutate(sampled = replace_na(sampled, FALSE))

p <- plot_data |>
  ggplot(aes(x = week_index, y = factor(run_year), fill = sampled,
             text = paste0("Run year: ", run_year, "<br>Week index: ", week_index))) +
  geom_tile(color = "white", linewidth = 0.3) +
  scale_fill_manual(values = c("TRUE" = "#4CAF50", "FALSE" = "#EEEEEE"),
                    labels = c("Sampled", "Not sampled")) +
  scale_x_continuous(breaks = c(1, 5, 10, 15, 20, 25, 31),
                     labels = c("wk45", "wk49", "wk3", "wk8", "wk13", "wk18", "wk22")) +
  facet_wrap(~site, ncol = 2, scales = "free_y") +
  theme_minimal(base_size = 9) +
  theme(panel.grid = element_blank(), legend.position = "top") +
  labs(title = "Weeks sampled per run year by site",
       subtitle = "Week index 1 = week 45 (Nov); index 31 = week 22 (Jun)",
       x = "Week (run-year indexed)", y = "Run year", fill = NULL)

ggplotly(p, tooltip = "text")
```

---

# 6. weekly_juvenile_abundance_efficiency_data

Weekly efficiency data for BT-SPAS-X model input.

```{r wja-eff-prep}
wja_eff <- SRJPEdata::weekly_juvenile_abundance_efficiency_data
```

## 6a. Efficiency (recaptured / released) over time

```{r wja-eff-time}
wja_eff_rates <- wja_eff |>
  filter(!is.na(number_released), !is.na(number_recaptured), number_released > 0) |>
  mutate(
    efficiency = number_recaptured / number_released,
    approx_date = make_date(run_year, 1, 1) + weeks(week - 1)
  )

p14 <- wja_eff_rates |>
  ggplot(aes(x = approx_date, y = efficiency, color = stream)) +
  geom_point(alpha = 0.5) +
  theme_minimal() +
  facet_wrap(~stream, scales = "free_x") +
  labs(title = "Weekly efficiency over time by stream", x = NULL, y = "Efficiency")

ggplotly(p14)
```

## 6b. Efficiency outliers

```{r wja-eff-outliers}
eff_outliers <- wja_eff_rates |>
  group_by(stream, site) |>
  mutate(outlier = flag_outliers(efficiency)) |>
  ungroup() |>
  filter(outlier | efficiency > 1) |>
  select(run_year, week, stream, site, number_released, number_recaptured, efficiency) |>
  arrange(desc(efficiency))

kable(eff_outliers, caption = paste0("Efficiency outliers (n = ", nrow(eff_outliers), ")"))
```

## 6c. Trials per run year by stream

```{r wja-eff-trials}
p15 <- wja_eff |>
  filter(!is.na(number_released)) |>
  count(stream, site, run_year) |>
  ggplot(aes(x = run_year, y = n, color = site)) +
  geom_line() +
  geom_point(size = 1) +
  theme_minimal() +
  facet_wrap(~stream, scales = "free_y") +
  theme(legend.position = "none") +
  labs(title = "Efficiency trials per run year", x = "Run year", y = "Number of weekly efficiency obs")

ggplotly(p15)
```

## 6d. Standardized efficiency flow distribution

```{r wja-eff-flow}
p16 <- wja_eff |>
  filter(!is.na(standardized_efficiency_flow)) |>
  ggplot(aes(x = standardized_efficiency_flow, fill = stream)) +
  geom_histogram(bins = 40, alpha = 0.7, show.legend = FALSE) +
  facet_wrap(~stream, scales = "free_y") +
  theme_minimal() +
  labs(title = "Standardized efficiency flow distribution", x = "Standardized flow", y = "Count")

ggplotly(p16)
```

---

# 7. years_to_exclude_rst_data

Years flagged as incomplete for RST data.

```{r yex-rst-prep}
yex_rst <- SRJPEdata::years_to_exclude_rst_data
```

## 7a. Table of excluded years

```{r yex-rst-table}
kable(yex_rst |> arrange(stream, site, run_year), caption = "All excluded RST site-years")
```

## 7b. Excluded years by stream and site

```{r yex-rst-heatmap}
# All possible site-years from rst_catch
all_site_years <- SRJPEdata::rst_catch |>
  filter(!is.na(count)) |>
  mutate(
    week = week(date),
    run_year = ifelse(week >= 45, year(date) + 1, year(date))
  ) |>
  distinct(stream, site, run_year) |>
  filter(!site %in% c("powerhouse", "adams dam"))

site_year_status <- all_site_years |>
  left_join(
    yex_rst |> mutate(excluded = TRUE) |> select(stream, site, run_year, excluded),
    by = c("stream", "site", "run_year")
  ) |>
  mutate(excluded = replace_na(excluded, FALSE))

p17 <- site_year_status |>
  ggplot(aes(x = run_year, y = site, fill = excluded)) +
  geom_tile(color = "white") +
  scale_fill_manual(values = c("FALSE" = "#4CAF50", "TRUE" = "#F44336"),
                    labels = c("Included", "Excluded")) +
  theme_minimal() +
  labs(title = "RST site-year inclusion status", x = "Run year", y = NULL, fill = "Status")

ggplotly(p17)
```

## 7c. Number of excluded years by stream

```{r yex-rst-summary}
yex_rst |>
  count(stream, site, exclusion_type, apply_to, name = "n_excluded_years") |>
  arrange(desc(n_excluded_years)) |>
  kable(caption = "Excluded years by stream, site, and exclusion type")
```

