---
title: "Efficiency by origin of fish released"
format:
  html:
    embed-resources: true
---

This document compares RST efficiency trial values by origin of fish released
(`origin_released` in `SRJPEdata::weekly_efficiency`). The goal is to assess
whether hatchery, natural, or mixed-origin fish yield meaningfully different
efficiency estimates, which has implications for how efficiency trials are
selected or pooled in the SRJPE model.

Efficiency is calculated as `number_recaptured / number_released` for each
weekly trial.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

colors_full <-  c("#9A8822", "#F5CDB4", "#F8AFA8", "#FDDDA0", "#74A089",
                  "#899DA4", "#C93312", "#DC863B",
                  "#F1BB7B", "#FD6467", "#5B1A18",
                  "#D8B70A", "#02401B", "#A2A475",
                  "#E6A0C4", "#C6CDF7", "#D8A499", "#7294D4",
                  "#9986A5", "#EAD3BF", "#AA9486", "#B6854D", "#798E87"
)

origin_colors <- c("natural" = "#74A089", "hatchery" = "#C93312",
                   "mixed" = "#D8B70A", "unknown" = "#899DA4")
```

```{r load-data, include=FALSE}
library(tidyverse)

efficiency <- SRJPEdata::weekly_efficiency |>
  mutate(
    efficiency = number_recaptured / number_released,
    origin_released = factor(origin_released, levels = c("natural", "hatchery", "mixed"))
  )
```

# Overview of origin coverage

```{r origin-summary}
efficiency |>
  count(year_released, stream, site, origin_released, name = "n_trials") |>
  arrange(stream, site, year_released, origin_released) |>
  rename(Year = year_released, Stream = stream, Site = site,
         Origin = origin_released, `N trials` = n_trials) |>
  knitr::kable(caption = "Number of weekly trials by year, stream, site, and origin")
```

Most sites release only natural-origin fish. Hatchery and mixed-origin releases
are concentrated at Sacramento River (Knights Landing, Tisdale), Yuba River
(Hallwood), Deer Creek, Mill Creek, and recent years on the Feather River.
Clear Creek has no recorded origin for any trial (`NA` throughout).

```{r origin-na-note}
n_na <- sum(is.na(efficiency$origin_released))
n_total <- nrow(efficiency)
```

**`r n_na` of `r n_total` trials (`r round(100 * n_na / n_total)`%) have
`NA` for `origin_released`**, including all Clear Creek trials and older records
at Battle Creek, Sacramento River (pre-2015), and several Feather River sites.
These are excluded from origin comparisons below.

# Overall efficiency distribution by origin

```{r overall-boxplot, fig.height=4, fig.width=6}
efficiency |>
  filter(!is.na(origin_released), efficiency <= 1) |>
  ggplot(aes(x = origin_released, y = efficiency, fill = origin_released)) +
  geom_boxplot(outlier.shape = 21, outlier.size = 1.5, alpha = 0.8) +
  scale_fill_manual(values = origin_colors, guide = "none") +
  labs(x = "Origin", y = "Efficiency (recaptured / released)",
       title = "Efficiency by origin — all sites combined") +
  theme_bw()
```

```{r overall-summary-table}
efficiency |>
  filter(!is.na(origin_released), efficiency <= 1) |>
  group_by(origin_released) |>
  summarize(
    n_trials      = n(),
    median_eff    = round(median(efficiency, na.rm = TRUE), 4),
    mean_eff      = round(mean(efficiency, na.rm = TRUE), 4),
    sd_eff        = round(sd(efficiency, na.rm = TRUE), 4),
    pct_zero      = round(100 * mean(efficiency == 0), 1),
    .groups = "drop"
  ) |>
  knitr::kable(
    col.names = c("Origin", "N trials", "Median eff.", "Mean eff.", "SD", "% zero recapture"),
    caption = "Summary statistics by origin (efficiency > 1 excluded)"
  )
```

# Efficiency by origin within each stream

```{r by-stream-boxplot, fig.height=10, fig.width=9}
efficiency |>
  filter(!is.na(origin_released), efficiency <= 1) |>
  ggplot(aes(x = origin_released, y = efficiency, fill = origin_released)) +
  geom_boxplot(outlier.shape = 21, outlier.size = 1.2, alpha = 0.8) +
  scale_fill_manual(values = origin_colors, guide = "none") +
  facet_wrap(~ stream, scales = "free_y", ncol = 2) +
  labs(x = "Origin", y = "Efficiency",
       title = "Efficiency by origin within each stream") +
  theme_bw() +
  theme(strip.text = element_text(size = 9))
```

# Efficiency by origin within each site

```{r by-site-boxplot, fig.height=14, fig.width=10}
efficiency |>
  filter(!is.na(origin_released), efficiency <= 1) |>
  ggplot(aes(x = origin_released, y = efficiency, fill = origin_released)) +
  geom_boxplot(outlier.shape = 21, outlier.size = 1.2, alpha = 0.8) +
  scale_fill_manual(values = origin_colors, guide = "none") +
  facet_wrap(~ site, scales = "free_y", ncol = 3) +
  labs(x = "Origin", y = "Efficiency",
       title = "Efficiency by origin within each site") +
  theme_bw() +
  theme(strip.text = element_text(size = 8))
```

# Efficiency over time by origin

Comparing efficiency trends across years helps distinguish whether origin
differences are confounded with temporal trends (e.g., hatchery releases only
occurring in recent years at some sites).

```{r time-series-by-origin, fig.height=14, fig.width=10}
efficiency |>
  filter(!is.na(origin_released), efficiency <= 1) |>
  group_by(stream, site, year_released, origin_released) |>
  summarize(mean_efficiency = mean(efficiency, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = year_released, y = mean_efficiency,
             color = origin_released, shape = origin_released)) +
  geom_point(size = 2) +
  geom_line(linewidth = 0.5) +
  scale_color_manual(values = origin_colors) +
  facet_wrap(~ site, scales = "free_y", ncol = 2) +
  labs(x = "Year", y = "Mean annual efficiency",
       color = "Origin", shape = "Origin",
       title = "Mean annual efficiency by origin and site") +
  theme_bw() +
  theme(strip.text = element_text(size = 8),
        legend.position = "bottom")
```

# Sites with multiple origins — direct comparison

For sites where both hatchery and natural releases occur, we can make a more
direct comparison.

```{r multi-origin-sites}
multi_origin_sites <- efficiency |>
  filter(!is.na(origin_released)) |>
  distinct(site, origin_released) |>
  count(site) |>
  filter(n > 1) |>
  pull(site)
```

Sites with more than one origin: **`r paste(multi_origin_sites, collapse = ", ")`**.

```{r multi-origin-boxplot, fig.height=6, fig.width=9}
efficiency |>
  filter(!is.na(origin_released), efficiency <= 1,
         site %in% multi_origin_sites) |>
  ggplot(aes(x = origin_released, y = efficiency, fill = origin_released)) +
  geom_boxplot(outlier.shape = 21, outlier.size = 1.5, alpha = 0.8) +
  scale_fill_manual(values = origin_colors, guide = "none") +
  facet_wrap(~ site, scales = "free_y") +
  labs(x = "Origin", y = "Efficiency",
       title = "Efficiency by origin — sites with multiple origins only") +
  theme_bw() +
  theme(strip.text = element_text(size = 9))
```

```{r multi-origin-table}
efficiency |>
  filter(!is.na(origin_released), efficiency <= 1,
         site %in% multi_origin_sites) |>
  group_by(site, origin_released) |>
  summarize(
    n_trials   = n(),
    median_eff = round(median(efficiency, na.rm = TRUE), 4),
    mean_eff   = round(mean(efficiency, na.rm = TRUE), 4),
    sd_eff     = round(sd(efficiency, na.rm = TRUE), 4),
    .groups    = "drop"
  ) |>
  knitr::kable(
    col.names = c("Site", "Origin", "N trials", "Median eff.", "Mean eff.", "SD"),
    caption   = "Efficiency summary for sites with multiple origins"
  )
```

# Efficiency by week of year and origin

Seasonal patterns in efficiency may differ by origin if hatchery fish are
released in different weeks than natural fish.

```{r weekly-pattern, fig.height=6, fig.width=9}
efficiency |>
  filter(!is.na(origin_released), efficiency <= 1) |>
  group_by(week_released, origin_released) |>
  summarize(
    mean_eff = mean(efficiency, na.rm = TRUE),
    n        = n(),
    .groups  = "drop"
  ) |>
  filter(n >= 3) |>
  ggplot(aes(x = week_released, y = mean_eff,
             color = origin_released, linewidth = n)) +
  geom_line(linewidth = 0.8) +
  geom_point(aes(size = n), alpha = 0.7) +
  scale_color_manual(values = origin_colors) +
  labs(x = "Week of year", y = "Mean efficiency",
       color = "Origin", size = "N trials",
       title = "Mean efficiency by week and origin (all sites, n ≥ 3 trials per week)") +
  theme_bw() +
  theme(legend.position = "bottom")
```

# Summary

```{r summary-stats-wide}
efficiency |>
  filter(!is.na(origin_released), efficiency <= 1) |>
  group_by(origin_released) |>
  summarize(
    n_trials   = n(),
    n_sites    = n_distinct(site),
    n_years    = n_distinct(year_released),
    median_eff = round(median(efficiency, na.rm = TRUE), 4),
    mean_eff   = round(mean(efficiency, na.rm = TRUE), 4),
    .groups    = "drop"
  ) |>
  knitr::kable(
    col.names = c("Origin", "N trials", "N sites", "N years",
                  "Median eff.", "Mean eff."),
    caption   = "Cross-site summary by origin"
  )
```

Key takeaways:

- **Natural-origin** releases dominate the dataset and are available across nearly
  all sites and years.
- **Hatchery and mixed** releases are limited to a small number of sites
  (Sacramento River, Yuba River, Deer Creek, Mill Creek, and recent Feather River).
  Direct comparisons are constrained by this imbalance.
- Where both origins are present at the same site (e.g., Knights Landing,
  Tisdale, Hallwood, eye riffle, herringer riffle), mean efficiency values are
  generally similar between natural and hatchery fish, though variability is high.
- **Mixed-origin** trials have few replicates; differences from natural or
  hatchery should be interpreted cautiously.
- Origin differences are partially confounded with time: hatchery releases at
  most sites begin only in recent years, and efficiency has changed over time
  independently.
