---
title: "QC Feather Flows"
format:
    html:
        embed-resources: true
---

Kassie Henley (DWR) provided guidance that the CDEC GRL gage is not representative, especially during low flows. Instead they use Hatchery (FRB) + Oroville (ORF) + Thermalito Afterbay (TAO). FRB and ORF are available via CDEC but TAO previously has not been. They manage their own internal database which they provided to us in May 2025. 

- We used the DWR database to generate HFC flows (Hatch.Oroville + TAO) however this was different than the Excel Kassie provided. Why are they different? To answer this, I looked at the database in Access (to verify that no loading issues existed). I pulled the Oroville and Hatchery columns separately. 
- We also discovered that TAO data is available on USGS. I compared this data to that in the database to identify if they are similar. 

**Key Takeaways**

- There is an issue with the `Hatch.Oroville` column which is related to how dates were being read in though was not affecting the other columns. For the 2026 database I exported the `Flows Master` table to Excel first though data for 2024 for this column did not exist so I was not able to compare. Decided that this column should not be used and instead the `Hatchery` also called `Hatchery (CDEC ORF; USGS 11406930)` and `Oroville` also called `Oroville (CDEC TFB; USGS 11406999)` should be used. The better practice however would be to pull these directly from CDEC for use in the SRJPE.
- I compared the TAO data from USGS with the TAO data from the 2026 DWR database and there are some differences primarily in 2011, 2012, and 2017. I do not think the differences are big enough to warrant using data that is not publically available as it will slow down the workflow.
- ORF and TFB pulled directly from USGS/CDEC are a good match with the DWR database (as expected). We will pull data directly from these sources as it is more transparent.

# Compare April 2025 and Excel

When use Hatch.Oroville I do not get correct values but when use Hatchery and Oroville separately I do. I looked at the db in Access and it looks like some of the dates are not being loaded in correctly and that this is happening for the Hatch.Oroville column. 

```{r, include = F}
## April 2025 database
library(Hmisc)
library(lubridate)
library(tidyverse)
feather_flow_db <- here::here("data-raw", "TEMP_data", "feather_flow_db.accdb")
mdb.get(feather_flow_db, tables = T)
feather_flow_raw <- mdb.get(feather_flow_db, tables = "Flows Master")
feather_gage_raw <- mdb.get(feather_flow_db, tables = "Gauge Locations")
# Assume that Hatch.Oroville is FRB+ORF then need to add TAO
detach(package:Hmisc)
feather_flow <- feather_flow_raw |> 
  group_by(Date) |> 
  summarize(hatch_oroville = mean(Hatch.Oroville, na.rm = T),
            TAO = mean(TAO, na.rm = T)) |> 
  mutate(value = hatch_oroville+TAO) |> 
  select(-c(hatch_oroville, TAO)) |> 
  rename(date = Date) |> 
  filter(date <= "2025-03-31") |> 
  mutate(stream = "feather river", 
         site_group = "upper feather hfc",
         gage_agency = "DWR Feather",
         gage_number = "FRB+ORF+TAO",
         parameter = "flow",
         statistic = "mean")

feather_flow |> 
    filter(year(date) == 2024)
```

```{r, include = F}
## Excel from Kassie
feather_flow_excel <- readxl::read_xlsx(here::here("data-raw", "TEMP_data", "FR_Flows_2024.xlsx"))

feather_flow_excel |> 
    filter(year(Date) == 2024)
```

```{r, include = F}
compare <- feather_flow |> 
    select(date, value) |> 
    mutate(date = ymd(date)) |> 
    rename(april_2025_db = value) |> 
    left_join(feather_flow_excel |> 
    select(Date, `HFC (Total FR)`) |> 
    mutate(date = as_date(Date),
    feather_excel = `HFC (Total FR)`) |> 
        select(date, feather_excel)) |> 
    mutate(compare_difference = april_2025_db - feather_excel) |>  view()
```

Comparison of the 2024 Excel from Kassie and the Hatch.Oroville column
```{r, echo = F, message = F, warning = F}
knitr::kable(compare |> 
    filter(year(date) == 2024))
```


```{r, include = F}
## Look at other columns in 2025 DB
feather_flow2 <- feather_flow_raw |> 
  group_by(Date) |> 
  summarize(hatch = mean(Hatchery, na.rm = T),
            oro = mean(Oroville, na.rm = T),
            TAO = mean(TAO, na.rm = T)) |> 
  mutate(value = hatch + oro +TAO) |> 
  select(-c(hatch, oro, TAO)) |> 
  rename(date = Date) |> 
  filter(date <= "2025-03-31") |> 
  mutate(stream = "feather river", 
         site_group = "upper feather hfc",
         gage_agency = "DWR Feather",
         gage_number = "FRB+ORF+TAO",
         parameter = "flow",
         statistic = "mean")

feather_flow2 |> 
    filter(year(date) == 2024)

compare2 <- feather_flow2 |> 
    select(date, value) |> 
    mutate(date = ymd(date)) |> 
    rename(april_2025_db = value) |> 
    left_join(feather_flow_excel |> 
    select(Date, `HFC (Total FR)`) |> 
    mutate(date = as_date(Date),
    feather_excel = `HFC (Total FR)`) |> 
        select(date, feather_excel)) |> 
    mutate(compare_difference = april_2025_db - feather_excel) |>  view()
```

Comparison of the 2024 Excel from Kassie and the Hatchery, Oroville column
```{r, echo = F, message = F, warning = F}
knitr::kable(compare2 |> 
    filter(year(date) == 2024))
```

```{r, include = F}
feather_2026_db_export <- readxl::read_xlsx(here::here("data-raw","TEMP_data","Flows_2_23_26_flows_master.xlsx"), col_types = c("numeric","date","date","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric"))

feather_20261 <- feather_2026_db_export |>
    group_by(Date) |> 
  summarize(hatch = mean(`Hatchery (CDEC ORF; USGS 11406930)`, na.rm = T),
            oro = mean(`Oroville (CDEC TFB; USGS 11406999)`, na.rm = T),
            TAO = mean(TAO, na.rm = T)) |> 
  mutate(value = hatch + oro + TAO) |> 
  select(-c(hatch, oro, TAO)) |> 
  rename(date = Date) |> 
  filter(date <= "2025-03-31") |> 
  mutate(stream = "feather river", 
         site_group = "upper feather hfc",
         gage_agency = "DWR Feather",
         gage_number = "FRB+ORF+TAO",
         parameter = "flow",
         statistic = "mean")

compare3 <- feather_20261 |> 
    select(date, value) |> 
    mutate(date = as_date(date)) |> 
    rename(april_2026 = value) |> 
    left_join(feather_flow_excel |> 
    select(Date, `HFC (Total FR)`) |> 
    mutate(date = ymd(Date),
    feather_excel = `HFC (Total FR)`) |> 
        select(date, feather_excel)) |> 
    mutate(compare = april_2026 - feather_excel)

feather_20262 <- feather_2026_db_export |>
    group_by(Date) |> 
  summarize(hatch = mean(`USGS 11407000`, na.rm = T),
            TAO = mean(TAO, na.rm = T)) |> 
  mutate(value = hatch + TAO) |> 
  select(-c(hatch,TAO)) |> 
  rename(date = Date) |> 
  filter(date <= "2025-03-31") |> 
  mutate(stream = "feather river", 
         site_group = "upper feather hfc",
         gage_agency = "DWR Feather",
         gage_number = "FRB+ORF+TAO",
         parameter = "flow",
         statistic = "mean")

compare4 <- feather_20262 |> 
    select(date, value) |> 
    mutate(date = as_date(date)) |> 
    rename(april_2026 = value) |> 
    left_join(feather_flow_excel |> 
    select(Date, `HFC (Total FR)`) |> 
    mutate(date = ymd(Date),
    feather_excel = `HFC (Total FR)`) |> 
        select(date, feather_excel)) |> 
    mutate(compare = april_2026 - feather_excel) |> view()
```

# TAO on USGS

Can we use all publically available data?

To answer this we compare [USGS TAO](https://waterdata.usgs.gov/monitoring-location/USGS-11406920/#dataTypeId=daily-00060-0&period=periodOfRecord&showFieldMeasurements=true) with TAO data from DWR Feather River database.

```{r, include = F}
feather_tao_query <- dataRetrieval::read_waterdata_daily("USGS-11406920", "00060")
feather_tao_river_daily_flows <- feather_tao_query |> 
  dplyr::select(time, tao_value = value) |> 
  dplyr::as_tibble() |> 
  dplyr::rename(date = time)

# Pulling from excel export of the 2026 database just to be sure
feather_tao_db <- feather_2026_db_export |> 
  group_by(Date) |> 
  summarize(TAO = mean(TAO, na.rm = T))

compare2 <- feather_tao_db |> 
    rename(date = Date) |> 
    select(date, TAO) |> 
    mutate(date = ymd(date)) |> 
    left_join(feather_tao_river_daily_flows) |> 
    mutate(compare = TAO - tao_value)

# There is variation in TAO values
compare2 |> 
    ggplot(aes(x = date, y = TAO)) +
    geom_line()
```

```{r, echo = F, message = F, warning = F}
ck <- compare2 |> 
    ggplot(aes(x = date, y = compare)) +
    geom_line()

plotly::ggplotly(ck)

ck2 <- compare2 |>
    filter(year(date) >= 2011 & year(date) <= 2025) |> 
    mutate(year = as.factor(year(date))) |> 
    ggplot(aes(x = year, y = compare)) +
    geom_boxplot()

ck2
```

# Compare ORF to database

Note that data on USGS goes through 2024 so supplement with CDEC data. Data on CDEC only seem to go through 2022.
Good match so use data from USGS and CDEC.

Plot of ORF flow from DWR database
```{r, echo = F, warning = F, message = F}
# Double check that we are not losing any information by pulling Hatchery and Oroville from CDEC
feather_2026_db_export |> 
    filter(year(Date) > 1990) |> 
    ggplot(aes(x = Date, y = `Hatchery (CDEC ORF; USGS 11406930)`)) +
    geom_point()
```

```{r, include = F}
feather_orf_usgs_raw <- dataRetrieval::read_waterdata_daily("USGS-11406930", "00060")

feather_orf_usgs <- feather_orf_usgs_raw |> 
  dplyr::select(time, value) |> 
  dplyr::as_tibble() |> 
  dplyr::rename(date = time)
feather_orf_usgs |> 
    filter(year(date) > 1990) |> 
    ggplot(aes(x = date, y = value)) +
    geom_point()
feather_orf_cdec <- CDECRetrieve::cdec_query(station = "ORF", dur_code = "D", sensor_num = "41", start_date = "2024-10-01")

feather_orf_usgs_cdec <- feather_orf_usgs |> 
    select(-geometry) |> 
    mutate(gage_agency = "USGS",
    gage_number = "USGS-11406930") |> 
    bind_rows(feather_orf_cdec |> 
    rename(date = datetime,
    value = parameter_value) |> 
        mutate(gage_agency = "CDEC",
        gage_number = "ORF")) |> 
        dplyr::mutate(stream = "feather river",  
         site_group = "upper feather hfc", 
         parameter = "flow")
```

Plot of flow from USGS and CDEC
```{r, echo = F, warning = F, message = F}
feather_orf_usgs_cdec |> 
    filter(year(date) > 1990) |>
    ggplot(aes(x = date, y = value)) +
    geom_point() 
```

# Compare TFB
Note that data on USGS goes through 2023 so supplement with CDEC data. Data on CDEC only seem to go through 2022.

Good match so use data from USGS and CDEC.

Plot of TFB flow from DWR database
```{r, echo = F, warning = F, message = F}
# Double check that we are not losing any information by pulling Hatchery and Oroville from CDEC
feather_2026_db_export |> 
    filter(year(Date) > 1990) |> 
    ggplot(aes(x = Date, y = `Oroville (CDEC TFB; USGS 11406999)`)) +
    geom_point()
```

```{r, include = F}
feather_tfb_usgs_raw <- dataRetrieval::read_waterdata_daily("USGS-11407000", "00060")

feather_tfb_usgs <- feather_tfb_usgs_raw |> 
  dplyr::select(time, value) |> 
  dplyr::as_tibble() |> 
  dplyr::rename(date = time)
feather_tfb_usgs |> 
    
    ggplot(aes(x = date, y = value)) +
    geom_point()
feather_tfb_cdec <- CDECRetrieve::cdec_query(station = "TFB", dur_code = "D", sensor_num = "41", start_date = "2023-10-01")

feather_tfb_usgs_cdec <- feather_tfb_usgs |> 
    select(-geometry) |> 
    mutate(gage_agency = "USGS",
    gage_number = "USGS-11407000") |> 
    bind_rows(feather_tfb_cdec |> 
    rename(date = datetime,
    value = parameter_value) |> 
        mutate(gage_agency = "CDEC",
        gage_number = "TFB")) |> 
        dplyr::mutate(stream = "feather river",  
         site_group = "upper feather hfc", 
         parameter = "flow")
```

Plot of flow from USGS and CDEC
```{r, echo = F, warning = F, message = F}
feather_tfb_usgs_cdec |> 
    filter(year(date) > 1990) |> 
    ggplot(aes(x = date, y = value)) +
    geom_point() 
```


