---
title: "Mill Creek - redd data analysis and interpolation"
author: "Maddee Wiggins (FlowWest)"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 12, 
  fig.height = 12
)

theme_set(
  theme_minimal() +
    theme(
      plot.title = element_text(size = 16, face = "bold"),
      axis.title.x = element_text(size = 14),
      axis.title.y = element_text(size = 14), 
      legend.position = "bottom"
    )
)
```

**Objective**

Compile and standardize a long-term redd survey dataset across Mill Creek sections and years, and to address gaps in survey coverage using an interpolation approach that preserves both annual variation in spawning abundance and spatial patterns in redd distribution.

**Read in raw data**

Data: `adult-holding-redd-and-carcass-surveys_mill-creek_data-raw_Mill Creek SRCS Redd Counts by Section 1997-2020 Reformatted.xlsx`

```{r read_data}
adult_redd_long <- readxl::read_excel(here::here("data-raw", "analysis", "adult-holding-redd-and-carcass-surveys_mill-creek_data-raw_Mill Creek SRCS Redd Counts by Section 1997-2020 Reformatted.xlsx")) |> 
  janitor::clean_names() |> 
  mutate(across(matches("^x\\d{4}_redd_count$"), as.character)) |>
  pivot_longer(
    cols = matches("^x\\d{4}_redd_count$"),
    names_to = "year",
    values_to = "n_redds"
  ) |>
  mutate(
    year = as.integer(str_extract(year, "\\d{4}")),
    n_redds = na_if(n_redds, "n/s"),
    n_redds = na_if(n_redds, "*"),
    n_redds = parse_number(n_redds) 
  ) |>
  select(site = section, year, n_redds) |>
  filter(!str_detect(site, "^total")) |> 
  filter(site != "Total Redds Counted:") 

# Fill in Savercool Place  to Black Rock with McCarthy Place to Savercool Place
# from 1997 to 2006 since those rows were merged in the spreadsheet
adult_redd_long <- adult_redd_long |> 
  left_join(
    adult_redd_long |>
      filter(site == "McCarthy Place to Savercool Place", year %in% 1997:2006) |>
      select(year, n_redds_xxx = n_redds),
    by = "year"
  ) |>
  mutate(
    n_redds = if_else(
      site == "Savercool Place  to Black Rock" & year %in% 1997:2006 & is.na(n_redds),
      n_redds_xxx,
      n_redds
    )
  ) |>
  select(-n_redds_xxx) |> 
  glimpse()
```

**Visualize data gaps**

Years that were not surveyed are shown in grey. We can see that from the late 1990s through mid-2000s, the `Canyon Camp to Sooner Place` reach was not surveyed. 

```{r plot1}
ggplot(adult_redd_long) + 
  geom_tile(aes(x = year, y = site, fill = n_redds)) +
  scale_fill_viridis_c() + 
  labs(
    title = "Redd Counts by Site (raw data)",
    fill = "n_redds"
  )

```

Some survey sites may be more productive than others. If this is true, then missing one reach may affect annual totals more than if another reach (with lower productivity) were missing. The survey sites with the highest counts have good coverage across years (i.e. `Sooner Place to McCarthy Place`, `McCarthy Place to Savercool Place`, and `Savercool Place to Black Rock`), but `Canyon Camp to Sooner Place` sits above the median (`19 redds`). This implies that some sites contribute more redds to the annual total than others, and this spatial variation should be included in any interpolation analysis.

```{r compare_site_averages}
compare_reach_totals <- adult_redd_long |> 
  group_by(site) |> 
  summarise(mean_redds_by_site = round(mean(n_redds, na.rm = T))) |> 
  ungroup()

compare_reach_totals |> 
  arrange(desc(mean_redds_by_site)) |> 
  select(`Survey Reach` = site, `Mean Redds Per Site` = mean_redds_by_site) |> 
  knitr::kable()
```

**Interpolation Method**

Redd counts were not available for all river sections in every year because some sections were not surveyed (`n/s`). To estimate redd abundance in these missing site–year combinations, we used a scaled interpolation approach that maintains the strong year-to-year variation in overall redd production and a site's typical relative productivity.

We implemented the following method: 

1) For each year, we calculated that year's observed mean redds per surveyed site (`mean_obs_per_site`), which reflects basin-wide changes in spawning abundance.
2) For each site, we calculated a `relative_index`: the average of that site's observed counts divided by each corresponding year's `mean_obs_per_site`, then re-centered so the average site has an index of 1. This produces a relative index describing whether a site is typically above or below average, independent of which specific years it happened to be surveyed in.
3) Missing site–year redd counts were filled as `relative_index(site) * mean_obs_per_site(year)`, so a below-average section missing in a strong run year is filled higher than an average section missing in the same year, and vice versa.

## Results

```{r}
n_sites_total <- length(unique(adult_redd_long$site))

annual_totals <- adult_redd_long |>
  group_by(year) |>
  summarise(
    total_redds_obs = sum(n_redds, na.rm = TRUE),
    n_sites_obs     = sum(!is.na(n_redds)),
    .groups = "drop"
  )

annual_fill_value <- annual_totals |>
  mutate(
    mean_obs_per_site = total_redds_obs / n_sites_obs,
    n_sites_missing = n_sites_total - n_sites_obs
  )

# site's typical productivity relative to an average site, with each
# observed year's count first normalized by that year's mean_obs_per_site
# so a site isn't over/under-credited just because its observed years
# happened to be strong or weak run years
relative_index <- adult_redd_long |>
  left_join(annual_fill_value |> select(year, mean_obs_per_site), by = "year") |>
  filter(!is.na(n_redds)) |>
  mutate(site_year_ratio = n_redds / mean_obs_per_site) |>
  group_by(site) |>
  summarise(relative_index = mean(site_year_ratio), .groups = "drop") |>
  mutate(relative_index = relative_index / mean(relative_index))

adult_redd_scaled <- adult_redd_long |>
  left_join(annual_fill_value, by = "year") |>
  left_join(relative_index, by = "site") |>
  mutate(
    fill_value = round(relative_index * mean_obs_per_site, 0),
    n_filled = ifelse(is.na(n_redds), fill_value, n_redds),
    fill_method = ifelse(is.na(n_redds), "site_year_scaled", "observed"),
    fill_confidence = case_when(
      fill_method == "observed" ~ "high",
      n_sites_obs >= 10 ~ "medium",
      n_sites_obs >= 5 ~ "low",
      TRUE ~ "very_low"
    )
  )

annual_summary_filled <- adult_redd_scaled |>
  group_by(year) |>
  summarise(
    total_redds_obs = first(total_redds_obs),
    n_sites_missing = first(n_sites_missing),
    total_redds_filled = sum(n_filled),
    .groups = "drop"
  ) |>
  mutate(
    obs_fraction = total_redds_obs / total_redds_filled,
    redd_multiplier = ifelse(total_redds_obs > 0,
    total_redds_filled / total_redds_obs,
    NA_real_)
  )

knitr::kable(annual_summary_filled |> 
  mutate(obs_fraction = round(obs_fraction, 2),
  redd_multiplier = round(redd_multiplier, 2)))

write_csv(annual_summary_filled, here::here("data-raw", "analysis", "mill_redd_fill_table.csv"))

```

Fill confidence was assigned based on whether redd counts were directly observed or interpolated, and on how many sites were surveyed in that year. Observed values were given a high confidence rating. For interpolated values, confidence was determined by the number of sites surveyed that year: years with at least 10 sites surveyed were assigned medium confidence, years with at least 5 sites surveyed were assigned low confidence, and years with fewer than 5 sites surveyed were assigned very low confidence.

```{r}
ggplot(adult_redd_scaled |> 
         mutate(
           fill_confidence = factor(
             fill_confidence,
             levels = c("high", "medium", "low")
           )
         ),
       aes(x = year, y = site, fill = n_filled)) +
  geom_tile(
    aes(color = fill_confidence),
    linewidth = 1.2,
    width = 0.92,   
    height = 0.92
  ) +
  scale_fill_viridis_c(name = "Redd count") +
  # outline colors for confidence
  scale_color_manual(
    name = "Fill confidence",
    values = c(
      "high"   = "darkgreen",
      "medium" = "orange",
      "low"    = "darkred"
    )
  ) +
  labs(
    title = "Redd Counts by Site (Scaled Interpolation Filled)",
    x = "Year",
    y = "Site"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank()
  )

analysis_summary <- adult_redd_scaled |>
  group_by(site) |>
  mutate(n_years_present = sum(!is.na(n_redds))) |> 
  ungroup() |> 
  group_by(site, n_years_present) |> 
  summarise(total_redds = sum(n_redds, na.rm = TRUE),
            total_redds_with_interpolation = sum(n_filled, na.rm = TRUE), 
            .groups = "drop") 

knitr::kable(analysis_summary)
```

## Integration into final annual adult count

The dataset with `redd_multiplier` for each year is then multiplied by the redd counts provided directly from the stream teams (Data: `data-raw/helper-tables/mill_deer_adult_historical.csv`).