---
title: "Data Checks for Butte Creek"
output: html_document
date: "2025-05-27"
---

```{r setup, include=FALSE}
library(tidyverse)
library(SRJPEdata)
library(lubridate)

colors_full <-  c("#9A8822", "#F5CDB4", "#F8AFA8", "#FDDDA0", "#74A089", #Royal 2
                  "#899DA4", "#C93312", "#DC863B", # royal 1 (- 3)
                  "#F1BB7B", "#FD6467", "#5B1A18", # Grand Budapest 1 (-4)
                  "#D8B70A", "#02401B", "#A2A475", # Cavalcanti 1
                  "#E6A0C4", "#C6CDF7", "#D8A499", "#7294D4", #Grand Budapest 2
                  "#9986A5", "#EAD3BF", "#AA9486", "#B6854D", "#798E87" # Isle of dogs 2 altered slightly
)


filepath <- here::here("data-raw", "data-checks", "stream_team_review", "butte", "weekly_catch_b_GHCamp_vs_FW.xlsx")
gh_2015 <- readxl::read_xlsx(filepath, sheet = "2015")
gh_2016 <- readxl::read_xlsx(filepath, sheet = "2016")
gh_2018 <- readxl::read_xlsx(filepath, sheet = "2018")
gh_2019 <- readxl::read_xlsx(filepath, sheet = "2019")
gh_2020 <- readxl::read_xlsx(filepath, sheet = "2020")
gh_2021 <- readxl::read_xlsx(filepath, sheet = "2021")
gh_2022 <- readxl::read_xlsx(filepath, sheet = "2022")
gh_2023 <- readxl::read_xlsx(filepath, sheet = "2023")
gh_2024 <- readxl::read_xlsx(filepath, sheet = "2024")

gh_combined <- bind_rows(gh_2015,gh_2016,
                         gh_2018, gh_2019,
                         gh_2020, gh_2021,
                         gh_2022, gh_2023,
                         gh_2024)

temp_catch <- read_csv(here::here("data-raw", "data-checks", "stream_team_review", "temp_catch0527.csv"))

# filter to chinook
weekly_edi_catch1 <- temp_catch |> 
  mutate(week = week(date),
         year = year(date),
         species = tolower(species)) |> 
  filter(species %in% c("chinook", "chinook salmon") | is.na(species)) |> 
  group_by(stream, site, week, year) |> 
  summarize(count = sum(count, na.rm = T))

weekly_rst_catch <- SRJPEdata::rst_catch |> 
  mutate(week = week(date),
         year = year(date)) |> 
  group_by(stream, site, week, year) |> 
  summarize(count = sum(count, na.rm = T))

butte_compare3 <- gh_combined |> 
  rename(year = Year,
         week = julianWeek) |> 
  full_join(SRJPEdata::weekly_juvenile_abundance_catch_data |> 
              filter(stream == "butte creek", site == "okie dam") |> 
              select(year, week, count)) |> 
  mutate(p_diff = ((abs(SumOfn - count))/((SumOfn + count)/2)) * 100)
write_csv(butte_compare3, here::here("data-raw", "data-checks", "stream_team_review", "butte", "butte_compare3.csv"))

butte_compare2 <- gh_combined |> 
  rename(year = Year,
         week = julianWeek) |> 
  full_join(weekly_rst_catch |> 
              filter(stream == "butte creek", site == "okie dam") |> 
              select(year, week, count)) |> 
  mutate(p_diff = ((abs(SumOfn - count))/((SumOfn + count)/2)) * 100)
write_csv(butte_compare2, here::here("data-raw", "data-checks", "stream_team_review", "butte", "butte_compare2.csv"))
butte_compare1 <- gh_combined |> 
  rename(year = Year,
         week = julianWeek) |> 
  full_join(weekly_edi_catch1 |> 
              filter(stream == "butte creek", site == "okie dam") |> 
              select(year, week, count)) |> 
  mutate(p_diff = ((abs(SumOfn - count))/((SumOfn + count)/2)) * 100)
write_csv(butte_compare1, here::here("data-raw", "data-checks", "stream_team_review", "butte", "butte_compare1.csv"))
```

I compare our most up to date SRJPEdata with the excel sheet provided by Grant. I
do step by step comparisons to identify where discrepancies are coming from.

1. Compare with data straight from EDI (summarized by week and filted to Chinook)
2. Compare with SRJPEdata::rst_catch (filted to Chinook, remove adults, remove fyke only dates)
3. Compare with SRJPEdata::weekly_juvenile_abundance_catch_data (in addition to rst_catch filters, remove yearlings)

**Action Items**

- My biggest priority is to figure out why there are differences when we compare to EDI. This is 2 weeks.
- Additionally I want to confirm my assumptions about the discrepancies in comparison 2 and 3 with Grant.


## Comparison 1 (data from EDI)

There are 2 weeks with differences greater than 5%
- 2021 week 6
- 2023 week 52

```{r, echo = F, warning = F, message = F}
butte_compare1 |> 
  filter(year > 2014) |> 
  ggplot(aes(x = week, y = p_diff)) +
  geom_point() +
  scale_color_manual(values = colors_full) +
  facet_wrap(~year) +
  theme_bw() +
  labs(y = "percent difference")
```

```{r, echo = F, warning = F, message = F}
filter(butte_compare1, p_diff > 5)
```

```{r, include = F}
filter(temp_catch, site == "okie dam", year(date) == 2021, week(date) == 6) |> view()
```

## Comparison 2 (data from rst_catch)

There are 41 weeks with differences.

- Is this coming from the removal of adults? Maybe in some cases
- Is this coming from the removal of fyke only trapping? Yes in some cases. For instance, year 2018, week 11: Grant has 1471 and we have 1267. Only the fyke was fishing on 2018-03-13 and 2018-03-14 and the catch on those days was 204 which is the difference between our two values.

```{r, echo = F, warning = F, message = F}
butte_compare2 |> 
  filter(year > 2014) |> 
  ggplot(aes(x = week, y = p_diff)) +
  geom_point() +
  scale_color_manual(values = colors_full) +
  facet_wrap(~year) +
  theme_bw() +
  labs(y = "percent difference")
```

```{r, echo = F, warning = F, message = F}
filter(butte_compare2, p_diff > 5)
```

```{r, include = F}
filter(temp_catch, site == "okie dam", year(date) == 2018, week(date) == 11) |> view()

filter(temp_catch, site == "okie dam", (date == as_date("2018-03-13") | date == as_date("2018-03-14"))) |> summarize(sum = sum(count))
```

## Comparison 3 (data from weekly_juvenile_abundance_catch_data)

There are 52 weeks with differences.

The differences would be the same as for rst_catch with some additional discrepancies from the removal of yearlings.

```{r, echo = F, warning = F, message = F}
butte_compare3 |> 
  filter(year > 2014) |> 
  ggplot(aes(x = week, y = p_diff)) +
  geom_point() +
  scale_color_manual(values = colors_full) +
  facet_wrap(~year) +
  theme_bw() +
  labs(y = "percent difference")
```

```{r, echo = F, warning = F, message = F}
filter(butte_compare3, p_diff > 5)
```
